### 2024-04-26 paper/main.tex
Add paper

\documentclass{article}
\include{setup}

\title{Empirical study of an algorithm for finding Hamiltonian cycles in 3-regular planar graphs.}
\author{Kristoffer Højelse}
\date{February 2024}

\begin{document}

\maketitle

\begin{abstract}
	An algorithm in the making considers the problem of finding Hamiltonian cycles in 3-regular planar graphs.
\end{abstract}

\section{Description of the problem}

	\textsc{The Hamiltonian Cycle Problem} is a computational problem that comes in a few variants, with slightly different input and output types.

	\textsc{The Hamiltonian Cycle Decision problem} asks given a simple graph, to output; yes or no to the question; Does there exist a Hamiltonian Cycle in the graph? Variants might ask for an example/witness/certificate of one such cycle, to make verifying the result easy.

	\textsc{The Hamiltonian Cycle Counting problem} asks given a simple graph, to output; the number of unique Hamiltonian cycles. Solving this problem you get the decision variant for free.

	\begin{definition}
		\textsc{The Hamiltonian Cycle Counting Problem}

		Input: Given a simple undirected connected graph $G$.

		Output: The number of unique Hamiltonian cycles.

		A pair of Hamiltonian cycles are not unique if and only if you can obtain the other by circular permutations and inversions of the associated vertex sequences.
	\end{definition}

	The simplest algorithm I can come up with that solves this problem has a factorial time complexity $\Theta(N!)$.

	Enumerate all $N!$ permutations of the $N$ vertices and count how many of those vertex sequences are valid Hamiltonian cycles, by checking firstly that the foremost and final vertex of the sequence are adjacent and secondly that every consecutive pair of vertices in the sequence are adjacent. Then return this count not before dividing by a factor of $2N$, as there are $N$ circular permutations and $2$ inversions of a sequence of $N$ elements.

	A Python program of this algorithm is in the appendix \ref{appx:brute-force-count-ham-cyc}.

	You cannot do much better\ref{} for this general problem. But if we constrain the input we can do much better.

\section{The domain}
	The algorithm described in this paper solves \textsc{The Hamiltonian Cycle Counting Problem} with a very contained subset of all graphs; simple undirected connected planar cubic graphs.

	Some informal definitions of the aforementioned properties of graphs:

	\begin{enumerate}
		\item A graph is called \textbf{simple} if and only if it has no parallel edges and no self-loops.
		\item A graph is called \textbf{undirected} if and only if all its edges can be traversed in both directions.
		\item A graph is called \textbf{connected} (or 1-vertex-connected) if and only if there exists a path between any two vertices.
		\item A graph is called \textbf{planar} if and only if there's a way to draw the graph in 2 dimensions such that no pair of edges crosses.
		\item A graph is called \textbf{cubic} if and only if every vertex has exactly three neighbors.
	\end{enumerate}

	Most variants of \textsc{The Hamiltonian Cycle Problem} too require the input graph to be simple, undirected, and connected. These problems probably have no known polynomial-time solutions.

	The algorithm exploits the two remaining attributes, planar and cubic, to obtain a polynomial running time.

\section{Description of the algorithm}

	Overview:

	\begin{enumerate}
		\item Count Hamiltonian Cycles by Dynamic Programming over a minimal Branch Decomposition.
		\item Compute a minimal Branch Decomposition from a minimal Carving Decomposition of the Medial graph.
		\item Compute a minimal Carving Decomposition by repeatedly Contracting edges that doesn't increase the Carving Width.
		\item Compute the Carving Width with the Rat Catching algorithm which needs to compute a dual graph.
		\item Compute a dual graph.
	\end{enumerate}

	\subsection{Dual graph}

		A subroutine of the algorithm computes the dual graph of a graph.

		\begin{definition}
			The dual graph $G*$ of a planar graph $G$ is a graph with a vertex $f*$ for each face $f$ of $G$ and an edge $e*$ for each edge $e$ that separates a face $f_1$ of $G$ and a face $f_2$ of $G$.
		\end{definition}

		\begin{corollary}
			If multiple edges separate $f_1$ and $f_2$ there will be multiple edges between $f_1*$ and $f_2*$.
		\end{corollary}

		\begin{corollary}\label{def:dual_coro_loop}
			If $e$ separates $f_1$ and $f_2$ and are the same face, $e*$ will be a self-loop.
		\end{corollary}

		For the algorithm in this paper, the class of graphs that will be given as input is medial graphs of simple undirected connected planar cubic graphs.

		I therefore claim, for now without any proof or argument, that;

		\begin{claim}
			Corollary \ref{def:dual_coro_loop} will be irrelevant for any implementation of the algorithm.
		\end{claim}

	\subsection{Medial graph}

		A subroutine of the algorithm computes the medial graph of a graph from the class; simple connected planar cubic graphs.

		\begin{definition}
			The medial graph $M(G)$ of a connected plane graph $G$ with a vertex $e*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $c*$ between a pair of vertices $e_1*$, $e_2*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
		\end{definition}

		todo: argue that simple connected planar cubic graphs are a valid substitution for a connected plane.

		todo: identify a more succinct description of the class; medial of simple connected planar cubic graphs

\section{Appendix.}

\subsection{Brute force count Hamiltonian Cycles}\label{appx:brute-force-count-ham-cyc}

\end{document}

### 2024-04-26 paper/setup.tex
Add paper

% General
	\usepackage{graphicx} % Required for inserting images
	\setlength{\parindent}{0pt}
	\setlength{\parskip}{8pt} % Vertical space between paragraphs
	\setlength{\baselineskip}{18pt} % Vertical space between lines
	\usepackage{amsthm}
	\usepackage{mdframed}
	\usepackage{hyperref} % Add the hyperref package for clickable links
	\newcommand\bigO{\mathrm{O}}

% Define the theorem-like environments
	\newtheorem{theorem}{Theorem}[section]  % Numbering within sections
	\newtheorem{definition}[theorem]{Definition}  % Numbering with theorems
	\newtheorem{corollary}[theorem]{Corollary}  % Numbering with theorems
	\newtheorem{claim}[theorem]{Claim}  % Numbering with theorems
	\newtheorem{lemma}[theorem]{Lemma}  % Numbering with theorems
	\newtheorem{axiom}[theorem]{Axiom}  % Numbering with theorems
	\newtheorem{conjecture}[theorem]{Conjecture}  % Numbering with theorems

% Define a box style for the theorem-like environments
	\mdfdefinestyle{theoremstyle}{%
		linewidth=0pt,
		backgroundcolor=white,
		roundcorner=5pt,
		innertopmargin=6pt,
		innerbottommargin=0pt,
		innerrightmargin=32pt,
		innerleftmargin=16pt,
	}	
% Enclose theorem-like environments within boxes
	\surroundwithmdframed[style=theoremstyle]{theorem}
	\surroundwithmdframed[style=theoremstyle]{definition}
	\surroundwithmdframed[style=theoremstyle]{corollary}
	\surroundwithmdframed[style=theoremstyle]{claim}

% Set margins
	\usepackage{geometry}
	\geometry{
		left=3cm,
		right=3cm,
		top=2cm,
		bottom=2cm
	}

% Codeblock
	\usepackage{minted}
	\usepackage{xcolor}
	\definecolor{mintedBg}{rgb}{0.97,0.97,0.98}
	\setminted[py]{
		bgcolor=mintedBg,
		framesep=3mm,
		fontsize=\footnotesize,
		linenos,
		breaklines,
	}
	\newcommand{\pyfile}[1]{%
		\begin{filecontents*}{#1}
		\end{filecontents*}
		\noindent\textbf{#1}\par
		\inputminted{py}{#1}%
	}

### 2024-04-26 paper/build.sh

@@ -1,0 +1,1 @@
+latexmk -pvc -shell-escape -pdf -output-directory="./build" main.tex

### 2024-04-26 gen_1-connected_simple_cubic_planar.py
Move stuff

@@ -1,74 +1,0 @@
-import random
-import sys
-
-VERTEX_COUNT = int(sys.argv[1])
-
-N = 3
-adj = {
-	1: [2, 3],
-	2: [3, 1],
-	3: [1, 2]
-}
-faces = [(1, 2, 3)]
-
-dual = {
-	(3,2,1): [(1,2,3)],
-	(1,2,3): [(3,2,1)]
-}
-
-while len(faces)+1 <= VERTEX_COUNT-2:
-	# choose random face
-	face = faces.pop(random.randint(0, len(faces)-1))
-
-	N += 1
-
-	# insert new vertex into face
-	adj[N] = list(face)
-
-	# insert neighbor in clockwise order
-	a = face[0]
-	b = face[1]
-	c = face[2]
-	adj[a].insert(adj[a].index(b)+1%3, N)
-	adj[b].insert(adj[b].index(c)+1%3, N)
-	adj[c].insert(adj[c].index(a)+1%3, N)
-
-	# append new faces
-	new_faces = [
-		(a, b, N),
-		(b, c, N),
-		(c, a, N)
-	]
-	for new_face in new_faces:
-		faces.append(new_face)
-
-	# update dual
-	for neighbor_face in dual[face]:
-		face_idx = dual[neighbor_face].index(face)
-		for new_face in new_faces:
-			if new_face[0] in neighbor_face and new_face[1] in neighbor_face:
-				dual[neighbor_face].insert(face_idx, new_face)
-			elif new_face[1] in neighbor_face and new_face[2] in neighbor_face:
-				dual[neighbor_face].insert(face_idx, new_face)
-			elif new_face[2] in neighbor_face and new_face[0] in neighbor_face:
-				dual[neighbor_face].insert(face_idx, new_face)
-		dual[neighbor_face].remove(face)
-
-	dual[new_faces[0]] = [new_faces[1], new_faces[2], *list(filter(lambda neighbor_face: a in neighbor_face and b in neighbor_face, dual[face]))]
-	dual[new_faces[1]] = [new_faces[2], new_faces[0], *list(filter(lambda neighbor_face: b in neighbor_face and c in neighbor_face, dual[face]))]
-	dual[new_faces[2]] = [new_faces[0], new_faces[1], *list(filter(lambda neighbor_face: c in neighbor_face and a in neighbor_face, dual[face]))]
-
-	dual.pop(face, None)
-
-
-faces = [(3,2,1), *faces]
-
[…]

### 2024-04-26 gen_3-connected_simple_cubic_planar.py

@@ -1,89 +1,0 @@
-# Author: Andreas Björklund
-from random import random
-import sys
-
-if len(sys.argv)>1:
-  DEPTH=(int(sys.argv[1])-4)//2
-else:  
-  DEPTH=5
-FRAC=0.9
-
-def start_instance():
-  # Original graph K_4  
-  #   0
-  #
-  #   3
-  #1     2
-  C=[[2,3,1], [0,3,2], [1,3,0], [0,2,1]] 
-  return C  
-                         
-# Generate random cubic 3-connected planar graph
-def random_instance(C):
- 
-  # Triangle
-  D=[[0,4,5],[1,5,3],[2,3,4]]
-  
-  # Single Bridge
-  # 041
-  # 253
-  F=[[0,1,5],[2,4,3]]
-  
-  B = C[:]
-
-  for rounds in range(1):
-    n=len(B)
-    who = int(random()*n)
-    if random() < FRAC:
-      nx = int(random()*3)
-      cycle=[who]
-      while B[who][nx]!=cycle[0]:
-        cycle.append(B[who][nx])
-        ix=[x for x in range(3) if B[B[who][nx]][x]==who]
-        ix=(ix[0]+2)%3
-        who = B[who][nx]
-        nx = ix
-    
-      m=len(cycle)
-      p = int(random()*(m-3))+2;
-      T = cycle[:2] + cycle[p:p-2:-1]
-      G = B[:]
-      H = [(list(map(sum, zip(x, [n-4,n-4,n-4])))) for x in F]
-      for i in range(4):
-        for ix2 in range(3):
-          if B[T[i]][ix2]==T[i^1]:
-            G[T[i]]= (G[T[i]][:ix2]+[n+(i // 2)]+G[T[i]][ix2+1:])
-        for j in range(len(F)):
-          for k in range(3):
-            if F[j][k]==i:
-              H[j] = (H[j][:k]+[T[i]]+H[j][k+1:])
-      B=G+H
-    else:
-      H = [(list(map(sum, zip(x, [n-3,n-3,n-3])))) for x in D]
-      G = B[:]
-      for i in range(3):
-        for ix2 in range(3):
-          if B[B[who][i]][ix2]==who:
-            G[B[who][i]] = (G[B[who][i]][0:ix2]+[n+i]+G[B[who][i]][ix2+1:])
-        for j in range(len(D)):
-          for k in range(3):
-            if D[j][k]==i:
[…]

### 2024-04-26 gen_all_n_vertex_graphs.sh

@@ -1,1 +1,0 @@
-./plantri.bin -c1d $(($1/2+2)) | python3 process_planar_code.py

### 2024-04-26 process_planar_code.py

@@ -1,24 +1,0 @@
-import sys
-
-# skip header
-byte_count = 0
-while byte_count < 15:
-	byte_s = sys.stdin.buffer.read(1)
-	if not byte_s:
-		break
-	byte_count += 1
-
-# read graphs
-graph_idx = 1
-while True:
-	byte_N = sys.stdin.buffer.read(1)
-	if not byte_N:
-		exit()
-	N = int.from_bytes(byte_N, byteorder='little')
-
-	adj = sys.stdin.buffer.read(N*4)
-	
-	with open(f"./graphs/{N}/{graph_idx}.bin", "xb") as outfile:
-		outfile.write(byte_N)
-		outfile.write(adj)
-	graph_idx += 1

### 2024-04-26 code/count_hamcyc_brute_force.py
Update alot of stuff

@@ -1,7 +1,7 @@
 import itertools
-from parse_graph import parse_graph_to_adj
+from parse_graph import parse_text_to_adj
 
-G: dict[int, list[int]] = parse_graph_to_adj()
+G: dict[int, list[int]] = parse_text_to_adj()
 vertex_set = G.keys()
 N = len(vertex_set)
 

### 2024-04-26 code/dual_graph.py

@@ -1,51 +1,67 @@
 from parse_graph import parse_text_to_adj, adj_to_text
 
 # assume planar graph
+# assume not reflexive graph
 # assume clockwise or counterclockwise ordering
-def find_faces(adj: dict[int, list[int]]) -> list[list[tuple[int, int]]]:
-	edges = set([(i, j) for i in adj for j in adj[i]])
-	faces = []
-	while len(edges) > 0:
-		first_edge = edges.pop()
-		v = first_edge[1]
-		face = [first_edge]
-		next_edge = first_edge
-		while True:
-			u = adj[next_edge[1]][(adj[next_edge[1]].index(next_edge[0])-1)%len(adj[next_edge[1]])]
-			next_edge = (v, u)
-			if next_edge == first_edge:
-				faces.append(face)
-				break
-			edges.remove(next_edge)
-			face.append(next_edge)
-			v = u
-	return faces
-
 def dual_graph(G: dict[int, list[int]]):
-	faces = [tuple(f) for f in find_faces(G)]
-	nodes = [i+1 for i in range(len(faces))]
-	node_to_face = dict([(i+1, f) for i,f in enumerate(faces)])
-	face_to_node = dict([(f, i+1) for i,f in enumerate(faces)])
+	G2 = dict([(x, ys.copy()) for x,ys in G.items()])
 
-	edge_to_node = dict([(e, node) for node,face in node_to_face.items() for e in face])
+	# rotate neighbors such that parallel edges are consecutive
+	def rotated(l, n):
+		return l[n:] + l[:n]
+	for x,ys in G2.items():
+		for i,y in enumerate(ys):
+			if y != ys[0]:
+				G2[x] = rotated(ys, i)
+				break
 
-	edge_to_link = dict()
-	link_to_edges = dict()
-	dual = dict([(i, []) for i in nodes])
-	for face in faces:
-		for (u, v) in face:
-			dual[edge_to_node[(u,v)]].append(edge_to_node[(v,u)])
-			edge_to_link[tuple(sorted((u,v)))] = tuple(sorted((edge_to_node[(u,v)], edge_to_node[(v,u)])))
+	# assign edge ids
+	G3 = dict([(x, ys.copy()) for x,ys in G2.items()])
+	def nth_to_last_idx_of(l, x, n):
+		return len(l) - 1 - (l[::-1].index(x) + n)
+	next_edge_id = 1
+	edges = dict()
[…]

### 2024-04-26 code/is_connected.py

@@ -1,0 +1,29 @@
+from parse_graph import parse_text_to_adj
+
+G = parse_text_to_adj()
+
+def is_connected(G):
+	# make directed graph undirected
+	es = set()
+	for x,ys in G.items():
+		for y in ys:
+			es.add(tuple(sorted([x,y])))
+
+	component_count = 0
+	# DFS
+	visited = set()
+	for v in G.keys():
+		if v not in visited:
+			stack = [v]
+			visited = set()
+			while stack:
+				v = stack.pop()
+				visited.add(v)
+				for u in G[v]:
+					if u not in visited:
+						stack.append(u)
+			component_count += 1
+	
+	return component_count == 1
+
+print(is_connected(G))

### 2024-04-26 code/is_cubic.py

@@ -1,0 +1,13 @@
+import networkx as nx
+from parse_graph import parse_text_to_adj, adj_to_nx
+
+G = adj_to_nx(parse_text_to_adj())
+
+def is_3_regular(G):
+	for v in G.nodes:
+		xs = list(G[v])
+		if len(xs) != 3:
+			return False
+	return True
+
+print(is_3_regular(G))

### 2024-04-26 code/is_planar.py

@@ -1,0 +1,5 @@
+import networkx as nx
+from parse_graph import parse_text_to_adj, adj_to_nx
+
+G = adj_to_nx(parse_text_to_adj())
+print(nx.is_planar(G))

### 2024-04-26 code/is_simple.py

@@ -1,0 +1,22 @@
+import networkx as nx
+from parse_graph import parse_text_to_adj, adj_to_nx
+
+G = adj_to_nx(parse_text_to_adj())
+
+def is_simple(G):
+	for v in G.nodes:
+		es = list(G.edges(v))
+
+		# check for self loops
+		if (v, v) in es:
+			return False
+
+		# check for parallel edges
+		e = es[0]
+		for e2 in es[1:]:
+			if e2 == e:
+				return False
+			e = e2
+	return True
+
+print(is_simple(G))

### 2024-04-26 code/is_undirected.py

@@ -1,0 +1,12 @@
+from parse_graph import parse_text_to_adj
+
+G = parse_text_to_adj()
+
+def is_undirected(G):
+	for x,ys in G.items():
+		for y in ys:
+			if x not in G[y]:
+				return False
+	return True
+
+print(is_undirected(G))

### 2024-04-26 code/medial_graph.py

@@ -15,9 +15,9 @@ def medial_graph(G: dict[int, list[int]]):
 			medial[nodes[i]].append(nodes[(i+1)%len(nodes)])
 	return medial, node_to_edge, edge_to_node
 
-# G = parse_text_to_adj()
-# medial, node_to_edge, edge_to_node = medial_graph(G)
-# adj_to_text(medial)
-# print(len(edge_to_node))
-# for e,x in edge_to_node.items():
-# 	print(e[0], e[1], x)
+G = parse_text_to_adj()
+medial, node_to_edge, edge_to_node = medial_graph(G)
+adj_to_text(medial)
+print(len(edge_to_node))
+for e,x in edge_to_node.items():
+	print(e[0], e[1], x)

### 2024-04-26 code/parse_graph.py

@@ -1,0 +1,45 @@
+import networkx as nx
+import sys
+
+def parse_bin_to_adj() -> dict[int, list[int]]:
+	adj = dict()
+	n = ord(sys.stdin.buffer.read(1))
+	for i in range(1, n+1):
+		adj[i] = []
+	i = 1
+	while i <= n:
+		x = ord(sys.stdin.buffer.read(1))
+		if x == 0:
+			i += 1
+			continue
+		adj[i].append(x)
+	return adj
+
+def parse_text_to_adj() -> dict[int, list[int]]:
+	adj = dict()
+	n = int(input())
+	for _ in range(n):
+		ys = list(map(int, input().split()))
+		x = ys[0]
+		adj[x] = []
+		for y in ys[1:]:
+			adj[x].append(y)
+	return adj
+
+def adj_to_text(adj):
+	print(len(adj))
+	for v,xs in adj.items():
+		print(v, *xs)
+
+def adj_to_nx(adj):
+	G = nx.MultiDiGraph()
+	for v,xs in adj.items():
+		G.add_node(v)
+		for x in xs:
+			G.add_edge(v, x)
+	return G
+
+def adj_to_bytes(adj):
+	print(chr(len(adj)), end="")
+	for v,xs in adj.items():
+		print("".join(map(chr, [*xs])), end="\x00")

### 2024-04-26 code/util/andreas_to_bin.py

@@ -1,5 +1,5 @@
 import sys
-from parse_graph import adj_to_bytes
+from code.parse_graph import adj_to_bytes
 
 adj = dict()
 

### 2024-04-26 code/util/bin_to_adj.py

@@ -1,4 +1,4 @@
-from parse_graph import parse_graph_to_adj
+from code.parse_graph import parse_graph_to_adj
 
 for x,ys in parse_graph_to_adj().items():
 	print(x, ' '.join(map(str, ys)))

### 2024-04-26 code/util/create_png.py

@@ -1,7 +1,7 @@
 import matplotlib.pyplot as plt
 import networkx as nx
 import sys
-from parse_graph import parse_graph_to_nx
+from code.parse_graph import parse_graph_to_nx
 
 G = parse_graph_to_nx()
 

### 2024-04-26 code/util/is_isomorphic.py

@@ -1,5 +1,5 @@
 import networkx as nx
-from parse_graph import parse_graph_to_nx
+from code.parse_graph import parse_graph_to_nx
 
 G1 = parse_graph_to_nx()
 

### 2024-04-26 code/util/parse_graph.py

@@ -1,44 +1,0 @@
-import networkx as nx
-import sys
-
-def parse_graph_to_nx():
-	G = nx.MultiDiGraph()
-
-	n = ord(sys.stdin.buffer.read(1))
-	for i in range(n):
-		G.add_node(i+1)
-	i = 1
-	while i <= n:
-		x = ord(sys.stdin.buffer.read(1))
-		if x == 0:
-			i += 1
-			continue
-		G.add_edge(i, x)
-	return G
-
-def adj_to_nx(adj):
-	G = nx.MultiDiGraph()
-	for v,xs in adj.items():
-		G.add_node(v)
-		for x in xs:
-			G.add_edge(v, x)
-	return G
-
-def parse_graph_to_adj() -> dict[int, list[int]]:
-	adj = dict()
-	n = ord(sys.stdin.buffer.read(1))
-	for i in range(1, n+1):
-		adj[i] = []
-	i = 1
-	while i <= n:
-		x = ord(sys.stdin.buffer.read(1))
-		if x == 0:
-			i += 1
-			continue
-		adj[i].append(x)
-	return adj
-
-def adj_to_bytes(adj):
-	print(chr(len(adj)), end="")
-	for v,xs in adj.items():
-		print("".join(map(chr, [*xs])), end="\x00")

### 2024-04-26 code/util/print_bin.py

@@ -1,4 +1,4 @@
-from parse_graph import adj_to_bytes
+from code.parse_graph import adj_to_bytes
 adj = {
 	1: [2,3,4,5,6],
 	2: [3,1,6],

### 2024-04-26 code/util/render_dual.py

@@ -1,4 +1,4 @@
-from parse_graph import parse_graph_to_adj, adj_to_nx
+from code.parse_graph import parse_graph_to_adj, adj_to_nx
 from dual_graph import dual_graph
 import matplotlib.pyplot as plt
 import networkx as nx

### 2024-04-26 code/util/render_medial.py

@@ -1,4 +1,4 @@
-from parse_graph import parse_graph_to_adj, adj_to_nx
+from code.parse_graph import parse_graph_to_adj, adj_to_nx
 from medial_graph import medial_graph
 import matplotlib.pyplot as plt
 import networkx as nx

### 2024-04-26 code/util/render_planar.py

@@ -1,6 +1,6 @@
 import matplotlib.pyplot as plt
 import networkx as nx
-from parse_graph import parse_graph_to_nx
+from code.parse_graph import parse_graph_to_nx
 
 G = parse_graph_to_nx()
 

### 2024-04-26 paper/main.tex
Add paper

@@
 
-	A Python program of this algorithm is in the appendix \ref{appx:brute-force-count-ham-cyc}.
+	A Python program of this algorithm is in the appendix \ref{appx:count-hamcyc-brute-force}.
 
@@
 
-\subsection{Brute force count Hamiltonian Cycles}\label{appx:brute-force-count-ham-cyc}
+\subsection{Count Hamiltonian Cycles with Brute Force}
+\label{appx:count-hamcyc-brute-force}
+\pyfile{count_hamcyc_brute_force.py}
 

### 2024-04-26 paper/build.sh

@@ -1,1 +1,1 @@
-latexmk -pvc -shell-escape -pdf -output-directory="./build" main.tex
+latexmk -pvc -shell-escape -pdf main.tex

### 2024-05-21 code/Graph.py
Refactor

@@ -19,9 +19,6 @@ class Graph:
 
 					next_edgeid += 1
 
-	def incident_edges(self, v: int) -> list[int]:
-		return self.adj_edges[v]
-
 	def V(self) -> list[int]:
 		return list(self.adj_edges.keys())
 	

### 2024-05-21 code/branch_decomposition.py

@@ -1,0 +1,24 @@
+from parse_graph import parse_text_to_adj
+from medial_graph import medial_graph
+from carving_decomposition import carving_decomposition
+
+# Construct a branch decomposition of a graph
+def branch_decomposition(G_adj: dict[int, list[int]]):
+	# Contruct the carving decomposition of the medial graph
+	M, node_to_vertexpair, vertexpair_to_node = medial_graph(G_adj)
+	cd = carving_decomposition(M)
+
+	# Convert the carving decomposition of M to a branch decomposition of G
+
+	def decomp(t):
+		if isinstance(t, int):
+			return node_to_vertexpair[t]
+		return tuple([decomp(a) for a in t])
+	
+	bd = decomp(cd)
+	return bd
+
+if __name__ == "__main__":
+	adj = parse_text_to_adj()
+	bd = branch_decomposition(adj)
+	print(bd)

### 2024-05-21 code/branch_width.py

@@ -1,80 +1,23 @@
-from carving_width import carving_width
-from contraction import contraction
+from branch_decomposition import branch_decomposition
 from parse_graph import parse_text_to_adj, adj_to_text
-from dual_graph import dual_graph
-from medial_graph import medial_graph
-from Graph import Graph
-
-# Find a contraction that does not increase the carving width
-def nonincreasing_cw_contraction(G: Graph, cw1):
-	for x,ys in G.adj().items():
-		for y in ys:
-			if x < y:
-				G2, z = contraction(G, x, y)
-				cw2 = carving_width(G2)
-				if cw2 <= cw1:
-					return G2, (x, y), cw2, z
-	return None, None, None, None
-
-# Contract edges that do not increase the carving width
-# until only 3 vertices remain
-def gradient_descent_contractions(G: Graph) -> Graph:
-	G2 = G.copy()
-	cw1 = carving_width(G)
-	# print("cw1", cw1)
-	# exit()
-	edges = dict()
-	while True:
-		G3, e, cw2, z = nonincreasing_cw_contraction(G2, cw1)
-		if G3 is not None and len(G3.V()) >= 3:
-			G2 = G3
-			cw1 = cw2
-			edges[z] = e
-		if len(G2.V()) == 3:
-			return G2, edges
-
-def carving_decomposition(G: Graph) -> tuple:
-	G2, edges = gradient_descent_contractions(G)
-
-	def decomp(x):
-		if x not in edges:
-			return x
-		a,b = edges[x]
-		return (decomp(a), decomp(b))
-
-	a,b,c = G2.V()
-	cd = (decomp(a), decomp(b), decomp(c))
-	return cd
-
-def branch_decomposition(adj: dict[int, list[int]]):
-	M, node_to_vertexpair, vertexpair_to_node = medial_graph(adj)
-	cd = carving_decomposition(M)
-
-	def decomp(x):
-		if isinstance(x, int):
-			return node_to_vertexpair[x]
-		return tuple([decomp(a) for a in x])
-	
-	bd = decomp(cd)
-	return bd
 
 def branch_width_of_branch_decomposition(bd):
-	t = dict()
-
+	# Create an adjacency list from the branch decomposition
+	T_adj = dict()
 	def aux(subtree, depth, name):
 		if len(subtree) == 2 and isinstance(subtree[0], int) and isinstance(subtree[1], int):
-			t[subtree] = []
+			T_adj[subtree] = []
[…]

### 2024-05-21 code/carving_decomposition.py

@@ -1,0 +1,53 @@
+from carving_width import carving_width
+from contraction import contraction
+from parse_graph import parse_text_to_adj, adj_to_text
+from dual_graph import dual_graph
+from medial_graph import medial_graph
+from Graph import Graph
+
+# Find a contraction that does not increase the carving width
+def nonincreasing_cw_contraction(G: Graph, cw1):
+	for es in G.E():
+		u, v = G.edge_to_vertexpair[es]
+		G2, w = contraction(G, u, v)
+		cw2 = carving_width(G2)
+		if cw2 <= cw1:
+			return G2, (u, v), cw2, w
+	return None, None, None, None
+
+# Contract edges that do not increase the carving width
+# until only 3 vertices remain.
+# Return the resulting graph and the edges that were contracted
+def gradient_descent_contractions(G: Graph) -> Graph:
+	G2 = G.copy()
+	cw1 = carving_width(G)
+	edges = dict()
+	while True:
+		G3, e, cw2, w = nonincreasing_cw_contraction(G2, cw1)
+		if G3 is not None and len(G3.V()) >= 3:
+			G2 = G3
+			cw1 = cw2
+			edges[w] = e
+		if len(G2.V()) == 3:
+			return G2, edges
+
+# Contruct a carving decomposition of a graph
+def carving_decomposition(G: Graph) -> tuple:
+	G2, edges = gradient_descent_contractions(G)
+
+	# Construct the decomposition from the edges that were contracted
+
+	def decomp(x):
+		if x not in edges:
+			return x
+		a,b = edges[x]
+		return (decomp(a), decomp(b))
+
+	a,b,c = G2.V()
+	cd = (decomp(a), decomp(b), decomp(c))
+	return cd
+
+if __name__ == "__main__":
+	adj = parse_text_to_adj()
+	cd = carving_decomposition(adj)
+	print(cd)

### 2024-05-21 code/carving_width.py

@@ -5,24 +5,10 @@ from parse_graph import adj_to_text, adj_to_text_2, parse_text_to_adj
 from dual_graph import dual_graph
 
 def carving_width(G: Graph) -> int:
-	# f = open(f"../graphs/g/8-5/quiet/_medial.in", "w")
-	# f.write(adj_to_text_2(G.adj()))
-	# f.close()
-
 	D, edge_to_link, link_to_edge, node_to_face, edge_to_node = dual_graph(G)
 
-	# blag = G.adj()
-	# for x,ys in D.adj().items():
-	# 	blag[x] = ys
-	# f = open(f"../graphs/g/8-5/quiet/_medial_dual.in", "w")
-	# f.write(adj_to_text_2(blag))
-	# f.close()
-
-	# print("node_to_face", node_to_face)
-	# print("node_to_face", [(k, list(map(lambda v: G.edge_to_vertexpair[v][0], vs))) for k,vs in node_to_face.items()])
-
 	# When the rat-catcher is on edge e, edge f is noisy iff there is
-	# a closed walk of at most length k containing e* and f* in G* .
+	# a closed walk of length scrictly less than k containing e* and f* in G* .
 	# Return the un-noisy subgraph.
 	def noisy_links(l: int, k: int) -> set[int]:
 		s,t = D.edge_to_vertexpair[l]
@@ -79,10 +65,6 @@ def carving_width(G: Graph) -> int:
 		for x,ys in D.adj().items():
 			blah[x] = ys
 
-		# f = open(f"../graphs/g/8-5/quiet/k{str(k)}e{G.edge_to_vertexpair[e]}.in", "w")
-		# f.write(adj_to_text_2(blah))
-		# f.close()
-
 		components = []
 		unseen = set(quiet_subgraph.keys())
 
@@ -116,11 +98,13 @@ def carving_width(G: Graph) -> int:
 		if max([len(G.N(v)) for v in G.V()]) >= k:
 			return True
 
+		# Set up the game states
 		edge_set = edge_to_link.keys()
 
 		Te = set([(e, tuple(C)) for e in edge_set for C in quiet_components(e, k)])
 		Sr = set([(r, v) for r in node_to_face.keys() for v in G.V()])
 
+		# Set up the losing states
 		losing_eC = set()
 		losing_rv = set()
 
@@ -131,6 +115,7 @@ def carving_width(G: Graph) -> int:
[…]

### 2024-05-21 code/compare_bw.py

@@ -3,7 +3,7 @@ import time
 import datetime
 import math
 
-ns = [10,14,18,22,26,30,34,38,42,46,50,54,58,62,66,70,74,78,82,86,90,94,98]
+ns = [6,8,10,14,18,22,26,30,34,38,42,46,50,54,58,62,66,70,74,78,82,86,90,94,98]
 
 for n in ns:
 	with open(f"../graphs/g/{n}-5.in", 'rb') as f:
@@ -11,8 +11,7 @@ for n in ns:
 		r = subprocess.run(['python3', 'branch_width.py'], stdin=f, text=True, capture_output=True)
 		res = r.stdout.strip()
 		seconds_elapsed = time.time() - start_time
-		print("n", n)
-		print("bound", 2.122*math.sqrt(n))
+		print("n", n, end=' ')
+		# print("bound", 2.122*math.sqrt(n))
 		print(res)
-		print("time", str(datetime.timedelta(seconds=seconds_elapsed)))
-		print()
+		# print("time", str(datetime.timedelta(seconds=seconds_elapsed)))

### 2024-05-21 code/contraction.py

@@ -1,4 +1,3 @@
-
 from Graph import Graph
 from parse_graph import adj_to_text, parse_text_to_adj
 
@@ -8,10 +7,14 @@ def index_of_first(lst, pred):
 			return i
 	return None
 
+# assume G might have parallel edges
+# assume G do not have self-loops
+# assume adjacency list of G has clockwise ordering of neighbors
 def contraction(G: Graph, a: int, b: int) -> Graph:
 	# copy G
 	G1 = G.copy()
 
+	# create new vertex c
 	c = max(G1.adj_edges.keys()) + 1
 
 	# let every edge incident to a or b be incident to c instead
@@ -34,7 +37,7 @@ def contraction(G: Graph, a: int, b: int) -> Graph:
 
 	G1.adj_edges[c] = rotated_Ga + rotated_Gb
 
-	# remove c <-> c edges
+	# remove self-loops on c
 	G1.adj_edges[c] = [e for e in G1.adj_edges[c] if not (G1.edge_to_vertexpair[e][0] == G1.edge_to_vertexpair[e][1] == c)]
 	G1.edge_to_vertexpair = dict([(k,v) for k,v in G1.edge_to_vertexpair.items() if not (v[0] == v[1] == c)])
 

### 2024-05-21 code/count_hamcyc_dp_thrice.py

@@ -3,7 +3,7 @@ from parse_graph import parse_graph_to_adj
 
 def find_faces(adj: dict[int, list[int]]) -> list[list[(int, int)]]:
 	# assume planar graph
-	# assume clockwise or counterclockwise ordering
+	# assume clockwise ordering of neighbors
 	edges = set([(i, j) for i in adj for j in adj[i]])
 	faces = []
 	while len(edges) > 0:

### 2024-05-21 code/count_hamcyc_dp_twice.py

@@ -3,7 +3,7 @@ from parse_graph import parse_graph_to_adj
 
 def find_faces(adj: dict[int, list[int]]) -> list[list[(int, int)]]:
 	# assume planar graph
-	# assume clockwise or counterclockwise ordering
+	# assume clockwise ordering
 	edges = set([(i, j) for i in adj for j in adj[i]])
 	faces = []
 	while len(edges) > 0:

### 2024-05-21 code/dual_graph.py

@@ -17,13 +17,11 @@ def dual_graph(G: Graph) -> Graph:
 		edge_to_node[e] = next_nodeid
 		face = [e]
 		while True:
-			# print("face", face, list(map(lambda x: G.edge_to_vertexpair[x], face)))
 			u,v = G.edge_to_vertexpair[next_e]
 			idx = G.adj_edges[v].index(-next_e)
 			next_e = G.adj_edges[v][(idx-1)%len(G.adj_edges[v])]
 			if (next_e == e):
 				break
-			# print("next_e", next_e)
 			edges.remove(next_e)
 			face.append(next_e)
 			edge_to_node[next_e] = next_nodeid

### 2024-05-21 code/medial_graph.py

@@ -1,16 +1,18 @@
 from Graph import Graph
 from parse_graph import adj_to_text, parse_text_to_adj
 
-def medial_graph(adj: dict[int, list[int]]) -> Graph:
-	es = set([tuple(sorted((i, j))) for i in adj for j in adj[i]])
+# assume planar graph
+# assume clockwise ordering of neighbors
+def medial_graph(G_adj: dict[int, list[int]]) -> Graph:
+	half_edges = set([tuple(sorted((i, j))) for i in G_adj for j in G_adj[i]])
 
-	vertexpair_to_node = dict([(e, i+1) for i,e in enumerate(es)])
-	node_to_vertexpair = dict([(i+1, e) for i,e in enumerate(es)])
+	vertexpair_to_node = dict([(e, i+1) for i,e in enumerate(half_edges)])
+	node_to_vertexpair = dict([(i+1, e) for i,e in enumerate(half_edges)])
 
-	medial = dict([(i+1, []) for i in range(len(es))])
+	medial = dict([(i+1, []) for i in range(len(half_edges))])
 
-	for v,xs in adj.items():
-		nodes = [vertexpair_to_node[tuple(sorted((v, x)))] for x in xs]
+	for u,vs in G_adj.items():
+		nodes = [vertexpair_to_node[tuple(sorted((u, v)))] for v in vs]
 		for i in range(len(nodes)):
 			medial[nodes[i]].append(nodes[(i-1)%len(nodes)])
 			medial[nodes[i]].append(nodes[(i+1)%len(nodes)])

### 2024-05-21 code/parse_graph.py

@@ -1,4 +1,3 @@
-import networkx as nx
 import sys
 
 def parse_bin_to_adj() -> dict[int, list[int]]:
@@ -30,6 +29,12 @@ def adj_to_text(adj):
 	print(len(adj))
 	for v,xs in adj.items():
 		print(v, *xs)
+
+def adj_to_text_2(adj):
+	s = str(len(adj)) + "\n"
+	for v,xs in adj.items():
+		s += str(v) + " " + " ".join(map(str, xs)) + "\n"
+	return s
 
 def adj_to_nx(adj):
 	G = nx.MultiDiGraph()

### 2024-05-21 code/permute_vertices.py

@@ -1,0 +1,15 @@
+from parse_graph import parse_text_to_adj, adj_to_text
+adj = parse_text_to_adj()
+
+vertices = list(adj.keys())
+rotated = vertices[1:] + vertices[:1]
+
+mapping = dict([(vertices[i], rotated[i]) for i in range(len(vertices))])
+
+P = dict([(v, []) for v in vertices])
+
+for i in adj:
+	for j in adj[i]:
+		P[mapping[i]].append(mapping[j])
+
+adj_to_text(P)

### 2024-05-21 paper/main.tex
Change focus to Implementation of an optimal branch-decomposition algorithm for planar graphs.

@@
 
-\title{Empirical study of an algorithm for finding Hamiltonian cycles in 3-regular planar graphs.}
+\title{Implementation of an optimal branch-decomposition algorithm for planar graphs.}
 \author{Kristoffer Højelse}
@@
 \begin{abstract}
-	An algorithm in the making considers the problem of finding Hamiltonian cycles in 3-regular planar graphs.
+	Seymour and Thomas give an algorithm, the rat-catching algorithm, for deciding $bw(G) \leq c$ in $O(n^2)$ time, and by using it as a subroutine, an algorithm to compute an optimal branch-decomposition in $O(n^4)$ time.
+In this paper, I describe an implementation of this algorithm and publish the source code.
 \end{abstract}
 
+\section{Motivation}
+	Graph optimization problems may be solved efficiently for graphs of small branchwidth.
+
 \section{Description of the problem}
+	A branch decomposition $B$ of a graph $G$ is a connected tree where every edge of $G$ is a leaf in $B$ and every internal vertex of $B$ has exactly 3 neighbors.
 
-	\textsc{The Hamiltonian Cycle Problem} is a computational problem that comes in a few variants, with slightly different input and output types.
+	Removing any edge $e = \{u,v\}$ of $B$ partitions $B$ into 2 trees $B_u$ and $B_v$ and the intersection of the sets of vertices in the leaves of $B_u$ and $B_v$ is called a middle set and is associated with the edge $e$.
+Every edge of $B$ therefore, has an associated middle set.
+The maximal cardinality of any middle set among all middle sets of $B$ is the width of the branch decomposition $B$.
 
-	\textsc{The Hamiltonian Cycle Decision problem} asks given a simple graph, to output; yes or no to the question; Does there exist a Hamiltonian Cycle in the graph?
-Variants might ask for an example/witness/certificate of one such cycle, to make verifying the result easy.
+	There can be many branch decompositions of a graph $G$.
 
-	\textsc{The Hamiltonian Cycle Counting problem} asks given a simple graph, to output; the number of unique Hamiltonian cycles. Solving this problem you get the decision variant for free.
+	A minimal branch decomposition of $G$ is any branch decomposition of $G$ of minimal width among all branch decompositions of $G$.
 
 	\begin{definition}
-		\textsc{The Hamiltonian Cycle Counting Problem}
+		\textsc{The Minimal Branch Decomposition Problem}
 
@@
 
-		Output: The number of unique Hamiltonian cycles.
-
-		A pair of Hamiltonian cycles are not unique if and only if you can obtain the other by circular permutations and inversions of the associated vertex sequences.
+		Output: A minimal branch decomposition of $G$.
 	\end{definition}
 
-	The simplest algorithm I can come up with that solves this problem has a factorial time complexity $\Theta(N!)$.
-
-	Enumerate all $N!$ permutations of the $N$ vertices and count how many of those vertex sequences are valid Hamiltonian cycles, by checking firstly that the foremost and final vertex of the sequence are adjacent and secondly that every consecutive pair of vertices in the sequence are adjacent.
-Then return this count not before dividing by a factor of $2N$, as there are $N$ circular permutations and $2$ inversions of a sequence of $N$ elements.
-
-	A Python program of this algorithm is in the appendix \ref{appx:count-hamcyc-brute-force}.
-
-	You cannot do much better\ref{} for this general problem. But if we constrain the input we can do much better.
-
 \section{The domain}
-	The algorithm described in this paper solves \textsc{The Hamiltonian Cycle Counting Problem} with a very contained subset of all graphs; simple undirected connected planar cubic graphs.
+	The algorithm described in this paper solves \textsc{The Minimal Branch Decomposition Problem} with an additional constraint of the input graph being planar.
 
@@
 		\item A graph is called \textbf{connected} (or 1-vertex-connected) if and only if there exists a path between any two vertices.
-		\item A graph is called \textbf{planar} if and only if there's a way to draw the graph in 2 dimensions such that no pair of edges crosses.
-		\item A graph is called \textbf{cubic} if and only if every vertex has exactly three neighbors.
+		\item A graph is called \textbf{planar} if and only if there is a way to draw the graph in 2 dimensions such that no pair of edges crosses.
 	\end{enumerate}
 
-	Most variants of \textsc{The Hamiltonian Cycle Problem} too require the input graph to be simple, undirected, and connected. These problems probably have no known polynomial-time solutions.
+	Different subroutines of the algorithm that take a graph as argument, assume different properties of the graph.
 
-	The algorithm exploits the two remaining attributes, planar and cubic, to obtain a polynomial running time.
+	\subsection{Medial graph}
 
-\section{Description of the algorithm}
+		A subroutine of the algorithm computes the medial graph of a graph from the class; simple connected planar cubic graphs.
 
-	Overview:
+		\begin{definition}
+			The medial graph $M(G)$ of a connected plane graph $G$ with a vertex $e^*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $c*$ between a pair of vertices $e_1*$, $e_2*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
+		\end{definition}
 
-	\begin{enumerate}
-		\item Count Hamiltonian Cycles by Dynamic Programming over a minimal Branch Decomposition.
-		\item Compute a minimal Branch Decomposition from a minimal Carving Decomposition of the Medial graph.
-		\item Compute a minimal Carving Decomposition by repeatedly Contracting edges that doesn't increase the Carving Width.
-		\item Compute the Carving Width with the Rat Catching algorithm which needs to compute a dual graph.
-		\item Compute a dual graph.
-	\end{enumerate}
+		todo: argue that simple connected planar cubic graphs are a valid substitution for a connected plane.
+
+		todo: identify a more succinct description of the class; medial of simple connected planar cubic graphs
+
+		An implementation of this subroutine can be found in the appendix \ref{appx:medial-graph}.
 
@@
 		\begin{definition}
-			The dual graph $G*$ of a planar graph $G$ is a graph with a vertex $f*$ for each face $f$ of $G$ and an edge $e*$ for each edge $e$ that separates a face $f_1$ of $G$ and a face $f_2$ of $G$.
+			The dual graph $G*$ of a planar graph $G$ is a graph with a vertex $f*$ for each face $f$ of $G$ and an edge $e^*$ for each edge $e$ that separates a face $f_1$ of $G$ and a face $f_2$ of $G$.
 		\end{definition}
@@
 		\begin{corollary}\label{def:dual_coro_loop}
-			If $e$ separates $f_1$ and $f_2$ and are the same face, $e*$ will be a self-loop.
+			If $e$ separates $f_1$ and $f_2$ and are the same face, $e^*$ will be a self-loop.
 		\end{corollary}
@@
 
-	\subsection{Medial graph}
+		An implementation of this subroutine can be found in the appendix \ref{appx:dual-graph}.
 
-		A subroutine of the algorithm computes the medial graph of a graph from the class; simple connected planar cubic graphs.
+	\subsection{Edge contraction}
+
+		A subroutine of the algorithm computes the resulting graph from an edge contraction.
+
+		This operation is a bit different from the conventional understanding of an edge contraction.
 
 		\begin{definition}
-			The medial graph $M(G)$ of a connected plane graph $G$ with a vertex $e*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $c*$ between a pair of vertices $e_1*$, $e_2*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
+			Edge contraction.
+
+			Given an undirected Graph $G=\{V,E\}$ with no self-loops and pair of vertices $u,v \in V$ such that $\{u,v\} \in E$, remove all edges between $u$ and $v$ and update every edge $\{v,w\} \in E$ to be $\{u,w\}$.
 		\end{definition}
 
-		todo: argue that simple connected planar cubic graphs are a valid substitution for a connected plane.
+		The resulting graph might have parallel edges but will not have self-loops.
 
-		todo: identify a more succinct description of the class; medial of simple connected planar cubic graphs
+\section{Description of the algorithm}
 
+	The algorithm computes a minimal branch decomposition of a simple connected planar graph $G$, the width of this branch decomposition is the branch width of $G$.
+This section describes a high-level set of problems for the algorithm to tackle.
+
+	\begin{problem}\label{desc:mbd}
+		Given a simple connected planar graph $G$, output a minimal branch decomposition of $G$.
+	\end{problem}
+
+	Problem \ref{desc:mbd} is the overarching problem, that the algorithm solves, and can be broken down into many smaller subproblems.
+
+	For the first large step in the algorithm, we use the fact that a minimal carving decomposition of the medial of a graph $G$ can be translated into a minimal branch decomposition of $G$, by replacing the vertices in leaves of the decompositions with edges, using the mapping between edges and vertices from computing the medial graph.
+
+	The argument for why this is true can be found in \ref{}.
+
+	\begin{problem}
+		Given a graph $G$ and a minimal carving decomposition of a medial graph of $G$, output a minimal branch decomposition of $G$.
+	\end{problem}
+
+	For both branch- and carving-decompositions, I have chosen a data structure of tuples of tuples or integers.
+This has a straightforward translation to the Newick tree format, a concise notation for tree structures.
+
+	To then solve the above-mentioned problem, the implementation recursively returns a copy of any tuple, but returns a tuple of integers for any integer, using the mapping from medial node to vertex pair.
+
+	\pyfile{branch_decomposition.py}
+
+	\begin{problem}
+		Given a graph $G$, output a medial graph and a mapping from medial nodes to vertex pairs.
+	\end{problem}
+
+	The medial graph $M$ of $G$ is a graph where there is a node in $M$ for each edge in $G$, and an edge between two nodes if their edges are consecutive in a face of $G$.
+
+	Therefore if you assume that the adjacency list of $G$ is given such that the neighborhoods are given in clockwise ordering according to some plane embedding of $G$ then a pair of consecutive nodes in a neighborhood are also consecutive edges in some face of $G$.
+
+	\pyfile{medial_graph.py}
+
+	For the upcoming problems one needs to deal with parallel edges and be able to tell them apart, therefore the implementation uses a data structure that encapsulates an adjacency list of edges and a map from unique edge ids its vertexpair.
+
+	I have chosen to assign IDs such that if one halfedge has ID $i$ then the other halfedge has ID $-i$, therefore the absolute value $|i|$ uniquely identifies an undirected edge.
+
+	\pyfile{graph.py}
+
+	Doing a series of edge contractions (contraction of all edges between a pair of vertices) on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices can be reassembled into a minimal carving decomposition of $M$.
+
+	The argument for why this is true can be found in \ref{}.
+
+	\begin{problem}
+		Given a graph $G$ that might have parallel edges, output a graph resulting from a contraction of all edges between a pair of vertices, preserving clockwise ordering.
+	\end{problem}
+
+	As the resulting graph is later given as an argument to functions assuming a clockwise ordering of vertices, the implementation needs to preserve this invariant when contracting.
+
+	As the contraction is a contraction of ALL edges between a pair of vertices, the resulting graph will not exhibit any self-loops.
+I suspect reconciling this and the ordering invariant could be difficult, but luckily in this context, it is irrelevant.
+
+	For a contraction of vertices $a$ and $b$, I have chosen to create a new vertex ID $c$ instead of reusing $a$ or $b$ as this later makes assembling the carving decomposition easier.
+
+	Creating the neighborhood of $c$ from the neighborhoods of $a$ and $b$ is done by finding the first edge that the adjacency lists of $a$ and $b$ have in common, and then "rotating" the adjacency lists such that a concatenation of the lists preserve the ordering.
+This is where telling apart parallel edges is very useful.
+
+	\pyfile{contraction.py}
+
+	I will defer describing how to compute the carving width and focus on the following problem for now.
+
+	\begin{problem}
+		Given a graph $G$ and function to compute the carving width of a graph, output a minimal carving decomposition of $G$.
+	\end{problem}
+
+	The implementation finds a nonincreasing contraction by doing a linear search over every edge.
+No consideration has yet been given to any potential clever orderings of the edges that might improve the running time.
+
+	The sequence of contracted edges is found and reassembled into a minimal carving decomposition.
+
+	\pyfile{carving_decomposition.py}
+
+	\begin{problem}
+		Given a graph $M$ that might have parallel edges, output the carving width of $M$.
+	\end{problem}
+
+	The rat-catching algorithm decides $cw(M) \geq k$ with $k$ being in the positive integers.
+This is a monotonic boolean space, so you can perform a binary search to find the smallest $k$ where $cw(M) \geq k$ is true.
+
+	The proposition $cw(M) \geq k$ is true if and only if $\Delta(M) \geq k$ or the rat can evade the rat-catcher indefinitely, with noise-level $k$, in a particular game based on $M$.
+
+	\pyfile{carving_width.py}
+
+	\begin{problem}
+		Given a graph $G$ that might have parallel edges, output the dual of $G$.
+	\end{problem}
 \section{Appendix.}
@@
 
+	\subsection{Medial graph}
+		\label{appx:medial-graph}
+		\pyfile{medial_graph.py}
+
+	\subsection{Dual graph}
+		\label{appx:dual-graph}
+		\pyfile{dual_graph.py}
+
 \end{document}

### 2024-05-21 paper/setup.tex
Add paper
Change focus to Implementation of an optimal branch-decomposition algorithm for planar graphs.

@@
 	\newcommand\bigO{\mathrm{O}}
+	\usepackage[T1]{fontenc}
 
 % Define the theorem-like environments
+	% Define a new theorem style with upright text
+	\newtheoremstyle{plainupright} % Name of the style
+	{10pt}                        % Space above
+	{3pt}                        % Space below
+	{}                           % Body font (default is \itshape which is italics)
+	{}                           % Indent amount
+	{\bfseries}                  % Theorem head font
+	{.}                          % Punctuation after theorem head
+	{ }                          % Space after theorem head
+	{}                           % Theorem head spec (can be left empty, meaning 'normal')
+
+	\theoremstyle{plainupright}
+
 	\newtheorem{theorem}{Theorem}[section]  % Numbering within sections
@@
 	\newtheorem{conjecture}[theorem]{Conjecture}  % Numbering with theorems
+	\newtheorem{problem}[theorem]{Problem}  % Numbering with theorems
+
+% Enclose theorem-like environments within boxes
+	\surroundwithmdframed[style=theoremstyle]{theorem}
+	\surroundwithmdframed[style=theoremstyle]{definition}
+	\surroundwithmdframed[style=theoremstyle]{corollary}
+	\surroundwithmdframed[style=theoremstyle]{claim}
+	\surroundwithmdframed[style=theoremstyle]{lemma}
+	\surroundwithmdframed[style=theoremstyle]{axiom}
+	\surroundwithmdframed[style=theoremstyle]{conjecture}
+	\surroundwithmdframed[style=theoremstyle]{problem}
 
@@
 	}
-% Enclose theorem-like environments within boxes
-	\surroundwithmdframed[style=theoremstyle]{theorem}
-	\surroundwithmdframed[style=theoremstyle]{definition}
-	\surroundwithmdframed[style=theoremstyle]{corollary}
-	\surroundwithmdframed[style=theoremstyle]{claim}
 
@@
 	\usepackage{xcolor}
-	\definecolor{mintedBg}{rgb}{0.97,0.97,0.98}
+	\definecolor{mintedBg}{rgb}{0.98,0.98,0.98}
 	\setminted[py]{
@@
 		breaklines,
+		breakanywhere,
 	}
 	\newcommand{\pyfile}[1]{%
-		\begin{filecontents*}{#1}
-		\end{filecontents*}
-		\noindent\textbf{#1}\par
-		\inputminted{py}{#1}%
+		\textsf{%
+			\detokenize{#1}%
+		}%
+		\par
+		\inputminted{py}{../code/\detokenize{#1}}%
 	}

### 2024-05-27 paper/main.tex
Refactor code and update paper

--- 2024-04-26 paper/main.tex
@@
 
-\title{Empirical study of an algorithm for finding Hamiltonian cycles in 3-regular planar graphs.}
+\title{Implementation of a minimal branch-decomposition algorithm for simple planar graphs.}
 \author{Kristoffer Højelse}
@@
 \begin{abstract}
-	An algorithm in the making considers the problem of finding Hamiltonian cycles in 3-regular planar graphs.
+	Seymour and Thomas give an algorithm, the rat-catching algorithm, for deciding $bw(G) \leq c$ in $O(n^2)$ time, and by using it as a subroutine, an algorithm to compute an optimal branch-decomposition in $O(n^4)$ time.
+In this paper, I describe an implementation of this algorithm and publish the source code.
 \end{abstract}
+
+\section{Motivation}
+	Some graph optimization problems can be solved efficiently for graphs of small branchwidth.\ref{}
+
+	\todo{which? counting hamiltonian cycles of planar cubic graphs}
 
@@
 
-	\textsc{The Hamiltonian Cycle Problem} is a computational problem that comes in a few variants, with slightly different input and output types.
-
-	\textsc{The Hamiltonian Cycle Decision problem} asks given a simple graph, to output; yes or no to the question; Does there exist a Hamiltonian Cycle in the graph?
-Variants might ask for an example/witness/certificate of one such cycle, to make verifying the result easy.
-
-	\textsc{The Hamiltonian Cycle Counting problem} asks given a simple graph, to output; the number of unique Hamiltonian cycles. Solving this problem you get the decision variant for free.
+	The main computational problem of this paper is \textsc{The Planar Minimal Branch Decomposition Problem}.
 
 	\begin{definition}
-		\textsc{The Hamiltonian Cycle Counting Problem}
+		\textsc{The Planar Minimal Branch Decomposition Problem}
+
+		Input: Given a simple undirected connected planar graph $G$.
+
+		Output: A minimal branch decomposition of $G$.
+	\end{definition}
+
+	Here is some informal definitions to unpack the aforementioned properties of graphs.
+
+	A graph is called \textbf{simple} if and only if it has no parallel edges and no self-loops.
+
+	A graph is called \textbf{undirected} if and only if all its edges can be traversed in both directions.
+
+	A graph is called \textbf{connected} (or 1-vertex-connected) if and only if there exists a path between any two vertices.
+
+	A graph is called \textbf{planar} if and only if there is a way to draw the graph in 2 dimensions such that no pair of edges crosses.
+
+	A \textbf{branch decomposition} $B$ of a graph $G$ is a tree where every edge of $G$ is a leaf in $B$ and every internal vertex of $B$ has exactly 3 neighbors. $B$ is an unrooted binary tree.
+
+	Removing any edge $e = \{u,v\}$ of $B$ partitions $B$ into 2 trees $B_u$ and $B_v$ and the intersection of the sets of vertices in the leaves of $B_u$ and $B_v$ is called a middle set.
+Every edge of $B$ has an associated middle set.
+The maximal cardinality of any middle set among all middle sets of $B$ is the width of $B$.
+
+	There can be many branch decompositions of a graph $G$.
+
+	A \textbf{minimal branch decomposition} of $G$ is any branch decomposition of $G$ of minimal width among all branch decompositions of $G$.
+
+	Without the contraint of planarity on the input graph, then it is NP-complete to determine whether a general graph $G$ has a branch decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.
+
+	\begin{definition}\label{def:mbdp}
+		\textsc{The Minimal Branch Decomposition Problem}
 
@@
 
-		Output: The number of unique Hamiltonian cycles.
-
-		A pair of Hamiltonian cycles are not unique if and only if you can obtain the other by circular permutations and inversions of the associated vertex sequences.
+		Output: A minimal branch decomposition of $G$.
 	\end{definition}
 
-	The simplest algorithm I can come up with that solves this problem has a factorial time complexity $\Theta(N!)$.
+	The algorithm described in this paper solves \textsc{The Planar Minimal Branch Decomposition Problem} which can be computed in polynomial time.
 
-	Enumerate all $N!$ permutations of the $N$ vertices and count how many of those vertex sequences are valid Hamiltonian cycles, by checking firstly that the foremost and final vertex of the sequence are adjacent and secondly that every consecutive pair of vertices in the sequence are adjacent.
-Then return this count not before dividing by a factor of $2N$, as there are $N$ circular permutations and $2$ inversions of a sequence of $N$ elements.
+	The width of a minimal branch decomposition of $G$ called the branch width of $G$.
 
-	A Python program of this algorithm is in the appendix \ref{appx:brute-force-count-ham-cyc}.
+\section{The algorithm}
+	This section describes the algorithm given by Seymour and Thomas\ref{} by indentifying a set of practical problems and subproblems and how they relate.
 
-	You cannot do much better\ref{} for this general problem. But if we constrain the input we can do much better.
+	Problem \ref{def:mbdp} is the overarching problem, that the algorithm solves, and can be broken down into many smaller subproblems.
 
-\section{The domain}
-	The algorithm described in this paper solves \textsc{The Hamiltonian Cycle Counting Problem} with a very contained subset of all graphs; simple undirected connected planar cubic graphs.
+	Considering a graph $G$, you can compute a minimal branch decomposition of $G$ from a minimal carving decomposition of the medial of $G$, by replacing the vertices in leaves of the decomposition with edges, using the mapping between edges and vertices from computing the medial graph.
+\ref{} Therefore problem \ref{def:mbdp} break down into problems \ref{problem:mcd-to-mbd}, \ref{problem:medial} and \ref{problem:mcd}.
 
-	Some informal definitions of the aforementioned properties of graphs:
+	\begin{problem}\label{problem:mcd-to-mbd}
+		Given a minimal carving decomposition of a medial graph of $G$, output a minimal branch decomposition of $G$.
+	\end{problem}
 
-	\begin{enumerate}
-		\item A graph is called \textbf{simple} if and only if it has no parallel edges and no self-loops.
-		\item A graph is called \textbf{undirected} if and only if all its edges can be traversed in both directions.
-		\item A graph is called \textbf{connected} (or 1-vertex-connected) if and only if there exists a path between any two vertices.
-		\item A graph is called \textbf{planar} if and only if there's a way to draw the graph in 2 dimensions such that no pair of edges crosses.
-		\item A graph is called \textbf{cubic} if and only if every vertex has exactly three neighbors.
-	\end{enumerate}
+	\begin{problem}\label{problem:medial}
+		Given a graph $G$, output a medial graph and a bijectional mapping between medial nodes and vertex pairs.
+	\end{problem}
 
-	Most variants of \textsc{The Hamiltonian Cycle Problem} too require the input graph to be simple, undirected, and connected. These problems probably have no known polynomial-time solutions.
+	\begin{problem}\label{problem:mcd}
+		Given a plane graph $M$ and function to compute the carving width of a graph, output a minimal carving decomposition of $M$.
+	\end{problem}
 
-	The algorithm exploits the two remaining attributes, planar and cubic, to obtain a polynomial running time.
+	Implementing a function to solve \ref{problem:mcd-to-mbd} is decribed in \ref{impl:mcd-to-mbd}.
 
-\section{Description of the algorithm}
+	To solve \ref{problem:medial} is a matter of following the definition.
 
-	Overview:
+	I will refer to vertices and edges of the medial graph as "nodes" and "links" in an attempt of disambiguation.
 
-	\begin{enumerate}
-		\item Count Hamiltonian Cycles by Dynamic Programming over a minimal Branch Decomposition.
-		\item Compute a minimal Branch Decomposition from a minimal Carving Decomposition of the Medial graph.
-		\item Compute a minimal Carving Decomposition by repeatedly Contracting edges that doesn't increase the Carving Width.
-		\item Compute the Carving Width with the Rat Catching algorithm which needs to compute a dual graph.
-		\item Compute a dual graph.
-	\end{enumerate}
-
-	\subsection{Dual graph}
-
-		A subroutine of the algorithm computes the dual graph of a graph.
+	Informally; The medial graph is a graph where there is a node in the medial for each edge, and an edge between two nodes if their corresponding edges are consecutive in a face of the graph.
 
 	\begin{definition}
-			The dual graph $G*$ of a planar graph $G$ is a graph with a vertex $f*$ for each face $f$ of $G$ and an edge $e*$ for each edge $e$ that separates a face $f_1$ of $G$ and a face $f_2$ of $G$.
+		The medial graph $M(G)$ of a connected plane graph $G$ is a graph with a vertex $e^*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $d^*$ between a pair of vertices $e_1^*$, $e_2^*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
+	\end{definition}
+
+	Computing a medial graph is described in \ref{impl:medial}.
+
+	To solve \ref{problem:mcd} \ref{} gives an contraction algorithm.
+
+	Doing a series of edge contractions (contraction of all edges between a pair of vertices) on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices can be assembled into a minimal carving decomposition of $M$.
+
+	I will defer describing exactly how to assemble a minimal carving decomposition to \ref{impl:mcd}.
+
+	The contraction algorithm depends on a function to compute a contraction and a function to compute the carving width. This is problems \ref{problem:contraction} and \ref{problem:cw}.
+
+	\begin{problem}\label{problem:contraction}
+		Given a graph $M$ that might have parallel edges, output a graph resulting from a contraction of all edges between a pair of vertices.
+	\end{problem}
+
+	\begin{problem}\label{problem:cw}
+		Given a plane graph $M$ that might have parallel edges, output the carving width of $M$.
+	\end{problem}
+
+	First consider problem \ref{problem:contraction}.
+
+	Informally; a contraction is merging two vertices into one and letting incident edges connect to the new vertex.
+The contraction in this paper differs a bit from conventional definitions of an edge- and vertex-contractions, as other definitions might result in self-loops when contracting one of multiple parallel edges.
+
+	\begin{definition}
+		Contraction.
+
+		Given an undirected Graph $G=\{V,E\}$ with no self-loops and pair of vertices $u,v \in V$ such that $\{u,v\} \in E$, remove all edges between $u$ and $v$ and update every edge $\{v,w\} \in E$ to be $\{u,w\}$.
+	\end{definition}
+
+	Computing a medial graph is described in \ref{impl:medial}.
+
+	Now consider problem \ref{problem:cw}.
+
+	The rat-catching algorithm decides $cw(M) \geq k$ with $k$ being in the positive integers.
+This is a monotonic boolean space, so you can perform a binary search to find the smallest $k$ where $cw(M) \geq k$ is true.
+
+	The rat-catching algorithm can be described as a game of two players, the rat and rat-catcher.
+Considering a graph $M$, the edges of a face can be thought of as walls of a room and vertices as the corner of some rooms.
+The rat moves from corner to corner along the walls and the rat-catcher moves from room to room through some wall.
+The rat-catcher can force the rat away from some walls by making noise.
+A round of this game is played with some noise-level $k$.
+The rat-catcher wins if they can force the rat to be in some wall of the room that they are in, with noise-level $k$, and the rat wins if there is a strategy whereby the rat can escape indefinitely.
+
+	Additionally, if $\Delta(M) \geq k$ then the rat wins, the argument for why is this is true is glossed over in \ref{BGZ15}. This is discussed in section \ref{}.
+
+	So assuming $\Delta(M) < k$ the game is played.
+
+	For some noise-level and location of the rat-catcher, exactly which edges are noisy and which are quiet, is definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
+
+	An edge $e$ is called quiet iff. $e$ is not noisy.
+
+	\begin{definition}\label{def:edge-quiet}
+		When the rat-catcher is on some edge $e_1$, then edge $e_2$ is noisy iff. there is a closed walk of length scrictly less than $k$ containing $e_1^*$ and $e_2^*$ in the dual $M^*$.
+	\end{definition}
+
+	\begin{definition}\label{def:face-quiet}
+		When the rat-catcher is on some room $f$, then edge $e$ is noisy iff. there is a closed walk of length scrictly less than $k$ containing $f^*$ and $e^*$ in the dual $M^*$.
+	\end{definition}
+
+	A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise-level $k$ and some $e \in E(M)$, is a subgraph of $M$ with $V(Q(M, k, e)) = V(M)$ and
+
+	$E(Q(M, k, e)) = \{e_1 \mid$ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
+
+	\begin{problem}\label{problem:noisy-subgraph}
+		Given a plane graph $M$ that might have parallel edges, an edge $e \in E(M)$, and noise-level $k \in \mathbb{N}$, output the quiet subgraph $Q(M, k, e)$.
+	\end{problem}
+
+	Problem \ref{problem:noisy-subgraph} depends on a function for computing the dual of a graph. Computing a dual graph is problem \ref{problem:dual}.
+
+	\begin{problem}\label{problem:dual}
+		Given a plane graph $M = \{V, E\}$ that might have parallel edges, output the dual of $M$.
+	\end{problem}
+
+	\begin{definition}
+		The dual graph $M^*$ of a plane graph $M$ is a graph where a face $f$ of $M$ is a vertex $f^*$ of $M^*$ and an edge $e$ that separates two faces $f_1$,$f_2$ of $M$ is an edge $e^*$ of $M^*$ between $f_1^*$ and $f_2^*$.
 	\end{definition}
@@
 	\begin{corollary}
-			If multiple edges separate $f_1$ and $f_2$ there will be multiple edges between $f_1*$ and $f_2*$.
+		If multiple edges separate $f_1$ and $f_2$ there will be parallel edges between $f_1*$ and $f_2*$.
 	\end{corollary}
@@
 	\begin{corollary}\label{def:dual_coro_loop}
-			If $e$ separates $f_1$ and $f_2$ and are the same face, $e*$ will be a self-loop.
+		If the same face is on both sides of $e$, then $e^*$ will be a self-loop.
 	\end{corollary}
 
-		For the algorithm in this paper, the class of graphs that will be given as input is medial graphs of simple undirected connected planar cubic graphs.
+	For the algorithm in this paper, the class of graphs that will be given as input is medial graphs of simple undirected connected planar graphs. Therefore I claim \ref{claim:irri}.
 
-		I therefore claim, for now without any proof or argument, that;
-
-		\begin{claim}
+	\begin{claim}\label{claim:irri}
 		Corollary \ref{def:dual_coro_loop} will be irrelevant for any implementation of the algorithm.
@@
 
-	\subsection{Medial graph}
+	The game states and possible moves, for some graph $M$ and some noise-level $k$, can be described as a graph $H(M, k)$.
 
-		A subroutine of the algorithm computes the medial graph of a graph from the class; simple connected planar cubic graphs.
+	Let $F(M)$ be the set of faces of $M$.
 
-		\begin{definition}
-			The medial graph $M(G)$ of a connected plane graph $G$ with a vertex $e*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $c*$ between a pair of vertices $e_1*$, $e_2*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
-		\end{definition}
+	Let $S$ be every possible state when the rat-catcher is in a face some of which might be losing states. $S = \{ (f, v) \mid v \in V(M) \land f$ is a face of $ M \}$.
 
-		todo: argue that simple connected planar cubic graphs are a valid substitution for a connected plane.
+	Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \mid e \in E(M) \land C$ is a connected component of $Q(M, k, e) \}$.
 
-		todo: identify a more succinct description of the class; medial of simple connected planar cubic graphs
+	Computing the quiet subgraph requires the dual graph.
+
+	With the graph $H$, the only missing peace of the rat-catching algorithm is how to determine the outcome.
+
+	You can mark states of the graph $H$ that are losing states, and then repeatedly mark any state that leads to a losing state, until either every state is marked or no more states can be marked.
+If every state is marked then the rat-catcher wins, otherwise the rat wins.
+
+\section{The implementation}
+
+	For the upcoming problems one needs to deal with parallel edges and be able to tell them apart, therefore the implementation uses a data structure that encapsulates an adjacency list of edges and a map from unique edge ids its vertexpair.
+
+	I have chosen to assign IDs such that if one halfedge has ID $i$ then the other halfedge has ID $-i$, therefore the absolute value $|i|$ uniquely identifies an undirected edge.
+
+	\pyfile{graph.py}{0}{100}
+
+	\subsection{Computing a minimal branch decomposition}\label{impl:mcd-to-mbd}
+
+		Solving problem \ref{problem:mcd-to-mbd}.
+
+		For both branch- and carving-decompositions, I have chosen a data structure of tuples of tuples or integers.
+This has a straightforward translation to the Newick tree format, a concise notation for tree structures.
+
+		To then solve the above-mentioned problem, the implementation recursively returns a copy of any tuple, but returns a tuple of integers for any integer, using the mapping from medial node to vertex pair.
+
+		\pyfile{branch_decomposition.py}{5}{19}
+
+	\subsection{Computing a medial graph}\label{impl:medial}
+
+		Solving problem \ref{problem:medial}.
+
+		I assume that the input graph $G$ is a planar graph given as an adjacency list such that the neighborhoods are given in clockwise ordering according to some plane embedding of $G$.
+
+		Given this format, any two consecutive edges $e_1$ and $e_2$ in some face of $G$ are therefore consecutive vertices in the neighborhood of the vertex that $e_1$ and $e_2$ share.
+
+		The implementation adds all medial links around some vertex for each vertex in $G$.
+
+		The clockwise ordering of neighborhoods of $G$ becomes counterclockwise orderings of neighborhoods of the medial $M$.
+The medial graph of a plane graph is 4-regular \ref{}.
+From the perpective of some medial node $v$, in some single iteration of the loop on line 14, two links are added to the neighborhood of $v$ in counterclockwise ordering, and latere the two links added to the neighborhood of $v$ also in counterclockwise ordering.
+
+		\pyfile{medial_graph.py}{4}{22}
+
+	\subsection{Computing a minimal carving decomposition}\label{impl:mcd}
+
+		Solving problem \ref{problem:mcd}.
+
+		The implementation finds a nonincreasing contraction by doing a linear search over every edge.
+No consideration has yet been given to any potential clever orderings of the edges that might improve the running time.
+
+		The sequence of contracted edges is found and reassembled into a minimal carving decomposition.
+
+		The "contraction" function returns a new unique vertex ID, therefore by saving which vertex is a contraction of which vertex pair in the "edges" dictionary, contructing the decomposition is then a matter of recursively expanding any vertices that was a result of a contraction into a tuple of the vertex pair that is was composed of.
+Repeating this until all only vertices of $M$ remain gives a carving decomposition in newick-like nested tuple format.
+
+		\pyfile{carving_decomposition.py}{8}{48}
+
+	\subsection{Contraction}\label{impl:contraction}
+
+		Solving problem \ref{problem:contraction}.
+
+		As the resulting graph is later given as an argument to functions assuming a clockwise or counterclockwise ordering of vertices, the implementation needs to preserve this invariant when contracting.
+
+		As this contraction is a contraction of ALL edges between a pair of vertices, the resulting graph will not exhibit any self-loops.
+I suspect reconciling this and the ordering invariant could be difficult, but luckily in this context, it is irrelevant.
+
+		For a contraction of vertices $a$ and $b$, I have chosen to create a new vertex ID $c$ instead of reusing $a$ or $b$ as this later makes assembling the carving decomposition easier.
+
+		First update any edges incident to either $a$ or $b$.
+Then creating the neighboorhood of the new vertex $c$ from the contraction of vertices $a$ and $b$, is done by firstly finding any shared edge $e$.
+In this implementation the first shared edge $e$ in the neighborhood of $a$.
+This edge has some ID $e$ and the other halfedge with ID $-e$ will therefore be in the neighborhood of $b$.
+Now  "rotating" the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0 in both lists means that a concatenation of the lists will preserve the ordering around the new vertex $c$.
+And finally remove any edges between $a$ and $b$.
+
+		This is where telling apart parallel edges, which the Graph class allows, becomes very useful.
+Infering where to stitch together the neighborhoods to preserve the ordering, just from a normal adjacency list, becomes a way harder problem.
+
+		\pyfile{contraction.py}{4}{49}
+
+	\subsection{Carving width and the rat cathing algorithm}\label{impl:cw}
+
+		Solving problem \ref{problem:cw}
+
+		The vertices of game state graph $H$ is initialized by computing the elements of $T$ and $S$, while edges of $H$ are not explicitly kept in any datastructure, but instead checked while playing the game.
+
+		Losing states - the tuples $(f,v) \in S$ where $v \in f$ - are marked as losing.
+
+		The outcome of the game is computed by marking states as losing.
+
+		Considering a tuple $(e, C) \in T$, if all $(f, v)$ where $v \in V(C)$ is losing then $(e, C)$ is losing.
+
+		Considering a tuple $(f, v)$, if there exists a tuple $(e, C)$ which is losing where $e \in f$ and $v \in V(C)$ then $(f, v)$ is losing.
+
+		\pyfile{carving_width.py}{7}{9}
+		\pyfile{carving_width.py}{78}{130}
+
+	\subsection{Quiet subgraph}\label{impl:quiet}
+
+		Solving problem \ref{problem:noisy-subgraph}.
+
+		Using definition \ref{def:edge-quiet}: When the rat-catcher is on some edge $e_1$, then edge $e_2$ is noisy iff.
+there is a closed walk of length scrictly less than $k$ containing $e_1^*$ and $e_2^*$ in the dual $M^*$.
+
+		Let $s_1$ and $t_1$ be the vertexpair for the link $e_1^*$ and let $s_2$ and $t_2$ be the vertexpair for the link $e_2^*$.
+
+		\begin{claim}
+			The shortest closed walk that includes both $e_1^*$ and $e_2^*$ has the same length as either $$d(s_1, s_2) + d(t_1, t_2) + 2$$ or $$d(s_1, t_1) + d(s_2, t_2) + 2$$.
+Where $d(u, v)$ is the length of the shortest path from $u$ to $v$.
+		\end{claim}
+
+		The single source shortest distances can then be computed using a breadth-first approach.
+
+		Using the mapping from links of the dual to edges, and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained in the natural way.
+
+		Computing the quiet subgraph and the connected components thereof is done with a depth-first search approach.
+
+		The edges of the connected components are irrelevant for the rest of the algorithm, so only a list of vertices are returned for each connected component.
+
+		\pyfile{carving_width.py}{10}{76}
+
+	\subsection{Dual graph}\label{impl:dual}
+
+		Solving problem \ref{problem:dual}
+
+		No other path of the implementation need the assumption that the dual is planar, therefore no clockwise or counterclockwise ordering of the neighborhoods of the adjacency list is needed.
+
+		The dual has a vertex for each face of the input graph.
+The faces are found by selecting an unmarked halfedge, and then marking all the edges of the face it belongs to, by following the edges which are just next to eachother in the ordered neighborhoods.
+
+		The next halfedge $e_{i+1}$ after the current halfedge $e_i = \{u, v\}$ is the edge just before $-e_i$ in the ordered neighborhood around $v$.
+
+		\pyfile{dual_graph.py}{21}{22}
+
+		\pyfile{dual_graph.py}{4}{50}
 
 \section{Appendix.}
-
-\subsection{Brute force count Hamiltonian Cycles}\label{appx:brute-force-count-ham-cyc}
+	\pyfile{branch_decomposition.py}{0}{1000}
+	\pyfile{branch_width_brute_force.py}{0}{1000}	
+	\pyfile{branch_width.py}{0}{1000}
+	\pyfile{carving_decomposition.py}{0}{1000}
+	\pyfile{carving_width_brute_force.py}{0}{1000}
+	\pyfile{carving_width.py}{0}{1000}
+	\pyfile{contraction.py}{0}{1000}
+	\pyfile{dual_graph.py}{0}{1000}
+	\pyfile{Graph.py}{0}{1000}
+	\pyfile{medial_graph.py}{0}{1000}
+	\pyfile{parse_graph.py}{0}{1000}
+	\pyfile{parse_newick.py}{0}{1000}
 

### 2024-05-27 code/bw_from_bd.py

@@ -1,4 +1,4 @@
-from code.branch_width.branch_width import branch_width_of_branch_decomposition
+from branch_width import branch_width_of_branch_decomposition
 from parse_newick import parse_newick
 
 s = input()

### 2024-05-27 code/carving_decomposition.py

@@ -6,7 +6,7 @@ from medial_graph import medial_graph
 from Graph import Graph
 
 # Find a contraction that does not increase the carving width
-def nonincreasing_cw_contraction(G: Graph, cw1):
+def nonincreasing_cw_contraction(G: Graph, cw1: int) -> tuple:
 	for es in G.E():
 		u, v = G.edge_to_vertexpair[es]
 		G2, w = contraction(G, u, v)
@@ -23,11 +23,11 @@ def gradient_descent_contractions(G: Graph) -> Graph:
 	cw1 = carving_width(G)
 	edges = dict()
 	while True:
-		G3, e, cw2, w = nonincreasing_cw_contraction(G2, cw1)
+		G3, uv, cw2, w = nonincreasing_cw_contraction(G2, cw1)
 		if G3 is not None and len(G3.V()) >= 3:
 			G2 = G3
 			cw1 = cw2
-			edges[w] = e
+			edges[w] = uv
 		if len(G2.V()) == 3:
 			return G2, edges
 

### 2024-05-27 code/carving_width.py

@@ -7,9 +7,9 @@ from dual_graph import dual_graph
 def carving_width(G: Graph) -> int:
 	D, edge_to_link, link_to_edge, node_to_face, edge_to_node = dual_graph(G)
 
-	# When the rat-catcher is on edge e, edge f is noisy iff there is
-	# a closed walk of length scrictly less than k containing e* and f* in G* .
-	# Return the un-noisy subgraph.
+	# If the rat-catcher is on edge e1, then edge e2 is noisy iff there is
+	# a closed walk of length scrictly less than k containing e1* and e2* in the dual G*.
+
 	def noisy_links(l: int, k: int) -> set[int]:
 		s,t = D.edge_to_vertexpair[l]
 		links = link_to_edge.keys()
@@ -56,15 +56,6 @@ def carving_width(G: Graph) -> int:
 			quiet_subgraph[u].append(e1)
 			quiet_subgraph[v].append(-e1)
 
-		blah = {v: [] for v in G.V()}
-		for e1 in edges:
-			u,v = G.edge_to_vertexpair[e1]
-			blah[u].append(v)
-			blah[v].append(u)
-
-		for x,ys in D.adj().items():
-			blah[x] = ys
-
 		components = []
 		unseen = set(quiet_subgraph.keys())
 
@@ -99,41 +90,41 @@ def carving_width(G: Graph) -> int:
 			return True
 
 		# Set up the game states
-		edge_set = edge_to_link.keys()
+		halfedges = edge_to_link.keys()
 
-		Te = set([(e, tuple(C)) for e in edge_set for C in quiet_components(e, k)])
-		Sr = set([(r, v) for r in node_to_face.keys() for v in G.V()])
+		T = set([(e, tuple(C)) for e in halfedges for C in quiet_components(e, k)])
+		S = set([(f, v) for f in node_to_face.keys() for v in G.V()])
 
 		# Set up the losing states
-		losing_eC = set()
-		losing_rv = set()
+		losing_T = set()
+		losing_S = set()
 
-		for (r, v) in Sr:
-			if v in flatten([G.edge_to_vertexpair[e] for e in node_to_face[r]]):
-				losing_rv.add((r,v))
+		for (f, v) in S:
+			if v in flatten([G.edge_to_vertexpair[e] for e in node_to_face[f]]):
+				losing_S.add((f,v))
 
-		if len(Te) == len(losing_eC) or len(Sr) == len(losing_rv):
[…]

### 2024-05-27 code/contraction.py

@@ -1,12 +1,6 @@
 from Graph import Graph
 from parse_graph import adj_to_text, parse_text_to_adj
 
-def index_of_first(lst, pred):
-	for i, v in enumerate(lst):
-		if pred(v):
-			return i
-	return None
-
 # assume G might have parallel edges
 # assume G do not have self-loops
 # assume adjacency list of G has clockwise ordering of neighbors
@@ -27,7 +21,14 @@ def contraction(G: Graph, a: int, b: int) -> Graph:
 			G1.edge_to_vertexpair[e] = (u, c)
 
 	# create neighborhood of c
-	first_shared_edge = G1.adj_edges[a][index_of_first(G1.adj_edges[a], lambda e: G1.edge_to_vertexpair[e][0] == c and G1.edge_to_vertexpair[e][1] == c)]
+	def index_of_first(lst, pred):
+		for i, v in enumerate(lst):
+			if pred(v):
+				return i
+		return None
+
+	index_of_first_shared_edge = index_of_first(G1.adj_edges[a], lambda e: G1.edge_to_vertexpair[e][0] == c and G1.edge_to_vertexpair[e][1] == c)
+	first_shared_edge = G1.adj_edges[a][index_of_first_shared_edge]
 
 	idx1 = G1.adj_edges[a].index(first_shared_edge)
 	rotated_Ga = G1.adj_edges[a][idx1:] + G1.adj_edges[a][:idx1]

### 2024-05-27 code/medial_graph.py

@@ -4,12 +4,12 @@ from parse_graph import adj_to_text, parse_text_to_adj
 # assume planar graph
 # assume clockwise ordering of neighbors
 def medial_graph(G_adj: dict[int, list[int]]) -> Graph:
-	half_edges = set([tuple(sorted((i, j))) for i in G_adj for j in G_adj[i]])
+	vertexpairs = set([tuple(sorted((i, j))) for i in G_adj for j in G_adj[i]])
 
-	vertexpair_to_node = dict([(e, i+1) for i,e in enumerate(half_edges)])
-	node_to_vertexpair = dict([(i+1, e) for i,e in enumerate(half_edges)])
+	vertexpair_to_node = dict([(e, i+1) for i,e in enumerate(vertexpairs)])
+	node_to_vertexpair = dict([(i+1, e) for i,e in enumerate(vertexpairs)])
 
-	medial = dict([(i+1, []) for i in range(len(half_edges))])
+	medial = dict([(i+1, []) for i in range(len(vertexpairs))])
 
 	for u,vs in G_adj.items():
 		nodes = [vertexpair_to_node[tuple(sorted((u, v)))] for v in vs]

### 2024-05-27 paper/main.tex
Fix grammar

@@
 
-	\todo{which? counting hamiltonian cycles of planar cubic graphs}
+	\todo{which? counting Hamiltonian cycles of planar cubic graphs}
 
@@
 
-	Here is some informal definitions to unpack the aforementioned properties of graphs.
+	Here are some informal definitions to unpack the aforementioned properties of graphs.
 
@@
 
-	Without the contraint of planarity on the input graph, then it is NP-complete to determine whether a general graph $G$ has a branch decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.
+	Without the constraint of planarity on the input graph, then it is NP-complete to determine whether a general graph $G$ has a branch decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.
 
@@
 
-	The width of a minimal branch decomposition of $G$ called the branch width of $G$.
+	The width of a minimal branch decomposition of $G$ is called the branch width of $G$.
 
 \section{The algorithm}
-	This section describes the algorithm given by Seymour and Thomas\ref{} by indentifying a set of practical problems and subproblems and how they relate.
+	This section describes the algorithm given by Seymour and Thomas\ref{} by identifying a set of practical problems and subproblems and how they relate.
 
@@
 
-	Implementing a function to solve \ref{problem:mcd-to-mbd} is decribed in \ref{impl:mcd-to-mbd}.
+	Implementing a function to solve \ref{problem:mcd-to-mbd} is described in \ref{impl:mcd-to-mbd}.
 
@@
 
-	I will refer to vertices and edges of the medial graph as "nodes" and "links" in an attempt of disambiguation.
+	I will refer to vertices and edges of the medial graph as "nodes" and "links" in an attempt at disambiguation.
 
-	Informally; The medial graph is a graph where there is a node in the medial for each edge, and an edge between two nodes if their corresponding edges are consecutive in a face of the graph.
+	Informally; The medial graph is a graph where there is a node in the medial for each edge, and an edge between two nodes if their corresponding edges are consecutive in some face of the graph.
 
@@
 
-	To solve \ref{problem:mcd} \ref{} gives an contraction algorithm.
+	To solve \ref{problem:mcd} \ref{} gives a contraction algorithm.
 
@@
 	Informally; a contraction is merging two vertices into one and letting incident edges connect to the new vertex.
-The contraction in this paper differs a bit from conventional definitions of an edge- and vertex-contractions, as other definitions might result in self-loops when contracting one of multiple parallel edges.
+The contraction in this paper differs a bit from conventional definitions of edge- and vertex-contractions, as other definitions might result in self-loops when contracting one of multiple parallel edges.
 
@@
 	The rat-catching algorithm can be described as a game of two players, the rat and rat-catcher.
-Considering a graph $M$, the edges of a face can be thought of as walls of a room and vertices as the corner of some rooms.
+Considering a graph $M$, the edges of a face can be thought of as walls of a room and vertices as the corners of some rooms.
 The rat moves from corner to corner along the walls and the rat-catcher moves from room to room through some wall.
@@
 
-	Additionally, if $\Delta(M) \geq k$ then the rat wins, the argument for why is this is true is glossed over in \ref{BGZ15}. This is discussed in section \ref{}.
+	Additionally, if $\Delta(M) \geq k$ then the rat wins. The argument for why this is true is glossed over in \ref{BGZ15}. This is discussed in section \ref{}.
 
@@
 
-	For some noise-level and location of the rat-catcher, exactly which edges are noisy and which are quiet, is definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
+	For some noise-level and location of the rat-catcher, exactly which edges are noisy and which are quiet is definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
 
@@
 
-	With the graph $H$, the only missing peace of the rat-catching algorithm is how to determine the outcome.
+	With the graph $H$, the only missing piece of the rat-catching algorithm is how to determine the outcome.
 
@@
 
-	I have chosen to assign IDs such that if one halfedge has ID $i$ then the other halfedge has ID $-i$, therefore the absolute value $|i|$ uniquely identifies an undirected edge.
+	I have chosen to assign IDs such that if one half-edge has ID $i$ then the other half-edge has ID $-i$, therefore the absolute value $|i|$ uniquely identifies an undirected edge.
 
@@
 
-		The clockwise ordering of neighborhoods of $G$ becomes counterclockwise orderings of neighborhoods of the medial $M$.
+		The clockwise ordering of neighborhoods of $G$ becomes counterclockwise ordering of neighborhoods of the medial $M$.
 The medial graph of a plane graph is 4-regular \ref{}.
-From the perpective of some medial node $v$, in some single iteration of the loop on line 14, two links are added to the neighborhood of $v$ in counterclockwise ordering, and latere the two links added to the neighborhood of $v$ also in counterclockwise ordering.
+From the perspective of some medial node $v$, in some single iteration of the loop on line 14, two links are added to the neighborhood of $v$ in counterclockwise ordering, and later the two other links are added to the neighborhood of $v$ also in counterclockwise ordering.
 
@@
 
-		The "contraction" function returns a new unique vertex ID, therefore by saving which vertex is a contraction of which vertex pair in the "edges" dictionary, contructing the decomposition is then a matter of recursively expanding any vertices that was a result of a contraction into a tuple of the vertex pair that is was composed of.
-Repeating this until all only vertices of $M$ remain gives a carving decomposition in newick-like nested tuple format.
+		The "contraction" function returns a new unique vertex ID, therefore by saving which vertex is a contraction of which vertex pair in the "edges" dictionary, constructing the decomposition is then a matter of recursively expanding any vertices that were a result of a contraction into a tuple of the vertex pair that is was composed of.
+Repeating this until all only vertices of $M$ remain gives a carving decomposition in Newick-like nested tuple format.
 
@@
 
-		First update any edges incident to either $a$ or $b$.
-Then creating the neighboorhood of the new vertex $c$ from the contraction of vertices $a$ and $b$, is done by firstly finding any shared edge $e$.
+		First, update any edges incident to either $a$ or $b$.
+Then creating the neighborhood of the new vertex $c$ from the contraction of vertices $a$ and $b$, is done by firstly finding any shared edge $e$.
 In this implementation the first shared edge $e$ in the neighborhood of $a$.
-This edge has some ID $e$ and the other halfedge with ID $-e$ will therefore be in the neighborhood of $b$.
+This edge has some ID $e$ and the other half-edge with ID $-e$ will therefore be in the neighborhood of $b$.
 Now  "rotating" the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0 in both lists means that a concatenation of the lists will preserve the ordering around the new vertex $c$.
-And finally remove any edges between $a$ and $b$.
+And finally, remove any edges between $a$ and $b$.
 
 		This is where telling apart parallel edges, which the Graph class allows, becomes very useful.
-Infering where to stitch together the neighborhoods to preserve the ordering, just from a normal adjacency list, becomes a way harder problem.
+Inferring where to stitch together the neighborhoods to preserve the ordering, just from a normal adjacency list, becomes a way harder problem.
 
@@
 
-		The vertices of game state graph $H$ is initialized by computing the elements of $T$ and $S$, while edges of $H$ are not explicitly kept in any datastructure, but instead checked while playing the game.
+		The vertices of the game state graph $H$ are initialized by computing the elements of $T$ and $S$, while edges of $H$ are not explicitly kept in any data structure, but instead checked while playing the game.
 
@@
 
-		Let $s_1$ and $t_1$ be the vertexpair for the link $e_1^*$ and let $s_2$ and $t_2$ be the vertexpair for the link $e_2^*$.
+		Let $s_1$ and $t_1$ be the vertex pair for the link $e_1^*$ and let $s_2$ and $t_2$ be the vertex pair for the link $e_2^*$.
 
@@
 
-		No other path of the implementation need the assumption that the dual is planar, therefore no clockwise or counterclockwise ordering of the neighborhoods of the adjacency list is needed.
+		No other path of the implementation needs the assumption that the dual is planar, therefore no clockwise or counterclockwise ordering of the neighborhoods of the adjacency list is needed.
 
 		The dual has a vertex for each face of the input graph.
-The faces are found by selecting an unmarked halfedge, and then marking all the edges of the face it belongs to, by following the edges which are just next to eachother in the ordered neighborhoods.
+The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to, by following the edges which are just next to each other in the ordered neighborhoods.
 

### 2024-05-27 paper/main.tex
Fix more grammar

@@
 The rat-catcher can force the rat away from some walls by making noise.
-A round of this game is played with some noise-level $k$.
-The rat-catcher wins if they can force the rat to be in some wall of the room that they are in, with noise-level $k$, and the rat wins if there is a strategy whereby the rat can escape indefinitely.
+A round of this game is played with some noise level $k$.
+The rat-catcher wins if they can force the rat to be in some wall of the room that they are in, with noise level $k$, and the rat wins if there is a strategy whereby the rat can escape indefinitely.
 
@@
 
-	For some noise-level and location of the rat-catcher, exactly which edges are noisy and which are quiet is definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
+	For some noise level and location of the rat-catcher, exactly which edges are noisy and which are quiet are definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
 
@@
 
-	A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise-level $k$ and some $e \in E(M)$, is a subgraph of $M$ with $V(Q(M, k, e)) = V(M)$ and
+	A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise level $k$ and some $e \in E(M)$, is a subgraph of $M$ with $V(Q(M, k, e)) = V(M)$ and
 
@@
 	\begin{problem}\label{problem:noisy-subgraph}
-		Given a plane graph $M$ that might have parallel edges, an edge $e \in E(M)$, and noise-level $k \in \mathbb{N}$, output the quiet subgraph $Q(M, k, e)$.
+		Given a plane graph $M$ that might have parallel edges, an edge $e \in E(M)$, and noise level $k \in \mathbb{N}$, output the quiet subgraph $Q(M, k, e)$.
 	\end{problem}
@@
 
-	The game states and possible moves, for some graph $M$ and some noise-level $k$, can be described as a graph $H(M, k)$.
+	The game states and possible moves, for some graph $M$ and some noise level $k$, can be described as a graph $H(M, k)$.
 
@@
 
-		The edges of the connected components are irrelevant for the rest of the algorithm, so only a list of vertices are returned for each connected component.
+		The edges of the connected components are irrelevant for the rest of the algorithm, so only a list of vertices is returned for each connected component.
 

### 2024-06-10 paper/main.tex
Update paper

@@
 
-\title{Implementation of a minimal branch-decomposition algorithm for simple planar graphs.}
+\title{Implementation of a minimal branch-decomposition algorithm for planar graphs.}
 \author{Kristoffer Højelse}
@@
 
+% west https://dwest.web.illinois.edu/grammar.html
+% The abstract states the results as fully as possible in a brief presentation.
+Crucial specialized terms the reader needs to know to understand the statements should be defined.
+The abstract stands on its own, especially in the age of electronic communication where it may be separate from the rest of the paper, and hence it contains no numbered reference to the bibliography.
+
 \begin{abstract}
@@
 
-\section{Motivation}
-	Some graph optimization problems can be solved efficiently for graphs of small branchwidth.\ref{}
+% The first section of the paper is an "Introduction" that should motivate the problem, discuss related results, state the results more completely, and perhaps summarize the techniques or the structure of the paper or crucial definitions.
+
+\section{Introduction}
+	Some graph optimization problems can be solved efficiently for graphs of small branchwidth.\cite{CNP+11}
+
+	Pino\cite{Pin16} applies branch decompositions.
+
+	Seymour and Thomas\cite{ST93} give the rat-catching algorithm.
+
+	Bian, Gu and Zhu\cite{BGZ15} describe and benchmark some implementations.
 
 	\todo{which? counting Hamiltonian cycles of planar cubic graphs}
+
+\section{Preliminaries}
+
+	A \textit{graph} $G$ is a 3-tuple $G = (V, E, \psi)$ consisting of a vertex set $V$, an edge set $E$ and a function $\psi$, where $V \subset \mathbb{N}^+$ and where $E \subset \mathbb{N}^+$ and where $\psi \st E \to \{\{u,v\} \st u,v \in V \land u \neq v \}$.
+
+	\textit{Note that other authors might instead name this an Undirected Labelled Multigraph (not permitting self-loops) (edges having own identity).}
+
+	A graph $G$ is \textit{simple}, if it has no parallel edges; $\neg \exists e_1,e_2 \in E_G$ where $e_1 \neq e_2$ such that $\psi_G(e_1) = \psi_G(e_1)$.
+
+	A \textit{subgraph} $H$ of a graph $G$, is a graph where some vertices and edges might be missing; a graph where $V_H \subseteq V_G$ and where $E_H \subseteq E_G$ and where $\psi_H$ is a subfunction of $\psi_G$; if $e \in E_H$ then $\psi_H(e) = \psi_G(e)$.
+
+	A graph is \textit{connected} if there exists a path connecting any two vertices.
+
+	A \textit{component} of a graph, is a connected subgraph.
+
+	A graph is \textit{planar} if there exists a planar embedding of the graph.
+
+	A \textit{plane graph} (syn.
+\textit{planar embedding of the graph}), is a 2D drawing of a planar graph such that no edges are crossing.
+In other words, for a planar graph $G$ a \textit{plane graph}, is a set of points $P$ and line segments $L$, with a one-to-one correspondence between $V(G)$ and $P$ and with a one-to-one correspondence between $E(G)$ and $L$, such that one of the endpoints of the line segments for the corresponding edges that are incident to some vertex $v$ occupy the same coordinate as the point corresponding to $v$ and such that line segments only pairwise intersect at their endpoints.
+
+	\begin{definition}
+		(\textit{Contraction})
+
+		Given an graph $G$ and pair of vertices $u,v \in G_V$ such that $\{u,v\}$ is in the image of $\psi_G$, remove all edges incident to $u$ and $v$ and update every edge $\{v,w\} \in E$ to be $\{u,w\}$.
+	\end{definition}
+
+	\begin{definition}
+		(\textit{Dual Graph})
+
+		The dual graph $M^*$ of a plane graph $M$ is a graph where a face $f$ of $M$ is a vertex $f^*$ of $M^*$ and an edge $e$ that separates two faces $f_1$,$f_2$ of $M$ is an edge $e^*$ of $M^*$ incident to $f_1^*$ and $f_2^*$.
+	\end{definition}
+
+	\begin{corollary}
+		If multiple edges separate $f_1$ and $f_2$ there will be parallel edges incident to $f_1*$ and $f_2*$.
+	\end{corollary}
+
+	\begin{definition}
+		(\textit{Medial Graph})
+
+		The medial graph $M(G)$ of a connected plane graph $G$ is a graph with a vertex $e^*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $d^*$ incident to a pair of vertices $e_1^*$, $e_2^*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
+	\end{definition}
+
+	\begin{corollary}
+		A medial graph is a 4-regular plane graph.
+	\end{corollary}
+
+	A \textit{Branch Decomposition} $B$ of a graph $G$ is a tree where every edge of $G$ is a leaf in $B$ and every internal vertex of $B$ has exactly 3 neighbors. $B$ is an unrooted binary tree.
+
+	Removing any edge $e = \{u,v\}$ of $B$ partitions $B$ into 2 trees $B_u$ and $B_v$ and the intersection of the sets of vertices in the leaves of $B_u$ and $B_v$ is called a middle set.
+Every edge of $B$ has an associated middle set.
+The maximal cardinality of any middle set among all middle sets of $B$ is the width of $B$.
+
+	There can be many branch decompositions of a graph $G$.
+
+	A \textit{Minimal Branch Decomposition} of $G$ is any branch decomposition of $G$ of minimal width among all branch decompositions of $G$.
+
+	Without the constraint of planarity on the input graph, then it is NP-complete to determine whether a general graph $G$ has a branch decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.
+
+	Let $n = |V(G)|$ and $m = |E(G)|$.
+
+	A \textit{Carving Decomposition}
+
+	A \textit{Minimal Carving Decomposition}
 
@@
 
-	A graph is called \textbf{simple} if and only if it has no parallel edges and no self-loops.
-
-	A graph is called \textbf{undirected} if and only if all its edges can be traversed in both directions.
-
-	A graph is called \textbf{connected} (or 1-vertex-connected) if and only if there exists a path between any two vertices.
-
-	A graph is called \textbf{planar} if and only if there is a way to draw the graph in 2 dimensions such that no pair of edges crosses.
-
-	A \textbf{branch decomposition} $B$ of a graph $G$ is a tree where every edge of $G$ is a leaf in $B$ and every internal vertex of $B$ has exactly 3 neighbors. $B$ is an unrooted binary tree.
-
-	Removing any edge $e = \{u,v\}$ of $B$ partitions $B$ into 2 trees $B_u$ and $B_v$ and the intersection of the sets of vertices in the leaves of $B_u$ and $B_v$ is called a middle set.
-Every edge of $B$ has an associated middle set.
-The maximal cardinality of any middle set among all middle sets of $B$ is the width of $B$.
-
-	There can be many branch decompositions of a graph $G$.
-
-	A \textbf{minimal branch decomposition} of $G$ is any branch decomposition of $G$ of minimal width among all branch decompositions of $G$.
-
-	Without the constraint of planarity on the input graph, then it is NP-complete to determine whether a general graph $G$ has a branch decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.
-
 	\begin{definition}\label{def:mbdp}
@@
 
-	The algorithm described in this paper solves \textsc{The Planar Minimal Branch Decomposition Problem} which can be computed in polynomial time.
+	The algorithm described in this paper solves \textsc{The Planar Minimal Branch Decomposition Problem}, which can be computed in polynomial time.
 
@@
 \section{The algorithm}
-	This section describes the algorithm given by Seymour and Thomas\ref{} by identifying a set of practical problems and subproblems and how they relate.
+	This section describes the algorithm given by Seymour and Thomas\cite{ST93} by identifying a set of practical problems and subproblems and how they relate.
 
@@
 
-	Informally; The medial graph is a graph where there is a node in the medial for each edge, and an edge between two nodes if their corresponding edges are consecutive in some face of the graph.
-
-	\begin{definition}
-		The medial graph $M(G)$ of a connected plane graph $G$ is a graph with a vertex $e^*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $d^*$ between a pair of vertices $e_1^*$, $e_2^*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
-	\end{definition}
-
 	Computing a medial graph is described in \ref{impl:medial}.
@@
 
-	Doing a series of edge contractions (contraction of all edges between a pair of vertices) on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices can be assembled into a minimal carving decomposition of $M$.
+	Doing a series of edge contractions (contraction of all edges incident to a pair of vertices) on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices can be assembled into a minimal carving decomposition of $M$.
 
@@
 	\begin{problem}\label{problem:contraction}
-		Given a graph $M$ that might have parallel edges, output a graph resulting from a contraction of all edges between a pair of vertices.
+		Given a graph $M$ and a pair of vertices $\{u, v\}$, output a graph resulting from a contraction of all edges incident to $u$ and $v$ if $u$ and $v$ are neighbors.
 	\end{problem}
@@
 The contraction in this paper differs a bit from conventional definitions of edge- and vertex-contractions, as other definitions might result in self-loops when contracting one of multiple parallel edges.
-
-	\begin{definition}
-		Contraction.
-
-		Given an undirected Graph $G=\{V,E\}$ with no self-loops and pair of vertices $u,v \in V$ such that $\{u,v\} \in E$, remove all edges between $u$ and $v$ and update every edge $\{v,w\} \in E$ to be $\{u,w\}$.
-	\end{definition}
 
@@
 
-	$E(Q(M, k, e)) = \{e_1 \mid$ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
+	$E(Q(M, k, e)) = \{e_1 \st $ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
 
@@
 
-	\begin{definition}
-		The dual graph $M^*$ of a plane graph $M$ is a graph where a face $f$ of $M$ is a vertex $f^*$ of $M^*$ and an edge $e$ that separates two faces $f_1$,$f_2$ of $M$ is an edge $e^*$ of $M^*$ between $f_1^*$ and $f_2^*$.
-	\end{definition}
-
-	\begin{corollary}
-		If multiple edges separate $f_1$ and $f_2$ there will be parallel edges between $f_1*$ and $f_2*$.
-	\end{corollary}
-
-	\begin{corollary}\label{def:dual_coro_loop}
-		If the same face is on both sides of $e$, then $e^*$ will be a self-loop.
-	\end{corollary}
-
-	For the algorithm in this paper, the class of graphs that will be given as input is medial graphs of simple undirected connected planar graphs. Therefore I claim \ref{claim:irri}.
-
-	\begin{claim}\label{claim:irri}
-		Corollary \ref{def:dual_coro_loop} will be irrelevant for any implementation of the algorithm.
-	\end{claim}
-
 	The game states and possible moves, for some graph $M$ and some noise level $k$, can be described as a graph $H(M, k)$.
@@
 
-	Let $S$ be every possible state when the rat-catcher is in a face some of which might be losing states. $S = \{ (f, v) \mid v \in V(M) \land f$ is a face of $ M \}$.
+	Let $S$ be every possible state when the rat-catcher is in a face some of which might be losing states. $S = \{ (f, v) \st v \in V(M) \land f$ is a face of $ M \}$.
 
-	Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \mid e \in E(M) \land C$ is a connected component of $Q(M, k, e) \}$.
+	Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \st e \in E(M) \land C$ is a component of $Q(M, k, e) \}$.
 
@@
 
-		As this contraction is a contraction of ALL edges between a pair of vertices, the resulting graph will not exhibit any self-loops.
+		As this contraction is a contraction of ALL edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
 I suspect reconciling this and the ordering invariant could be difficult, but luckily in this context, it is irrelevant.
@@
 Now  "rotating" the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0 in both lists means that a concatenation of the lists will preserve the ordering around the new vertex $c$.
-And finally, remove any edges between $a$ and $b$.
+And finally, remove any edges connecting $a$ and $b$.
 
@@
 
-		Considering a tuple $(f, v)$, if there exists a tuple $(e, C)$ which is losing where $e \in f$ and $v \in V(C)$ then $(f, v)$ is losing.
+		Considering a tuple $(f, v)$, if there exists a tuple $(e, C)$ that is losing where $e \in f$ and $v \in V(C)$ then $(f, v)$ is losing.
 
@@
 			The shortest closed walk that includes both $e_1^*$ and $e_2^*$ has the same length as either $$d(s_1, s_2) + d(t_1, t_2) + 2$$ or $$d(s_1, t_1) + d(s_2, t_2) + 2$$.
-Where $d(u, v)$ is the length of the shortest path from $u$ to $v$.
+Where $d(u, v)$ is the length of the shortest $u,v$-path.
 		\end{claim}
@@
 
-		Using the mapping from links of the dual to edges, and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained in the natural way.
+		Using the mapping from links to edges, and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained in the natural way.
 
-		Computing the quiet subgraph and the connected components thereof is done with a depth-first search approach.
+		Computing the quiet subgraph and the components thereof is done with a depth-first search approach.
 
-		The edges of the connected components are irrelevant for the rest of the algorithm, so only a list of vertices is returned for each connected component.
+		The edges of the components are irrelevant for the rest of the algorithm, so only a list of vertices is returned for each component.
 
@@
 		The dual has a vertex for each face of the input graph.
-The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to, by following the edges which are just next to each other in the ordered neighborhoods.
+The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to, by following the edges that are just next to each other in the ordered neighborhoods.
 
@@
 
-\section{Appendix.}
+\newpage
+\section{References}
+	\printbibliography
+
+\newpage
+\section{Appendix}
 	\pyfile{branch_decomposition.py}{0}{1000}

### 2024-06-10 code/carving_decomposition.py
Remove unused imports

@@ -1,8 +1,6 @@
 from carving_width import carving_width
 from contraction import contraction
 from parse_graph import parse_text_to_adj, adj_to_text
-from dual_graph import dual_graph
-from medial_graph import medial_graph
 from Graph import Graph
 
 # Find a contraction that does not increase the carving width

### 2024-06-24 paper/images/dual.tex
Add some examples to prelim

\begin{tikzpicture}
	\begin{scope}[xshift=0cm, yshift=0cm]
		\node (G) at (1, 6) {$G$};

		\node[circle, draw] (a1) at (0, 4) {1};
		\node[circle, draw] (a2) at (0, 2) {2};
		\node[circle, draw] (a3) at (0, 0) {3};
		\node[circle, draw] (a4) at (2, 2) {4};
		\node[circle, draw] (a5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (a1) -- node[left] {1} (a2);
		\draw[line width=0.5mm] (a2) -- node[left] {2} (a3);
		\draw[line width=0.5mm] (a2) -- node[above] {3} (a4);
		\draw[line width=0.5mm] (a4) -- node[right] {4} (a5);
		\draw[line width=0.5mm] (a3) to[bend left=40] node[above] {5} (a5);
		\draw[line width=0.5mm] (a3) to[bend right=40] node[below] {6} (a5);
		\draw[line width=0.5mm] (a4) to[in=125,out=55,loop, min distance=2cm] node[above] {7} (a4);
	\end{scope}

	\begin{scope}[xshift=6cm, yshift=0cm]
		\node (G) at (1, 6) {$G$ and $G^*$};

		\node[circle, draw] (a1) at (0, 4) {1};
		\node[circle, draw] (a2) at (0, 2) {2};
		\node[circle, draw] (a3) at (0, 0) {3};
		\node[circle, draw] (a4) at (2, 2) {4};
		\node[circle, draw] (a5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (a1) -- (a2);
		\draw[line width=0.5mm] (a2) -- (a3);
		\draw[line width=0.5mm] (a2) -- (a4);
		\draw[line width=0.5mm] (a4) -- (a5);
		\draw[line width=0.5mm] (a3) to[bend left=40] (a5);
		\draw[line width=0.5mm] (a3) to[bend right=40] (a5);
		\draw[line width=0.5mm] (a4) to[in=125,out=55,loop, min distance=2cm] (a4);

		\node[circle, draw, color=red, fill=white] (d1) at (0, 5) {1};
		\node[circle, draw, color=red, fill=white] (d2) at (2, 3) {2};
		\node[circle, draw, color=red, fill=white] (d3) at (1, 1.2) {3};
		\node[circle, draw, color=red, fill=white] (d4) at (1, 0) {4};

		\draw[line width=0.5mm, color=red] (d1) to[in=240, out=300, loop, min distance=2.5cm] (d1);
		\draw[line width=0.5mm, color=red] (d1) to[bend left=20] (d2);
		\draw[line width=0.5mm, color=red] (d1) to[bend left=30] (d3);
		\draw[line width=0.5mm, color=red] (d1) to[in=180, out=220, min distance=2cm] (d3);
		\draw[line width=0.5mm, color=red] (d1) to[in=0, out=0, min distance=3cm] (d3);
		\draw[line width=0.5mm, color=red] (d3) -- (d4);
		\draw[line width=0.5mm, color=red] (d1) to[in=240, out=200, min distance=3cm] (d4);
	\end{scope}

	\begin{scope}[xshift=12cm, yshift=0cm]
		\node (G) at (1, 6) {$G^*$};

		\node[circle, draw, color=red, fill=white] (d1) at (0, 5) {1};
		\node[circle, draw, color=red, fill=white] (d2) at (2, 3) {2};
		\node[circle, draw, color=red, fill=white] (d3) at (1, 1.2) {3};
		\node[circle, draw, color=red, fill=white] (d4) at (1, 0) {4};

		\draw[line width=0.5mm, color=red] (d1) to[in=240, out=300, loop, min distance=2.5cm] node[below] {1} (d1);
		\draw[line width=0.5mm, color=red] (d1) to[in=180, out=220, min distance=2cm] node[right] {2} (d3);
		\draw[line width=0.5mm, color=red] (d1) to[bend left=30] node[below left] {3} (d3);
		\draw[line width=0.5mm, color=red] (d1) to[in=0, out=0, min distance=3cm] node[right] {4} (d3);
		\draw[line width=0.5mm, color=red] (d3) to node[right] {5} (d4);
		\draw[line width=0.5mm, color=red] (d1) to[in=240, out=200, min distance=3cm] node[right] {6} (d4);
		\draw[line width=0.5mm, color=red] (d1) to[bend left=20] node[above right] {7} (d2);
	\end{scope}
\end{tikzpicture}

### 2024-06-24 paper/images/induced-subgraph.tex
Add some examples to prelim

\begin{tikzpicture}
	\begin{scope}[xshift=0cm, yshift=0cm]
		\node (G) at (1, 6) {$G$};

		\node[circle, draw] (1) at (0, 4) {1};
		\node[circle, draw] (2) at (0, 2) {2};
		\node[circle, draw] (3) at (0, 0) {3};
		\node[circle, draw] (4) at (2, 2) {4};
		\node[circle, draw] (5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (1) -- node[above] {1} (2);
		\draw[line width=0.5mm] (2) -- node[left] {2} (3);
		\draw[line width=0.5mm] (2) -- node[above] {3} (4);
		\draw[line width=0.5mm] (4) -- node[right] {4} (5);
		\draw[line width=0.5mm] (3) to[bend left=40] node[above] {5} (5);
		\draw[line width=0.5mm] (3) to[bend right=40] node[below] {6} (5);
		\draw[line width=0.5mm] (4) to[in=125,out=55,loop, min distance=1cm] node[above] {7} (4);
	\end{scope}

	\begin{scope}[xshift=6cm, yshift=0cm]
		\node (E0) at (0, 6) {$A$};
		\node (E1) at (0, 4) {\{1, 3, 4, 5\}};
	\end{scope}

	\begin{scope}[xshift=10cm, yshift=0cm]
		\node (G) at (1, 6) {$G[A]$};

		\node[circle, draw] (1) at (0, 4) {1};
		\node[circle, draw] (3) at (0, 0) {3};
		\node[circle, draw] (4) at (2, 2) {4};
		\node[circle, draw] (5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (4) -- node[right] {4} (5);
		\draw[line width=0.5mm] (3) to[bend left=40] node[above] {5} (5);
		\draw[line width=0.5mm] (3) to[bend right=40] node[below] {6} (5);
		\draw[line width=0.5mm] (4) to[in=125,out=55,loop, min distance=1cm] node[above] {7} (4);
	\end{scope}
\end{tikzpicture}

### 2024-06-24 paper/images/medial.tex
Add some examples to prelim

\begin{tikzpicture}
	\begin{scope}[xshift=0cm, yshift=0cm]
		\node (G) at (1, 6) {$G$};

		\node[circle, draw] (a1) at (0, 4) {1};
		\node[circle, draw] (a2) at (0, 2) {2};
		\node[circle, draw] (a3) at (0, 0) {3};
		\node[circle, draw] (a4) at (2, 2) {4};
		\node[circle, draw] (a5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (a1) -- node[left] {1} (a2);
		\draw[line width=0.5mm] (a2) -- node[left] {2} (a3);
		\draw[line width=0.5mm] (a2) -- node[above] {3} (a4);
		\draw[line width=0.5mm] (a4) -- node[right] {4} (a5);
		\draw[line width=0.5mm] (a3) to[bend left=40] node[above] {5} (a5);
		\draw[line width=0.5mm] (a3) to[bend right=40] node[below] {6} (a5);
		\draw[line width=0.5mm] (a4) to[in=125,out=55,loop, min distance=2cm] node[above] {7} (a4);
	\end{scope}

	\begin{scope}[xshift=6cm, yshift=0cm]
		\node (G) at (1, 6) {$G$ and $G^\times$};

		\node[circle, draw] (a1) at (0, 4) {1};
		\node[circle, draw] (a2) at (0, 2) {2};
		\node[circle, draw] (a3) at (0, 0) {3};
		\node[circle, draw] (a4) at (2, 2) {4};
		\node[circle, draw] (a5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (a1) -- (a2);
		\draw[line width=0.5mm] (a2) -- (a3);
		\draw[line width=0.5mm] (a2) -- (a4);
		\draw[line width=0.5mm] (a4) -- (a5);
		\draw[line width=0.5mm] (a3) to[bend left=40] (a5);
		\draw[line width=0.5mm] (a3) to[bend right=40] (a5);
		\draw[line width=0.5mm] (a4) to[in=125,out=55,loop, min distance=2cm] (a4);

		\node[circle, draw, color=blue, fill=white] (b1) at (0, 3) {1};
		\node[circle, draw, color=blue, fill=white] (b2) at (0, 1) {2};
		\node[circle, draw, color=blue, fill=white] (b3) at (1, 2) {3};
		\node[circle, draw, color=blue, fill=white] (b4) at (2, 1) {4};
		\node[circle, draw, color=blue, fill=white] (b5) at (1, 0.5) {5};
		\node[circle, draw, color=blue, fill=white] (b6) at (1, -0.5) {6};
		\node[circle, draw, color=blue, fill=white] (b7) at (2, 3.5) {7};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm, color=blue] (b1) to[in=125,out=55,loop, min distance=3cm] (b1);
		\draw[line width=0.5mm, color=blue] (b1) to[bend right=40, min distance=0.5cm] (b2);
		\draw[line width=0.5mm, color=blue] (b1) -- (b3);
		\draw[line width=0.5mm, color=blue] (b2) -- (b3);
		\draw[line width=0.5mm, color=blue] (b3) -- (b4);
		\draw[line width=0.5mm, color=blue] (b2) -- (b5);
		\draw[line width=0.5mm, color=blue] (b5) -- (b4);
		\draw[line width=0.5mm, color=blue] (b3) to[bend left=40, min distance=0.65cm] (b7);
		\draw[line width=0.5mm, color=blue] (b4) to[bend right=70, min distance=0.5cm] (b7);
		\draw[line width=0.5mm, color=blue] (b2) to[bend right=70, min distance=1.5cm] (b6);
		\draw[line width=0.5mm, color=blue] (b4) to[bend left=70, min distance=1.5cm] (b6);
		\draw[line width=0.5mm, color=blue] (b5) to[bend right=40] (b6);
		\draw[line width=0.5mm, color=blue] (b5) to[bend left=40] (b6);
	\end{scope}

	\begin{scope}[xshift=12cm, yshift=0cm]
		\node (G) at (1, 6) {$G^\times$};

		\node[circle, draw, color=blue, fill=white] (b1) at (0, 3) {1};
		\node[circle, draw, color=blue, fill=white] (b2) at (0, 1) {2};
		\node[circle, draw, color=blue, fill=white] (b3) at (1, 2) {3};
		\node[circle, draw, color=blue, fill=white] (b4) at (2, 1) {4};
		\node[circle, draw, color=blue, fill=white] (b5) at (1, 0.5) {5};
		\node[circle, draw, color=blue, fill=white] (b6) at (1, -0.5) {6};
		\node[circle, draw, color=blue, fill=white] (b7) at (2, 3.5) {7};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm, color=blue] (b1) to[in=125,out=55,loop, min distance=3cm] node[above] {1} (b1);
		\draw[line width=0.5mm, color=blue] (b1) to[bend right=40, min distance=0.5cm] node[left] {2} (b2);
		\draw[line width=0.5mm, color=blue] (b1) -- node[above right] {3} (b3);
		\draw[line width=0.5mm, color=blue] (b2) -- node[above left] {4} (b3);
		\draw[line width=0.5mm, color=blue] (b3) -- node[above right] {5} (b4);
		\draw[line width=0.5mm, color=blue] (b2) -- node[above right] {6} (b5);
		\draw[line width=0.5mm, color=blue] (b5) -- node[above left] {7} (b4);
		\draw[line width=0.5mm, color=blue] (b3) to[bend left=40, min distance=0.65cm] node[below right] {8} (b7);
		\draw[line width=0.5mm, color=blue] (b4) to[bend right=70, min distance=0.5cm] node[right] {9} (b7);
		\draw[line width=0.5mm, color=blue] (b2) to[bend right=70, min distance=1.5cm] node[below left] {10} (b6);
		\draw[line width=0.5mm, color=blue] (b4) to[bend left=70, min distance=1.5cm] node[below right] {11} (b6);
		\draw[line width=0.5mm, color=blue] (b5) to[bend right=40] node[left] {12} (b6);
		\draw[line width=0.5mm, color=blue] (b5) to[bend left=40] node[right] {13} (b6);
	\end{scope}
\end{tikzpicture}

### 2024-06-24 paper/images/multigraph.tex
Add some examples to prelim

\begin{tikzpicture}
    % Define the nodes with specific positions
    \begin{scope}[xshift=0cm, yshift=0cm]
        \node (G) at (1, 6) {A multi-graph $G$};

        \node[circle, draw, fill=black, minimum size=0.01mm] (1) at (0, 4) {};
        \node[circle, draw, fill=black, minimum size=0.01mm] (2) at (0, 2) {};
        \node[circle, draw, fill=black, minimum size=0.01mm] (3) at (0, 0) {};
        \node[circle, draw, fill=black, minimum size=0.01mm] (4) at (2, 2) {};
        \node[circle, draw, fill=black, minimum size=0.01mm] (5) at (2, 0) {};

        % Draw the edges between nodes
        \draw[line width=0.5mm] (1) -- (2);
        \draw[line width=0.5mm] (2) -- (3);
        \draw[line width=0.5mm] (2) -- (4);
        \draw[line width=0.5mm] (4) -- (5);
        \draw[line width=0.5mm] (3) to[bend left=40] (5);
        \draw[line width=0.5mm] (3) to[bend right=40] (5);
    \end{scope}
\end{tikzpicture}

### 2024-06-24 paper/main.tex
Add some examples to prelim

@@
 
-	A \textit{graph} $G$ is a 3-tuple $G = (V, E, \psi)$ consisting of a vertex set $V$, an edge set $E$ and a function $\psi$, where $V \subset \mathbb{N}^+$ and where $E \subset \mathbb{N}^+$ and where $\psi \st E \to \{\{u,v\} \st u,v \in V \land u \neq v \}$.
+	A \textit{graph} $G$ consists of a vertex set $V(G)$, and an edge set $\mathbb{E}(G)$ and a function $\phi_G$, where $V(G) \subset \mathbb{N}^+$ and where $\mathbb{E}(G) \subset \mathbb{N}^+$ and where $\phi_G \st \mathbb{E}(G) \to \{\{u,v\} \st u,v \in V(G)\}$.
 
-	\textit{Note that other authors might instead name this an Undirected Labelled Multigraph (not permitting self-loops) (edges having own identity).}
+	\textbf{Note.} Other authors might instead call this definition an undirected labelled pseudograph, with edges having own identity.
 
-	A graph $G$ is \textit{simple}, if it has no parallel edges; $\neg \exists e_1,e_2 \in E_G$ where $e_1 \neq e_2$ such that $\psi_G(e_1) = \psi_G(e_1)$.
+	\textbf{Note.} Regarding notation, $V$ and $\mathbb{E}$ are operations on graphs returning the vertex set and edge set respectively.
 
-	A \textit{subgraph} $H$ of a graph $G$, is a graph where some vertices and edges might be missing; a graph where $V_H \subseteq V_G$ and where $E_H \subseteq E_G$ and where $\psi_H$ is a subfunction of $\psi_G$; if $e \in E_H$ then $\psi_H(e) = \psi_G(e)$.
+	A \textit{drawing} of a graph $G$ is a node-link diagram in which the vertices are represented as disks and the edges are represented as line segments or curves in the Euclidean plane.
 
-	A graph is \textit{connected} if there exists a path connecting any two vertices.
+	Here is a drawing of a graph $G$.
+
+	\begin{center}
+		\input{images/unlabeled-graph.tex}
+	\end{center}
+
+	Here is a labeled drawing of the same graph $G$ and its function $\phi_G$.
+
+	\begin{center}
+		\input{images/graph.tex}
+	\end{center}
+
+	Let $E(G)$ return a multiset of all vertex-pairs of $G$; in other words, $E(G) = \{\phi_G(e) \st e \in \mathbb{E}(G)\}$.
+
+	A edge $e$ where $\phi_G(e) = \{u,v\}$, is a \textit{self-loop}, if $u = v$.
+
+	A graph $G$ is \textit{loop-less}, if no edge $e \in \mathbb{E}(G)$ is a self-loop.
+
+	A \textit{multi-graph} is a graph that is loop-less.
+
+	\begin{center}
+		\input{images/multigraph.tex}
+	\end{center}
+
+	A multi-graph $G$ is \textit{simple}, if it has no parallel edges; in other words, if elements of $E(G)$ are pair-wise distinct.
+
+	\begin{center}
+		\input{images/simple-graph.tex}
+	\end{center}
+
+	A \textit{subgraph} $H$ of a graph $G$, is a graph where some vertices and edges might be missing; in other words, is a graph where $V(H) \subseteq V(G)$ and where $\mathbb{E}(H) \subseteq \mathbb{E}(G)$ and where $\forall e \in \mathbb{E}(H), \phi_H(e) = \phi_G(e)$.
+
+	\begin{center}
+		\input{images/subgraph.tex}
+	\end{center}
+
+	For $A \subseteq V(G)$, we denote by $G[A]$ the subgraph induced by the subset of vertices $A$; in other words, $G[A]$ is the subgraph where $V(G[A]) = A$ and where $\mathbb{E}(G[A]) = \{ e \st e \in \mathbb{E}(G) \land |\phi_G(e) \cap A| = 2\}$ and where $\forall e \in E(G[A]), \phi_{G[A]}(e) = \phi_{G}(e)$.
+
+	\begin{center}
+		\input{images/induced-subgraph.tex}
+	\end{center}
+
+	A vertex $v \in V(G)$ and an edge $e \in \mathbb{E}(G)$ are \textit{incident} to each other, if $v \in \phi_G(e)$.
+Furthermore, two distinct edges $e_1,e_2 \in \mathbb{E}(G)$ are incident to each other, if $\phi_G(e_1) \cap \phi_G(e_2) \neq \emptyset$.
+
+	The \textit{degree} of a vertex $v$, denoted $\deg(v)$, is the number of times that an edge is incident to $v$. A self-loop is incident to the same vertex twice.
+
+	\begin{center}
+		\input{images/degree.tex}
+	\end{center}
+
+	The \textit{maximum degree} of a graph $G$, denoted $\Delta(G)$, is the maximal degree of any vertex of $G$.
+
+	A \textit{walk} of a graph $G$ is a list of edges, such that consecutive edges in the list are incident to each other.
+
+	An \textit{$s,t$-walk} is a walk, such that the first edge is incident to the vertex $s$ and such that the last edge is incident to the vertex $t$.
+
+	The \textit{length} of a walk is the number of edges in the walk.
+
+	An $s,t$-walk is \textit{closed}, if $s=t$.
+
+	A \textit{path} of a graph $G$, is a walk such that only consecutive edges of the walk are incident.
+
+	A \textit{cycle} of a graph $G$, is a path but where the first and the last edge are incident to each other.
+
+	A graph $G$ is \textit{connected} if there exists a $s,t$-walk for every pair of distinct vertices $s,t \in V(G)$.
 
@@
 
-	A graph is \textit{planar} if there exists a planar embedding of the graph.
+	A \textit{bijection} is a relation between two sets such that each element of either set is paired with exactly one element of the other set.
 
-	A \textit{plane graph} (syn.
-\textit{planar embedding of the graph}), is a 2D drawing of a planar graph such that no edges are crossing.
-In other words, for a planar graph $G$ a \textit{plane graph}, is a set of points $P$ and line segments $L$, with a one-to-one correspondence between $V(G)$ and $P$ and with a one-to-one correspondence between $E(G)$ and $L$, such that one of the endpoints of the line segments for the corresponding edges that are incident to some vertex $v$ occupy the same coordinate as the point corresponding to $v$ and such that line segments only pairwise intersect at their endpoints.
+	A \textit{plane graph} is a drawing of a graph, such that no edges are crossing.
+
+	A graph $G$ is \textit{planar}, if there exists a \textit{plane graph} of $G$.
 
@@
 
-		Given an graph $G$ and pair of vertices $u,v \in G_V$ such that $\{u,v\}$ is in the image of $\psi_G$, remove all edges incident to $u$ and $v$ and update every edge $\{v,w\} \in E$ to be $\{u,w\}$.
+		A contraction is a function that given a multi-graph $G$ and pair of vertices $u,v \in V(G)$ such that $\{u,v\} \in E(G)$, then for all edges $e \in \mathbb{E}(G)$ if $\phi_G(e) = \{u,v\}$ then removes $e$ else if $\phi_G(e) = \{v,w\}$ then $\phi_G(e) = \{u,w\}$, and finally returns the resulting graph.
 	\end{definition}
 
-	\begin{definition}
-		(\textit{Dual Graph})
-
-		The dual graph $M^*$ of a plane graph $M$ is a graph where a face $f$ of $M$ is a vertex $f^*$ of $M^*$ and an edge $e$ that separates two faces $f_1$,$f_2$ of $M$ is an edge $e^*$ of $M^*$ incident to $f_1^*$ and $f_2^*$.
-	\end{definition}
-
-	\begin{corollary}
-		If multiple edges separate $f_1$ and $f_2$ there will be parallel edges incident to $f_1*$ and $f_2*$.
-	\end{corollary}
+	\begin{center}
+		\input{images/contraction.tex}
+	\end{center}
 
@@
 
-		The medial graph $M(G)$ of a connected plane graph $G$ is a graph with a vertex $e^*$ for each edge $e$ of $G$ and for each face $f$ of $G$, there's an edge $d^*$ incident to a pair of vertices $e_1^*$, $e_2^*$ of $M(G)$ if $e_1$ and $e_2$ are consecutive in $f$.
+		The medial graph $G^\times$ of a connected plane graph $G$ is a graph such that there is a bijection between $V(G^\times)$ and $\mathbb{E}(G)$ and such that for each face $f$ of $G$, there's an edge $e^\times \in \mathbb{E}(G^\times)$ incident to a pair of vertices $u^\times,v^\times \in V(G^\times)$ if edges $u,v \in \mathbb{E}(G)$ are consecutive in $f$.
 	\end{definition}
+
+	\begin{center}
+		\input{images/medial.tex}
+	\end{center}
 
@@
 
-	A \textit{Branch Decomposition} $B$ of a graph $G$ is a tree where every edge of $G$ is a leaf in $B$ and every internal vertex of $B$ has exactly 3 neighbors. $B$ is an unrooted binary tree.
+	\begin{definition}
+		(\textit{Dual Graph})
 
-	Removing any edge $e = \{u,v\}$ of $B$ partitions $B$ into 2 trees $B_u$ and $B_v$ and the intersection of the sets of vertices in the leaves of $B_u$ and $B_v$ is called a middle set.
-Every edge of $B$ has an associated middle set.
-The maximal cardinality of any middle set among all middle sets of $B$ is the width of $B$.
+		The dual graph $G^*$ of a plane graph $G$ is a graph with a bijection between the set of faces of $G$ and $V(G^*)$ and a bijection between $\mathbb{E}(G)$ and $\mathbb{E}(G^*)$ such that an edge $e \in \mathbb{E}(G)$ that separates two faces $f_1$,$f_2$ of $G$ is an edge $e^* \in \mathbb{E}(G^*)$ incident to $f_1^*$ and $f_2^*$.
+	\end{definition}
 
-	There can be many branch decompositions of a graph $G$.
+	\begin{center}
+		\input{images/dual.tex}
+	\end{center}
+
+	A \textit{tree} is a connected graph with no cycles.
+
+	A \textit{leaf} $v$ of a tree $T$, is a vertex $v \in V(T)$ of degree 1.
+
+	An \textit{internal vertex} $v$ of a tree $T$, is a vertex $v \in V(T)$ of degree at least 2.
+
+	Let the \textit{leaf set} of a tree $T$, denoted $L(T)$, be the set of all vertices of $T$ that are also leaves.
+
+	A \textit{Branch Decomposition} $(B_G, \delta_G)$ of a graph $G$ consists of firstly, a tree $B_G$ where every internal vertex of $B_G$ has degree 3; in other words, $B_G$ is an unrooted binary tree, and secondly a bijection $\delta_G$ between $\mathbb{E}(G)$ and $L(B_G)$.
+
+	Removing any edge $e \in \mathbb{E}(B_G)$ partitions $B_G$ into 2 trees $P_e$ and $Q_e$.
+The set $L(P_e) \cap L(Q_e)$ is called a \textit{middle set} of $B_G$ given $e$, denoted $Z(B_G,e)$.
+The maximal cardinality of any middle set of $B_G$ given any $e$ of $B_G$ is the width of $B_G$; in other words, the width of $B_G$ is $\max\{ |Z(B, e)| \st e \in \mathbb{E}(B) \}$.
+
+	There might exist many branch decompositions of a graph $G$.
 
@@
 
-	Without the constraint of planarity on the input graph, then it is NP-complete to determine whether a general graph $G$ has a branch decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.
+	A \textit{Carving Decomposition} $(C_G, \lambda_G)$ of a graph $G$ consists of firstly, a tree unrooted binary tree $C_G$ and secondly a bijection $\lambda_G$ between $V(G)$ and $L(C_G)$.
 
-	Let $n = |V(G)|$ and $m = |E(G)|$.
+	Removing any edge $e \in \mathbb{E}(C_G)$ partitions $C_G$ into 2 trees $P_e$ and $Q_e$.
+The set $L(P_e) \cap L(Q_e)$ is called a \textit{middle set} of $C_G$ given $e$, denoted $Z(C_G,e)$.
+The maximal cardinality of any middle set of $C_G$ given any $e$ of $C_G$ is the width of $C_G$; in other words, the width of $C_G$ is $\max\{ |Z(C_G, e)| \st e \in \mathbb{E}(C_G) \}$.
 
-	A \textit{Carving Decomposition}
+	There might exist many carving decompositions of a graph $G$.
 
-	A \textit{Minimal Carving Decomposition}
+	A \textit{Minimal Carving Decomposition} of $G$ is any carving decomposition of $G$ of minimal width among all carving decompositions of $G$.
 
@@
 
-		Input: Given a simple undirected connected planar graph $G$.
+		Input: Given a simple connected planar graph $G$.
 
@@
 
-		Input: Given a simple undirected connected graph $G$.
+		Input: Given a simple connected graph $G$.
 
@@
 
-	Problem \ref{def:mbdp} is the overarching problem, that the algorithm solves, and can be broken down into many smaller subproblems.
+	Problem \ref{def:mbdp} is the overarching problem, and can be broken down into many smaller subproblems.
 
-	Considering a graph $G$, you can compute a minimal branch decomposition of $G$ from a minimal carving decomposition of the medial of $G$, by replacing the vertices in leaves of the decomposition with edges, using the mapping between edges and vertices from computing the medial graph.
-\ref{} Therefore problem \ref{def:mbdp} break down into problems \ref{problem:mcd-to-mbd}, \ref{problem:medial} and \ref{problem:mcd}.
+	Considering a graph $G$, you can compute a minimal branch decomposition $(B_G, \delta_G)$ of $G$ from a minimal carving decomposition $(C_{G^\times}, \lambda_{G^\times})$ of the medial graph $G^\times$ of $G$.
+Do this by replacing the vertices in leaves of the carving decomposition with edges, using the mapping between edges and vertices from computing the medial graph;
+
+	Therefore problem \ref{def:mbdp} break down into problems \ref{problem:mcd-to-mbd}, \ref{problem:medial} and \ref{problem:mcd}.
 
@@
 
-	A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise level $k$ and some $e \in E(M)$, is a subgraph of $M$ with $V(Q(M, k, e)) = V(M)$ and
+	A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise level $k$ and some $e \in \mathbb{E}(M)$, is a subgraph of $M$ with $V(Q(M, k, e)) = V(M)$ and
 
-	$E(Q(M, k, e)) = \{e_1 \st $ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
+	$\mathbb{E}(Q(M, k, e)) = \{e_1 \st $ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
 
 	\begin{problem}\label{problem:noisy-subgraph}
-		Given a plane graph $M$ that might have parallel edges, an edge $e \in E(M)$, and noise level $k \in \mathbb{N}$, output the quiet subgraph $Q(M, k, e)$.
+		Given a plane graph $M$ that might have parallel edges, an edge $e \in \mathbb{E}(M)$, and noise level $k \in \mathbb{N}$, output the quiet subgraph $Q(M, k, e)$.
 	\end{problem}
@@
 
-	Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \st e \in E(M) \land C$ is a component of $Q(M, k, e) \}$.
+	Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \st e \in \mathbb{E}(M) \land C$ is a component of $Q(M, k, e) \}$.
 
@@
 
-		I assume that the input graph $G$ is a planar graph given as an adjacency list such that the neighborhoods are given in clockwise ordering according to some plane embedding of $G$.
+		I assume that the input graph $G$ is a rotation system of a planar graph; an adjacency list such that the neighborhoods are given in clockwise ordering according to some plane embedding of $G$.
 
@@
 
-		As the resulting graph is later given as an argument to functions assuming a clockwise or counterclockwise ordering of vertices, the implementation needs to preserve this invariant when contracting.
+		As the resulting graph is later given as an argument to functions assuming a rotation system of a planar graph, the implementation needs to preserve this invariant when contracting.
 
 		As this contraction is a contraction of ALL edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
-I suspect reconciling this and the ordering invariant could be difficult, but luckily in this context, it is irrelevant.
+I suspect reconciling this and the rotation system could be difficult, but luckily in this context, it is irrelevant.
 
@@
 
+		% https://en.wikipedia.org/wiki/Rotation_system
+
 		The next halfedge $e_{i+1}$ after the current halfedge $e_i = \{u, v\}$ is the edge just before $-e_i$ in the ordered neighborhood around $v$.

### 2024-06-24 code/branch_decomposition.py
Update comments

@@ -9,7 +9,6 @@ def branch_decomposition(G_adj: dict[int, list[int]]):
 	cd = carving_decomposition(M)
 
 	# Convert the carving decomposition of M to a branch decomposition of G
-
 	def decomp(t):
 		if isinstance(t, int):
 			return node_to_vertexpair[t]

### 2024-06-24 code/count_hamcyc_dp_thrice.py

@@ -3,7 +3,7 @@ from parse_graph import parse_graph_to_adj
 
 def find_faces(adj: dict[int, list[int]]) -> list[list[(int, int)]]:
 	# assume planar graph
-	# assume clockwise ordering of neighbors
+	# assume adj is a rotation system
 	edges = set([(i, j) for i in adj for j in adj[i]])
 	faces = []
 	while len(edges) > 0:

### 2024-06-24 code/count_hamcyc_dp_twice.py

@@ -3,7 +3,7 @@ from parse_graph import parse_graph_to_adj
 
 def find_faces(adj: dict[int, list[int]]) -> list[list[(int, int)]]:
 	# assume planar graph
-	# assume clockwise ordering
+	# assume adj is a rotation system
 	edges = set([(i, j) for i in adj for j in adj[i]])
 	faces = []
 	while len(edges) > 0:

### 2024-06-24 code/dual_graph.py

@@ -1,15 +1,17 @@
 from Graph import Graph
 from parse_graph import adj_to_text, parse_text_to_adj
 
+# Assume G is a rotation system
 def dual_graph(G: Graph) -> Graph:
 	edges = [e for e in G.E()]
 
 	D = Graph()
 	edge_to_link = dict()
 	link_to_edge = dict()
-	node_to_face = dict()
-	edge_to_node = dict()
+	node_to_face = dict() # nodeid to edgeid list
+	edge_to_node = dict() # half-edge to the faceid/node to its either left/right
 
+	# Find faces
 	next_nodeid = -1
 	while edges:
 		e = edges.pop()
@@ -31,6 +33,7 @@ def dual_graph(G: Graph) -> Graph:
 	for i in node_to_face.keys():
 		D.adj_edges[i] = []
 
+	# Add edges to dual graph
 	next_linkid = 1
 	for i,f1 in node_to_face.items():
 		for j,f2 in node_to_face.items():
@@ -46,6 +49,9 @@ def dual_graph(G: Graph) -> Graph:
 					D.adj_edges[i].append(next_linkid)
 					D.adj_edges[j].append(-next_linkid)
 					next_linkid += 1
+
+	# todo make edge_to_link and link_to_edge redundant
+	# by nameing edges and links the same
 
 	return D, edge_to_link, link_to_edge, node_to_face, edge_to_node
 

### 2024-06-24 code/medial_graph.py

@@ -2,7 +2,7 @@ from Graph import Graph
 from parse_graph import adj_to_text, parse_text_to_adj
 
 # assume planar graph
-# assume clockwise ordering of neighbors
+# assume G_adj is a rotation system
 def medial_graph(G_adj: dict[int, list[int]]) -> Graph:
 	vertexpairs = set([tuple(sorted((i, j))) for i in G_adj for j in G_adj[i]])
 

### 2024-08-12 paper/images/branch.tex
Update paper

\begin{tikzpicture}
	\begin{scope}[xshift=0cm, yshift=0cm]
		\node (G) at (1, 6) {$G$};

		\node[circle, draw] (a1) at (0, 4) {1};
		\node[circle, draw] (a2) at (0, 2) {2};
		\node[circle, draw] (a3) at (0, 0) {3};
		\node[circle, draw] (a4) at (2, 2) {4};
		\node[circle, draw] (a5) at (2, 0) {5};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm] (a1) -- node[left] {1} (a2);
		\draw[line width=0.5mm] (a2) -- node[left] {2} (a3);
		\draw[line width=0.5mm] (a2) -- node[above] {3} (a4);
		\draw[line width=0.5mm] (a4) -- node[right] {4} (a5);
		\draw[line width=0.5mm] (a3) -- node[above] {5} (a5);
	\end{scope}

	\begin{scope}[xshift=6cm, yshift=0cm]
		\node (G) at (1, 6) {Branch Decomposition of $G$};

		\node[circle, draw, color=black, fill=white] (b1) at (1, 4) {(1,2)};
		\node[circle, draw, color=black, fill=white] (b3) at (0, 3) {(2,4)};
		\node[circle, draw, color=black, fill=white] (i1) at (1, 3) {};
		\node[circle, draw, color=black, fill=white] (i2) at (1.5, 2) {};
		\node[circle, draw, color=black, fill=white] (i3) at (1, 1) {};
		\node[circle, draw, color=black, fill=white] (b2) at (2.5, 2) {(2,3)};
		\node[circle, draw, color=black, fill=white] (b4) at (0, 1) {(4,5)};
		\node[circle, draw, color=black, fill=white] (b5) at (1, 0) {(3,5)};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm, color=black] (b1) -- (i1);
		\draw[line width=0.5mm, color=black] (b3) -- (i1);
		\draw[line width=0.5mm, color=black] (i1) -- (i2);
		\draw[line width=0.5mm, color=black] (b2) -- (i2);
		\draw[line width=0.5mm, color=black] (i2) -- (i3);
		\draw[line width=0.5mm, color=black] (i3) -- (b4);
		\draw[line width=0.5mm, color=black] (i3) -- (b5);
	\end{scope}
\end{tikzpicture}

### 2024-08-12 paper/main.tex
Update paper

@@
 
-	A edge $e$ where $\phi_G(e) = \{u,v\}$, is a \textit{self-loop}, if $u = v$.
+	A \textit{self-loop}, is an edge $e$ where $\phi_G(e) = \{u,v\}$ and $u = v$.
 
-	A graph $G$ is \textit{loop-less}, if no edge $e \in \mathbb{E}(G)$ is a self-loop.
+	A graph $G$ is \textit{loop-less}, if no edge $e \in \mathbb{E}(G)$ is a self-loop. The edge with label 7, in the graph above, is an example of a self-loop.
 
@@
 
-	A \textit{walk} of a graph $G$ is a list of edges, such that consecutive edges in the list are incident to each other.
+	$$\Delta(G) = 4$$
 
-	An \textit{$s,t$-walk} is a walk, such that the first edge is incident to the vertex $s$ and such that the last edge is incident to the vertex $t$.
+	A \textit{walk} of a graph $G$ is a list $[v_0,e_1,v_1,...,e_k,v_k]$ where $v_0,v_1,...,v_k \in V(G)$ and for $1 \leq i \leq k$, $\phi_{G}(e_i) = \{v_{i-1}, v_i\}$.
+
+	A walk of $G$ $$[4,7,4,4,5,5,3,5,5,6,3]$$
 
@@
 
-	An $s,t$-walk is \textit{closed}, if $s=t$.
+	An \textit{s,t-walk} is a walk where $s = v_0$ and $t = v_k$.
 
-	A \textit{path} of a graph $G$, is a walk such that only consecutive edges of the walk are incident.
+	An $s,t$-walk is \textit{closed}, if $s = t$.
 
-	A \textit{cycle} of a graph $G$, is a path but where the first and the last edge are incident to each other.
+	A \textit{path} of a graph $G$, is a walk such that no vertex is repeated in the list.
+
+	A \textit{cycle} of a graph $G$, is an $s,t$-walk such that no vertex is repeated in the list except $s=t$.
 
@@
 
-	A \textit{bijection} is a relation between two sets such that each element of either set is paired with exactly one element of the other set.
+	A \textit{bijection} (or \textit{one-to-one correspondence}) is a relation between two sets such that each element of either set is paired with exactly one element of the other set.
 
@@
 
-	\begin{definition}
+	A \textit{rotation system} is an encoding of a graph, in particular, it is an adjacency list such that the neighborhood, around any vertex, is in clockwise ordering according to some plane embedding of $G$.
+
+	\begin{definition}\label{def:contraction}
 		(\textit{Contraction})
 
-		A contraction is a function that given a multi-graph $G$ and pair of vertices $u,v \in V(G)$ such that $\{u,v\} \in E(G)$, then for all edges $e \in \mathbb{E}(G)$ if $\phi_G(e) = \{u,v\}$ then removes $e$ else if $\phi_G(e) = \{v,w\}$ then $\phi_G(e) = \{u,w\}$, and finally returns the resulting graph.
+		A contraction is a function that given a multi-graph $G$ and pair of distinct vertices $u,v \in V(G)$ such that $\{u,v\} \in E(G)$, then for all edges $e \in \mathbb{E}(G)$ if $\phi_G(e) = \{u,v\}$ then removes $e$ else if $\phi_G(e) = \{v,w\}$ then $\phi_G(e) = \{u,w\}$, and finally returns the resulting graph.
 	\end{definition}
@@
 
-	\begin{definition}
+	\begin{definition}\label{def:medial}
 		(\textit{Medial Graph})
@@
 	\end{corollary}
+
+	I will refer to vertices and edges of the medial graph as "nodes" and "links" in an attempt at disambiguation.
 
@@
 
-	An \textit{internal vertex} $v$ of a tree $T$, is a vertex $v \in V(T)$ of degree at least 2.
+	Let the \textit{leaf set} of a tree $T$, denoted $L(T)$, be the subset of vertices $L(T) \subseteq V(T)$ that are also leaves of $T$.
 
-	Let the \textit{leaf set} of a tree $T$, denoted $L(T)$, be the set of all vertices of $T$ that are also leaves.
+	An \textit{internal vertex} $v$ of a tree $T$, is a vertex $v \in V(T) - L(T)$ that is not a leaf. An internal vertex therefore has degree at least 2.
 
-	A \textit{Branch Decomposition} $(B_G, \delta_G)$ of a graph $G$ consists of firstly, a tree $B_G$ where every internal vertex of $B_G$ has degree 3; in other words, $B_G$ is an unrooted binary tree, and secondly a bijection $\delta_G$ between $\mathbb{E}(G)$ and $L(B_G)$.
+	An \textit{unrooted binary tree} $T$, is a tree where every internal vertex has degree 3.
+
+	A \textit{Branch Decomposition} $(B_G, \delta_G)$ of a simple graph $G$ consists of firstly, an unrooted binary tree $B_G$ and secondly a bijection $\delta_G$ between $\mathbb{E}(G)$ and $L(B_G)$.
+
+	\begin{center}
+		\input{images/branch.tex}
+	\end{center}
 
@@
 
-	A \textit{Carving Decomposition} $(C_G, \lambda_G)$ of a graph $G$ consists of firstly, a tree unrooted binary tree $C_G$ and secondly a bijection $\lambda_G$ between $V(G)$ and $L(C_G)$.
+	A \textit{Carving Decomposition} $(C_G, \lambda_G)$ of a simple graph $G$ consists of firstly, an unrooted binary tree $C_G$ and secondly a bijection $\lambda_G$ between $V(G)$ and $L(C_G)$.
 
@@
 
-\section{Description of the problem}
+\section{Overview of the algorithm}
+
+	Given a simple connected planar graph $G$, the algorithm computes a minimal branch decomposition of $G$.
+
+	To compute a minimal branch decomposition of $G$, the algorithm first computes the medial graph $G^\times$ of $G$, secondly computes a minimal carving decomposition of $G^\times$, and finally computes a minimal branch decomposition of $G$ from the minimal carving decomposition of $G^\times$.
+
+	To compute the minimal carving decomposition of $G^\times$, the algorithm uses a contraction algorithm that finds a series of contractions that does not increase the carving width and then assembles the series of contractions into a minimal carving decomposition of $G^\times$.
+
+	To compute the carving width of $G$, the algorithm uses the rat-catching algorithm.
+By an analogy, the rat-catching algorithm can be described by a game of two players, the rat and rat-catcher, who take turns moving around on $G$.
+The game is played to determine, if the rat-catcher can corner the rat by making noise and thereby scaring away the rat from some subgraph, the size of which is dependent on $k$ or if the rat can escape indefinitely.
+Larger $k$ makes the rat-catching game easier for the rat-catcher.
+The smallest $k$ where the rat can escape indefinitely is the carving width of $G$.
+
+	To compute a noisy subgraph of $G$ given a noise level $k$ and a rat-catchers position, the algorithm considers the dual graph of $G$ and finds the set of dual edges reachable by a closed walk of length at most $k-1$, the dual of these edges is the noisy subgraph.
+
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=\textwidth]{./images/overview.png}
+		\caption{Overview of the algorithm}
+		\label{fig:overview}
+	\end{figure}
+
+\section{Data Structures and implementation considerations}
+
+	For many subroutines the algorithm needs to deal with parallel edges and be able to tell them apart, therefore the implementation encodes a graph as an adjacency list of edges and a map from unique edge ids its vertexpair.
+
+	I have chosen to assign IDs such that if one half-edge has ID $i$ then the other half-edge has ID $-i$, therefore the absolute value $|i|$ uniquely identifies an undirected edge.
+
+	\pyfile{graph.py}{0}{33}
+
+\section{The algorithm}
 
@@
 
-	\begin{definition}
+	\begin{definition}\label{def:mbdp}
 		\textsc{The Planar Minimal Branch Decomposition Problem}
@@
 
-	Here are some informal definitions to unpack the aforementioned properties of graphs.
+	The algorithm described in this paper solves \textsc{The Planar Minimal Branch Decomposition Problem} in polynomial time.
 
-	\begin{definition}\label{def:mbdp}
-		\textsc{The Minimal Branch Decomposition Problem}
-
-		Input: Given a simple connected graph $G$.
-
-		Output: A minimal branch decomposition of $G$.
-	\end{definition}
-
-	The algorithm described in this paper solves \textsc{The Planar Minimal Branch Decomposition Problem}, which can be computed in polynomial time.
-
-	The width of a minimal branch decomposition of $G$ is called the branch width of $G$.
-
-\section{The algorithm}
 	This section describes the algorithm given by Seymour and Thomas\cite{ST93} by identifying a set of practical problems and subproblems and how they relate.
@@
 	Considering a graph $G$, you can compute a minimal branch decomposition $(B_G, \delta_G)$ of $G$ from a minimal carving decomposition $(C_{G^\times}, \lambda_{G^\times})$ of the medial graph $G^\times$ of $G$.
-Do this by replacing the vertices in leaves of the carving decomposition with edges, using the mapping between edges and vertices from computing the medial graph;
 
-	Therefore problem \ref{def:mbdp} break down into problems \ref{problem:mcd-to-mbd}, \ref{problem:medial} and \ref{problem:mcd}.
+	Therefore problem \ref{def:mbdp} breaks down into subproblems \ref{problem:mcd-to-mbd}, \ref{problem:medial} and \ref{problem:mcd}.
+
+	\begin{problem}\label{problem:medial}
+		Given a plane graph $G$, output a medial graph $G^\times$, along with a bijectional mapping between medial nodes $V(G^\times)$ and vertex pairs $E(G)$.
+	\end{problem}
+
+	\begin{problem}\label{problem:mcd}
+		Given a plane graph $M$, output a minimal carving decomposition of $M$.
+	\end{problem}
 
@@
 
-	\begin{problem}\label{problem:medial}
-		Given a graph $G$, output a medial graph and a bijectional mapping between medial nodes and vertex pairs.
-	\end{problem}
+	With subroutines for all three problems, the implementation obtains a branch decomposition like so.
 
-	\begin{problem}\label{problem:mcd}
-		Given a plane graph $M$ and function to compute the carving width of a graph, output a minimal carving decomposition of $M$.
-	\end{problem}
+	\pyfile{branch_decomposition.py}{0}{100000}
 
-	Implementing a function to solve \ref{problem:mcd-to-mbd} is described in \ref{impl:mcd-to-mbd}.
+	Here are three sections, one for each subproblem.
 
-	To solve \ref{problem:medial} is a matter of following the definition.
+	\subsection{Medial graph}
 
-	I will refer to vertices and edges of the medial graph as "nodes" and "links" in an attempt at disambiguation.
+		Solving problem \ref{problem:medial}.
 
-	Computing a medial graph is described in \ref{impl:medial}.
+		Recall the definition \ref{def:medial}.
+
+		I assume that the input graph $G$ is a rotation system.
+Given this format, any two consecutive edges $w$ and $v$ in some face of $G$ are therefore consecutive vertices in the neighborhood of the vertex $a$ that $w$ and $v$ share.
+
+		\begin{figure}[H]
+			\centering
+			\includegraphics[width=\textwidth]{./images/medial-implementation-4-reg-argument.png}
+			\caption{Two iterations of the algorithm computing a medial rotation system}
+			\label{fig:overview}
+		\end{figure}
+
+		A medial graph is 4-regular; every node has degree 4.
+From the perspective of some medial node $v$, in some single iteration of the loop on line 14, two links are added to the neighborhood of $v$ in counterclockwise ordering, and later the two other links are added to the neighborhood of $v$ also in counterclockwise ordering.
+Therefore the medial graph $M$ is a rotation system.
+
+		\pyfile{medial_graph.py}{4}{24}
+
+	\subsection{Minimal Branch Decomposition from Minimal Carving Decomposition}
+
+		Solving problem \ref{problem:mcd-to-mbd}.
+
+		For both branch- and carving-decompositions, I have chosen a data structure of tuples of tuples or integers.
+This has a straightforward translation to the Newick tree format, a concise notation for tree structures.
+
+		To then solve the above-mentioned problem, the implementation recursively returns a copy of any tuple, but returns a tuple of integers for any integer, using the mapping from medial node to vertex pair.
+
+		\pyfile{branch_decomposition.py}{5}{19}
+
+	\subsection{Minimal Carving Decomposition}
 
@@
 
-	Doing a series of edge contractions (contraction of all edges incident to a pair of vertices) on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices can be assembled into a minimal carving decomposition of $M$.
+		By doing a series of edge contractions on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices, can be assembled into a minimal carving decomposition of $M$.
 
-	I will defer describing exactly how to assemble a minimal carving decomposition to \ref{impl:mcd}.
+		The implementation finds a nonincreasing contraction by doing a linear search over every edge.
+No consideration has yet been given to any potential clever orderings of the edges that might improve the running time.
+
+		The "contraction" function returns a new unique vertex ID, therefore by keeping which vertex is a contraction of which vertex pair in the "edges" dictionary, constructing the decomposition is then a matter of recursively expanding any vertices that were a result of a contraction into a tuple of the vertex pair that is was composed of.
+Repeating this until all only vertices of $M$ remain gives a carving decomposition in Newick-like nested tuple format.
+
+		\pyfile{carving_decomposition.py}{8}{48}
 
@@
 
-	First consider problem \ref{problem:contraction}.
+		\subsubsection{Contraction}
 
-	Informally; a contraction is merging two vertices into one and letting incident edges connect to the new vertex.
-The contraction in this paper differs a bit from conventional definitions of edge- and vertex-contractions, as other definitions might result in self-loops when contracting one of multiple parallel edges.
+			Solving problem \ref{problem:contraction}.
 
-	Computing a medial graph is described in \ref{impl:medial}.
+			As the resulting graph is later given as an argument to functions assuming a rotation system of a planar graph, the implementation needs to preserve this invariant when contracting.
 
-	Now consider problem \ref{problem:cw}.
+			Recall the definition \ref{def:contraction}.
+As this contraction is a contraction of ALL edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
+I suspect reconciling this and the rotation system could be difficult, but in this context, it is irrelevant.
 
-	The rat-catching algorithm decides $cw(M) \geq k$ with $k$ being in the positive integers.
-This is a monotonic boolean space, so you can perform a binary search to find the smallest $k$ where $cw(M) \geq k$ is true.
+			For a contraction of vertices $a$ and $b$, I have chosen to create a new vertex ID $c$, instead of reusing $a$ or $b$, as this makes the implementation, for assembling the carving decomposition, simpler.
+
+			First, update any edges incident to either $a$ or $b$.
+Then creating the neighborhood of the new vertex $c$ from the contraction of vertices $a$ and $b$, is done by firstly finding any shared edge $e$.
+In this implementation the first shared edge $e$ in the neighborhood of $a$.
+This edge has some ID $e$ and the other half-edge with ID $-e$ will therefore be in the neighborhood of $b$.
+Now  "rotating" the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0 in both lists means that a concatenation of the lists will preserve the ordering around the new vertex $c$.
+And finally, remove any edges connecting $a$ and $b$.
+
+			This is where telling apart parallel edges, which the Graph class allows, becomes very useful.
+Inferring where to stitch together the neighborhoods to preserve the ordering, solely from an adjacency list, is way less practical.
+
+			\pyfile{contraction.py}{4}{49}
+
+		\subsubsection{Carving Width}
+
+			Solving problem \ref{problem:cw}.
+
+			The rat-catching algorithm decides $cw(M) \geq k$ with $k \in \mathbb{N}^+$.
 
@@
 A round of this game is played with some noise level $k$.
-The rat-catcher wins if they can force the rat to be in some wall of the room that they are in, with noise level $k$, and the rat wins if there is a strategy whereby the rat can escape indefinitely.
+The rat-catcher wins the round if they can force the rat to be in some wall of the room that they are in, and the rat wins the round if there is a strategy whereby the rat can escape indefinitely.
 
@@
 
-	So assuming $\Delta(M) < k$ the game is played.
+			So if $\Delta(M) < k$, then the game is played to determine an outcome, otherwise the rat is said to win.
+
+			The outcomes of the rounds for $k=\{1,2,3,...\}$ is a monotonic boolean space, so you can perform a binary or linear search to find the smallest $k$ where $cw(M) \geq k$ is true.
+
+			\pyfile{carving_width.py}{0}{8}
+			\pyfile{carving_width.py}{132}{157}
+
+	\section{Does the rat win?}
+
+		We have arrived at the crux of the algorithm. Does the rat win for some integer $k$?
 
 		For some noise level and location of the rat-catcher, exactly which edges are noisy and which are quiet are definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
-
-	An edge $e$ is called quiet iff. $e$ is not noisy.
 
@@
 
+		An edge $e$ is called quiet iff. $e$ is not noisy.
+
 		\begin{definition}\label{def:face-quiet}
-		When the rat-catcher is on some room $f$, then edge $e$ is noisy iff. there is a closed walk of length scrictly less than $k$ containing $f^*$ and $e^*$ in the dual $M^*$.
+			When the rat-catcher is in some face $f$, then edge $e$ is noisy iff. there is a closed walk of length scrictly less than $k$ containing $f^*$ and $e^*$ in the dual $M^*$.
 		\end{definition}
@@
 
-\section{The implementation}
+		\textbf{impl. carving width and the rat cathing algorithm}
 
-	For the upcoming problems one needs to deal with parallel edges and be able to tell them apart, therefore the implementation uses a data structure that encapsulates an adjacency list of edges and a map from unique edge ids its vertexpair.
-
-	I have chosen to assign IDs such that if one half-edge has ID $i$ then the other half-edge has ID $-i$, therefore the absolute value $|i|$ uniquely identifies an undirected edge.
-
-	\pyfile{graph.py}{0}{100}
-
-	\subsection{Computing a minimal branch decomposition}\label{impl:mcd-to-mbd}
-
-		Solving problem \ref{problem:mcd-to-mbd}.
-
-		For both branch- and carving-decompositions, I have chosen a data structure of tuples of tuples or integers.
-This has a straightforward translation to the Newick tree format, a concise notation for tree structures.
-
-		To then solve the above-mentioned problem, the implementation recursively returns a copy of any tuple, but returns a tuple of integers for any integer, using the mapping from medial node to vertex pair.
-
-		\pyfile{branch_decomposition.py}{5}{19}
-
-	\subsection{Computing a medial graph}\label{impl:medial}
-
-		Solving problem \ref{problem:medial}.
-
-		I assume that the input graph $G$ is a rotation system of a planar graph; an adjacency list such that the neighborhoods are given in clockwise ordering according to some plane embedding of $G$.
-
-		Given this format, any two consecutive edges $e_1$ and $e_2$ in some face of $G$ are therefore consecutive vertices in the neighborhood of the vertex that $e_1$ and $e_2$ share.
-
-		The implementation adds all medial links around some vertex for each vertex in $G$.
-
-		The clockwise ordering of neighborhoods of $G$ becomes counterclockwise ordering of neighborhoods of the medial $M$.
-The medial graph of a plane graph is 4-regular \ref{}.
-From the perspective of some medial node $v$, in some single iteration of the loop on line 14, two links are added to the neighborhood of $v$ in counterclockwise ordering, and later the two other links are added to the neighborhood of $v$ also in counterclockwise ordering.
-
-		\pyfile{medial_graph.py}{4}{22}
-
-	\subsection{Computing a minimal carving decomposition}\label{impl:mcd}
-
-		Solving problem \ref{problem:mcd}.
-
-		The implementation finds a nonincreasing contraction by doing a linear search over every edge.
-No consideration has yet been given to any potential clever orderings of the edges that might improve the running time.
-
-		The sequence of contracted edges is found and reassembled into a minimal carving decomposition.
-
-		The "contraction" function returns a new unique vertex ID, therefore by saving which vertex is a contraction of which vertex pair in the "edges" dictionary, constructing the decomposition is then a matter of recursively expanding any vertices that were a result of a contraction into a tuple of the vertex pair that is was composed of.
-Repeating this until all only vertices of $M$ remain gives a carving decomposition in Newick-like nested tuple format.
-
-		\pyfile{carving_decomposition.py}{8}{48}
-
-	\subsection{Contraction}\label{impl:contraction}
-
-		Solving problem \ref{problem:contraction}.
-
-		As the resulting graph is later given as an argument to functions assuming a rotation system of a planar graph, the implementation needs to preserve this invariant when contracting.
-
-		As this contraction is a contraction of ALL edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
-I suspect reconciling this and the rotation system could be difficult, but luckily in this context, it is irrelevant.
-
-		For a contraction of vertices $a$ and $b$, I have chosen to create a new vertex ID $c$ instead of reusing $a$ or $b$ as this later makes assembling the carving decomposition easier.
-
-		First, update any edges incident to either $a$ or $b$.
-Then creating the neighborhood of the new vertex $c$ from the contraction of vertices $a$ and $b$, is done by firstly finding any shared edge $e$.
-In this implementation the first shared edge $e$ in the neighborhood of $a$.
-This edge has some ID $e$ and the other half-edge with ID $-e$ will therefore be in the neighborhood of $b$.
-Now  "rotating" the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0 in both lists means that a concatenation of the lists will preserve the ordering around the new vertex $c$.
-And finally, remove any edges connecting $a$ and $b$.
-
-		This is where telling apart parallel edges, which the Graph class allows, becomes very useful.
-Inferring where to stitch together the neighborhoods to preserve the ordering, just from a normal adjacency list, becomes a way harder problem.
-
-		\pyfile{contraction.py}{4}{49}
-
-	\subsection{Carving width and the rat cathing algorithm}\label{impl:cw}
-
-		Solving problem \ref{problem:cw}
+		\ref{problem:cw}
 
@@
 
-	\subsection{Quiet subgraph}\label{impl:quiet}
+		\textbf{impl. noisy-subgraph}
 
@@
 
-		No other path of the implementation needs the assumption that the dual is planar, therefore no clockwise or counterclockwise ordering of the neighborhoods of the adjacency list is needed.
+		No other path of the implementation needs the assumption that the dual is planar, therefore the output doesn't need to be a rotation system. This simplifies the implementation.
 
@@
 The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to, by following the edges that are just next to each other in the ordered neighborhoods.
-
-		% https://en.wikipedia.org/wiki/Rotation_system
 
@@
 \section{Appendix}
-	\pyfile{branch_decomposition.py}{0}{1000}
-	\pyfile{branch_width_brute_force.py}{0}{1000}	
-	\pyfile{branch_width.py}{0}{1000}
-	\pyfile{carving_decomposition.py}{0}{1000}
-	\pyfile{carving_width_brute_force.py}{0}{1000}
-	\pyfile{carving_width.py}{0}{1000}
-	\pyfile{contraction.py}{0}{1000}
-	\pyfile{dual_graph.py}{0}{1000}
-	\pyfile{Graph.py}{0}{1000}
-	\pyfile{medial_graph.py}{0}{1000}
-	\pyfile{parse_graph.py}{0}{1000}
-	\pyfile{parse_newick.py}{0}{1000}
+	% \pyfile{branch_decomposition.py}{0}{1000}
+	% \pyfile{branch_width_brute_force.py}{0}{1000}	
+	% \pyfile{branch_width.py}{0}{1000}
+	% \pyfile{carving_decomposition.py}{0}{1000}
+	% \pyfile{carving_width_brute_force.py}{0}{1000}
+	% \pyfile{carving_width.py}{0}{1000}
+	% \pyfile{contraction.py}{0}{1000}
+	% \pyfile{dual_graph.py}{0}{1000}
+	% \pyfile{Graph.py}{0}{1000}
+	% \pyfile{medial_graph.py}{0}{1000}
+	% \pyfile{parse_graph.py}{0}{1000}
+	% \pyfile{parse_newick.py}{0}{1000}
 

### 2024-08-14 paper/main.tex
Update paper

@@
 
-	For many subroutines the algorithm needs to deal with parallel edges and be able to tell them apart, therefore the implementation encodes a graph as an adjacency list of edges and a map from unique edge ids its vertexpair.
+	For many of the subroutines, the algorithm needs to deal with parallel edges and be able to tell them apart, therefore the implementation encodes a graph as an adjacency list of edges and a map from unique edge IDs its vertex-pair.
 
@@
 
-	Problem \ref{def:mbdp} is the overarching problem, and can be broken down into many smaller subproblems.
+	Problem \ref{def:mbdp} is the overarching problem. It can be broken down into many smaller subproblems.
 
-	Considering a graph $G$, you can compute a minimal branch decomposition $(B_G, \delta_G)$ of $G$ from a minimal carving decomposition $(C_{G^\times}, \lambda_{G^\times})$ of the medial graph $G^\times$ of $G$.
+	Considering a plane graph $G$, you can compute a minimal branch decomposition $(B_G, \delta_G)$ of $G$ from a minimal carving decomposition $(C_{G^\times}, \lambda_{G^\times})$ of the medial graph $G^\times$ of $G$.
 
-	Therefore problem \ref{def:mbdp} breaks down into subproblems \ref{problem:mcd-to-mbd}, \ref{problem:medial} and \ref{problem:mcd}.
+	Therefore problem \ref{def:mbdp} breaks down into subproblems \ref{problem:medial}, \ref{problem:mcd} and \ref{problem:mcd-to-mbd}.
 
 	\begin{problem}\label{problem:medial}
-		Given a plane graph $G$, output a medial graph $G^\times$, along with a bijectional mapping between medial nodes $V(G^\times)$ and vertex pairs $E(G)$.
+		Given a plane graph $G$, output a medial graph $G^\times$, along with a bijectional relation between medial nodes $V(G^\times)$ and vertex-pairs $E(G)$.
 	\end{problem}
@@
 
-		For both branch- and carving-decompositions, I have chosen a data structure of tuples of tuples or integers.
-This has a straightforward translation to the Newick tree format, a concise notation for tree structures.
+		For both branch- and carving-decompositions, I have chosen a data structure of tuples of either more tuples or integers.
+This has a straightforward translation to the Newick Tree Format, a concise notation for tree structures.
 
-		To then solve the above-mentioned problem, the implementation recursively returns a copy of any tuple, but returns a tuple of integers for any integer, using the mapping from medial node to vertex pair.
+		To then solve the above-mentioned problem, the implementation recursively returns either returns a copy of any tuple it encounters or returns a tuple of integers for any integer in encounters, using the relation node\_to\_vertexpair.
+
+		This effectively replaces the leaf nodes of the carving decomposition with vertex-pairs for the branch decomposition.
 
@@
 
-		The "contraction" function returns a new unique vertex ID, therefore by keeping which vertex is a contraction of which vertex pair in the "edges" dictionary, constructing the decomposition is then a matter of recursively expanding any vertices that were a result of a contraction into a tuple of the vertex pair that is was composed of.
-Repeating this until all only vertices of $M$ remain gives a carving decomposition in Newick-like nested tuple format.
+		The "contraction" function, given a pair of vertices $u,v$, returns a new unique vertex ID $w$ (instead of reusing either $u$ or $v$), therefore by keeping track of which vertex is a contraction of which vertex-pair, in say a dictionary, constructing the decomposition is then a matter of recursively looking up vertices in the dictionary.
+Repeating this until only vertices of $M$ remain gives a carving decomposition in the aforementioned Newick-like nested tuple format.
 
@@
 
-		The contraction algorithm depends on a function to compute a contraction and a function to compute the carving width. This is problems \ref{problem:contraction} and \ref{problem:cw}.
+		The contraction algorithm depends on a function to compute a contraction and a function to compute the carving width of a graph. This is problems \ref{problem:contraction} and \ref{problem:cw}.
 
 		\begin{problem}\label{problem:contraction}
-			Given a graph $M$ and a pair of vertices $\{u, v\}$, output a graph resulting from a contraction of all edges incident to $u$ and $v$ if $u$ and $v$ are neighbors.
+			Given a graph $M$ and a pair of vertices $\{u, v\}$, output the graph resulting from a contraction.
 		\end{problem}
@@
 			Recall the definition \ref{def:contraction}.
-As this contraction is a contraction of ALL edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
+As this contraction removes all edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
 I suspect reconciling this and the rotation system could be difficult, but in this context, it is irrelevant.
 
-			For a contraction of vertices $a$ and $b$, I have chosen to create a new vertex ID $c$, instead of reusing $a$ or $b$, as this makes the implementation, for assembling the carving decomposition, simpler.
+			For a contraction of vertices $a$ and $b$, I have chosen to create a new vertex ID $c$ (instead of, as the definition suggests, reusing either $a$ or $b$), as this makes the implementation, for assembling the carving decomposition, simpler.
 
-			First, update any edges incident to either $a$ or $b$.
-Then creating the neighborhood of the new vertex $c$ from the contraction of vertices $a$ and $b$, is done by firstly finding any shared edge $e$.
-In this implementation the first shared edge $e$ in the neighborhood of $a$.
-This edge has some ID $e$ and the other half-edge with ID $-e$ will therefore be in the neighborhood of $b$.
-Now  "rotating" the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0 in both lists means that a concatenation of the lists will preserve the ordering around the new vertex $c$.
+			First, let any edges incident to $a$ or $b$ be incident to $c$ instead.
+Then creating the neighborhood of the new vertex $c$ is done by firstly finding any shared edge $e$; an edge that was in the neighborhood of both $a$ and $b$.
+This edge has some ID $e$ and the other half-edge, ID $-e$, will therefore be in the neighborhood of $b$.
+Now "rotating" the lists representing the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0, in their respective lists, means that a concatenation of the lists will preserve the clockwise ordering around the new vertex $c$.
 And finally, remove any edges connecting $a$ and $b$.
 
-			This is where telling apart parallel edges, which the Graph class allows, becomes very useful.
-Inferring where to stitch together the neighborhoods to preserve the ordering, solely from an adjacency list, is way less practical.
+			This is where telling apart two edges that are parallel, by having edge IDs, becomes very useful.
+Inferring where to stitch together the neighborhoods to preserve the clockwise ordering, solely from an adjacency list, is way less practical.
 
@@
 			Solving problem \ref{problem:cw}.
+
+			The rat-catching algorithm decides whether $cw(M) \geq k$ with $k \in \mathbb{N}^+$.
+The boolean results of this algorithm for $k=\{1,2,3,...\}$ are monotonic, so you can perform a binary or linear search to find the smallest $k$ where $cw(M) \geq k$ is true.
+
+			\pyfile{carving_width.py}{0}{8}
+			\pyfile{carving_width.py}{132}{157}
+
+	\section{The Rat-Catching Algorithm}
 
@@
 		So if $\Delta(M) < k$, then the game is played to determine an outcome, otherwise the rat is said to win.
-
-			The outcomes of the rounds for $k=\{1,2,3,...\}$ is a monotonic boolean space, so you can perform a binary or linear search to find the smallest $k$ where $cw(M) \geq k$ is true.
-
-			\pyfile{carving_width.py}{0}{8}
-			\pyfile{carving_width.py}{132}{157}
-
-	\section{Does the rat win?}
 
@@
 
-		A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise level $k$ and some $e \in \mathbb{E}(M)$, is a subgraph of $M$ with $V(Q(M, k, e)) = V(M)$ and
+		A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise level $k$ and some $e \in \mathbb{E}(M)$, is a subgraph of $M$ with the vertex set $V(Q(M, k, e)) = V(M)$ and the edge set
 
@@
 		\begin{problem}\label{problem:noisy-subgraph}
-			Given a plane graph $M$ that might have parallel edges, an edge $e \in \mathbb{E}(M)$, and noise level $k \in \mathbb{N}$, output the quiet subgraph $Q(M, k, e)$.
+			Given a plane graph $M$ that might have parallel edges, an edge $e \in \mathbb{E}(M)$, and noise level $k \in \mathbb{N}^+$, output the quiet subgraph $Q(M, k, e)$.
 		\end{problem}
@@
 
-		Computing the quiet subgraph requires the dual graph.
+		With the graph of possible moves $H$, the only missing piece of the rat-catching algorithm is how to determine the outcome.
 
-		With the graph $H$, the only missing piece of the rat-catching algorithm is how to determine the outcome.
-
-		You can mark states of the graph $H$ that are losing states, and then repeatedly mark any state that leads to a losing state, until either every state is marked or no more states can be marked.
+		You can mark states/vertices of the graph $H$ that are losing states, and then repeatedly mark any state that leads to a losing state, until either every state is marked or no more states can be marked.
 If every state is marked then the rat-catcher wins, otherwise the rat wins.
 
-		\textbf{impl. carving width and the rat cathing algorithm}
-
-		\ref{problem:cw}
+		Solving problem \ref{problem:cw}.
 
@@
 
-		Losing states - the tuples $(f,v) \in S$ where $v \in f$ - are marked as losing.
+		Losing states (the tuples $(f,v) \in S$ where $v \in f$) are marked as losing.
 
@@
 
-		\textbf{impl. noisy-subgraph}
+		\subsubsection{Quiet subgraph}
 
@@
 
-		Using definition \ref{def:edge-quiet}: When the rat-catcher is on some edge $e_1$, then edge $e_2$ is noisy iff.
-there is a closed walk of length scrictly less than $k$ containing $e_1^*$ and $e_2^*$ in the dual $M^*$.
+			Recall the definition \ref{def:edge-quiet}.
 
-		Let $s_1$ and $t_1$ be the vertex pair for the link $e_1^*$ and let $s_2$ and $t_2$ be the vertex pair for the link $e_2^*$.
+			Let $s_1$ and $t_1$ be the vertex-pair for the link $e_1^*$ and let $s_2$ and $t_2$ be the vertex-pair for the link $e_2^*$.
 
 			\begin{claim}
-			The shortest closed walk that includes both $e_1^*$ and $e_2^*$ has the same length as either $$d(s_1, s_2) + d(t_1, t_2) + 2$$ or $$d(s_1, t_1) + d(s_2, t_2) + 2$$.
+				The shortest closed walk that includes both $e_1^*$ and $e_2^*$ is the minimum of either $$d(s_1, s_2) + d(t_1, t_2) + 2$$ or $$d(s_1, t_1) + d(s_2, t_2) + 2$$.
 Where $d(u, v)$ is the length of the shortest $u,v$-path.
@@
 
-	\subsection{Dual graph}\label{impl:dual}
+		\subsubsection{Dual graph}\label{impl:dual}
 
@@
 
-		The dual has a vertex for each face of the input graph.
-The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to, by following the edges that are just next to each other in the ordered neighborhoods.
+			The dual has a vertex for each face of the input graph. The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to.
 
-		The next halfedge $e_{i+1}$ after the current halfedge $e_i = \{u, v\}$ is the edge just before $-e_i$ in the ordered neighborhood around $v$.
+			The next halfedge $e_{i+1}$ after the current halfedge $e_i = \{u, v\}$ is the edge just before $-e_i$ in the neighborhood list for vertex $v$.
 
@@
 			\pyfile{dual_graph.py}{4}{50}
+
+\section{Results}
+
+	The algorithm has been implemented in Python 3.12.2. The source code is available at \url{https://github.com/hojelse/thesis}.
+
+	The implementation has been tested on a set of graphs, and the results are as expected. \todo{more}
 

### 2024-08-19 code/Graph.py
Update code

@@ -27,8 +27,8 @@ class Graph:
 		return list(self.edge_to_vertexpair.keys())
 
 	def N(self, v: int) -> list[int]:
-		return [self.edge_to_vertexpair[e][1] for e in self.adj_edges[v]]
-
+		return [self.edge_to_vertexpair[e][1] if self.edge_to_vertexpair[e][0] == v else self.edge_to_vertexpair[e][0] for e in self.adj_edges[v]]
+	
 	def adj(self) -> dict[int, list[int]]:
 		return dict([(x, self.N(x)) for x in self.adj_edges.keys()])
 	
@@ -41,3 +41,38 @@ class Graph:
 		H.adj_edges = adj_edges_deepcopy
 		H.edge_to_vertexpair = edge_to_vertexpair_deepcopy
 		return H
+
+	def from_lmg(self, input_string: str):
+		lines = input_string.strip().split('\n')
+		N, M = map(int, lines[0].split())
+
+		for line in lines[1:N+1]:
+			x, *ys = map(int, line.split())
+			self.adj_edges[x] = ys
+
+		for line in lines[N+1:]:
+			e, u, v = map(int, line.split())
+			self.edge_to_vertexpair[e] = (u, v)
+
+	def to_lmg(self) -> str:
+		s = str(len(self.adj_edges)) + " " + str(len(self.edge_to_vertexpair.keys())) + "\n"
+		for x, ys in self.adj_edges.items():
+			s += str(x) + " " + " ".join(map(lambda y: str(y), ys)) + "\n"
+		for e, (u, v) in self.edge_to_vertexpair.items():
+			s += str(e) + " " + str(u) + " " + str(v) + "\n"
+		return s
+
+	def __str__(self):
+		return self.to_lmg()
+	
+def read_lmg_from_stdin() -> str:
+	N,M = map(int, input().split())
+	stdin_multiline_string = f"{N}, {M}\n"
+	for _ in range(N+M):
+		stdin_multiline_string += input() + "\n"
+	return stdin_multiline_string
+
+if __name__ == "__main__":
+	G = Graph()
+	G.from_lmg(read_lmg_from_stdin())
+	print(G)

### 2024-08-19 code/branch_decomposition.py

@@ -1,12 +1,14 @@
+import log
 from parse_graph import parse_text_to_adj
-from medial_graph import medial_graph
+from medial_graph import medial_graph, medial_graph_2
 from carving_decomposition import carving_decomposition
 
 # Construct a branch decomposition of a graph
 def branch_decomposition(G_adj: dict[int, list[int]]):
 	# Contruct the carving decomposition of the medial graph
-	M, node_to_vertexpair, vertexpair_to_node = medial_graph(G_adj)
-	cd = carving_decomposition(M)
+	Gx, node_to_vertexpair = medial_graph(G_adj)
+
+	cd = carving_decomposition(Gx.copy())
 
 	# Convert the carving decomposition of M to a branch decomposition of G
 	def decomp(t):
@@ -15,6 +17,8 @@ def branch_decomposition(G_adj: dict[int, list[int]]):
 		return tuple([decomp(a) for a in t])
 	
 	bd = decomp(cd)
+
+	log.add("Branch decomposition: " + str(bd))
 	return bd
 
 if __name__ == "__main__":

### 2024-08-19 code/branch_width.py

@@ -1,3 +1,4 @@
+import log
 from branch_decomposition import branch_decomposition
 from parse_graph import parse_text_to_adj, adj_to_text
 
@@ -42,11 +43,14 @@ def branch_width_of_branch_decomposition(bd):
 			middle_set = len(a.intersection(b))
 			width = max(width, middle_set)
 	
+	log.add("Branch width of branch decomposition: " + str(width))
 	return width
 
 def branch_width(adj: dict[int, list[int]]):
 	bd = branch_decomposition(adj)
-	return branch_width_of_branch_decomposition(bd)
+	bw = branch_width_of_branch_decomposition(bd)
+	log.add("Branch width: " + str(bw))
+	return bw
 
 if __name__ == "__main__":
 	adj = parse_text_to_adj()

### 2024-08-19 code/carving_decomposition.py

@@ -1,3 +1,4 @@
+import log
 from carving_width import carving_width
 from contraction import contraction
 from parse_graph import parse_text_to_adj, adj_to_text
@@ -5,11 +6,12 @@ from Graph import Graph
 
 # Find a contraction that does not increase the carving width
 def nonincreasing_cw_contraction(G: Graph, cw1: int) -> tuple:
-	for es in G.E():
-		u, v = G.edge_to_vertexpair[es]
+	for e in G.E():
+		u, v = G.edge_to_vertexpair[e]
 		G2, w = contraction(G, u, v)
 		cw2 = carving_width(G2)
 		if cw2 <= cw1:
+			log.add(f"Graph after contracting edge {e}: \n{str(G2)}")
 			return G2, (u, v), cw2, w
 	return None, None, None, None
 
@@ -43,6 +45,8 @@ def carving_decomposition(G: Graph) -> tuple:
 
 	a,b,c = G2.V()
 	cd = (decomp(a), decomp(b), decomp(c))
+
+	log.add("Carving decomposition: " + str(cd))
 	return cd
 
 if __name__ == "__main__":

### 2024-08-19 code/carving_width.py

@@ -134,6 +134,7 @@ def carving_width(G: Graph) -> int:
 		r = 1
 		while True:
 			if rat_wins(r):
+				
 				l = r
 				r *= 2
 			else:

### 2024-08-19 code/log.py

@@ -1,0 +1,7 @@
+def init():
+	global log
+	log = []
+
+def add(o):
+	log.append(o)
+	print(o)

### 2024-08-19 code/main.py

@@ -1,0 +1,23 @@
+import log
+import branch_width
+import sys
+import Graph
+
+graph_file_name = sys.argv[1]
+
+log.init()
+log.add("Reading graph from file " + graph_file_name)
+with open(graph_file_name, 'r') as file:
+	graph_str = file.read()
+
+	G = Graph.Graph()
+	if graph_file_name.endswith(".m.in"):
+		G.from_lmg(graph_str)
+	else:
+		print("Unsupported file extension")
+		exit()
+
+	log.add("Graph: " + str(G))
+	log.add("Graph adjacency list: " + str(G.adj()))
+
+	bw = branch_width.branch_width(G.adj())

### 2024-08-19 code/medial_graph.py

@@ -1,8 +1,11 @@
+import log
 from Graph import Graph
-from parse_graph import adj_to_text, parse_text_to_adj
+from parse_graph import adj_to_text, adj_to_text_2, parse_text_to_adj
 
-# assume planar graph
+# assume G is simple
+# assume G is planar
 # assume G_adj is a rotation system
+# guarantee M is a rotation system
 def medial_graph(G_adj: dict[int, list[int]]) -> Graph:
 	vertexpairs = set([tuple(sorted((i, j))) for i in G_adj for j in G_adj[i]])
 
@@ -19,11 +22,67 @@ def medial_graph(G_adj: dict[int, list[int]]) -> Graph:
 
 	M = Graph()
 	M.from_adj(medial)
-	return M, node_to_vertexpair, vertexpair_to_node
+	# log.add("Medial graph: " + str(M))
+	# log.add("Node to vertexpair: " + str(node_to_vertexpair))
+	return M, node_to_vertexpair
+
+# assume G is simple
+# assume G is planar
+# assume G_adj is a rotation system
+# guarantee an edge of G have the same id as its corresponding medial node
+# guarantee M is a rotation system
+def medial_graph_2(G_adj: dict[int, list[int]]):
+	G = Graph()
+	G.from_adj(G_adj)
+
+	M = Graph()
+	M.adj_edges = dict((abs(n), []) for n in G.edge_to_vertexpair.keys())
+
+	# print(M.adj_edges)
+
+	edges = [e for e in G.E()]
+	faces = []
+	while len(edges) > 0:
+		e = edges.pop()
+		next_e = e
+		face = [e]
+		while True:
+			u,v = G.edge_to_vertexpair[next_e]
+			idx = G.adj_edges[v].index(-next_e)
+			next_e = G.adj_edges[v][(idx-1)%len(G.adj_edges[v])]
+			if (next_e == e):
+				break
+			edges.remove(next_e)
+			face.append(next_e)
+		faces.append(face)
+
+	next_linkid = 1
+	for face in faces:
+		for i in range(len(face)):
+			n1 = abs(face[i])
+			n2 = abs(face[(i-1)%len(face)])
+			M.adj_edges[n1].append(next_linkid)
+			M.adj_edges[n2].append(-next_linkid)
+			M.edge_to_vertexpair[next_linkid] = (n1, n2)
[…]

### 2024-08-19 paper/images/carving.tex
Update paper

--- 2024-08-12 paper/images/branch.tex
@@
 		\draw[line width=0.5mm] (a3) -- node[above] {5} (a5);
+
+		\node[opacity=0] at (-1, 3.5) (start) {};
+		\node[opacity=0] at (-1, 0.5) (end) {};
+		\draw[dotted, thick] (start) .. controls +(2,0) and +(2,0) .. (end);
 	\end{scope}
@@
 	\begin{scope}[xshift=6cm, yshift=0cm]
-		\node (G) at (1, 6) {Branch Decomposition of $G$};
+		\node (G) at (1, 6) {Carving Decomposition of $G$};
+		\node (G) at (1, 5.5) {of width 3};
 
-		\node[circle, draw, color=black, fill=white] (b1) at (1, 4) {(1,2)};
-		\node[circle, draw, color=black, fill=white] (b3) at (0, 3) {(2,4)};
-		\node[circle, draw, color=black, fill=white] (i1) at (1, 3) {};
-		\node[circle, draw, color=black, fill=white] (i2) at (1.5, 2) {};
-		\node[circle, draw, color=black, fill=white] (i3) at (1, 1) {};
-		\node[circle, draw, color=black, fill=white] (b2) at (2.5, 2) {(2,3)};
-		\node[circle, draw, color=black, fill=white] (b4) at (0, 1) {(4,5)};
-		\node[circle, draw, color=black, fill=white] (b5) at (1, 0) {(3,5)};
+		\node[circle, draw, color=black, fill=white] (b1) at (1, 4) {1};
+		\node[circle, draw, color=black, fill=white] (b3) at (0, 3) {2};
+		\node[circle, draw, color=black, fill=black] (i1) at (1, 3) {};
+		\node[circle, draw, color=black, fill=black] (i2) at (1.5, 2) {};
+		\node[circle, draw, color=black, fill=black] (i3) at (1, 1) {};
+		\node[circle, draw, color=black, fill=white] (b2) at (2.5, 2) {3};
+		\node[circle, draw, color=black, fill=white] (b4) at (0, 1) {4};
+		\node[circle, draw, color=black, fill=white] (b5) at (1, 0) {5};
 
@@
 		\draw[line width=0.5mm, color=black] (i3) -- (b5);
+
+		\node (G) at (-0.5, 4.25) {\{1,2,3\}};
+
+		\node[opacity=0] at (0,4) (start) {};
+		\node[opacity=0] at (0,2) (end) {};
+		\draw[dotted, thick] (start) to[bend left=60] (end);
+	\end{scope}
+
+	\begin{scope}[xshift=12cm, yshift=0cm]
+		\node (G) at (1, 6) {Minimum Carving Decomposition of $G$};
+		\node (G) at (1, 5.5) { of width 3};
+
+		\node[circle, draw, color=black, fill=white] (b1) at (1, 4) {1};
+		\node[circle, draw, color=black, fill=white] (b3) at (0, 3) {2};
+		\node[circle, draw, color=black, fill=black] (i1) at (1, 3) {};
+		\node[circle, draw, color=black, fill=black] (i2) at (1.5, 2) {};
+		\node[circle, draw, color=black, fill=black] (i3) at (1, 1) {};
+		\node[circle, draw, color=black, fill=white] (b2) at (2.5, 2) {3};
+		\node[circle, draw, color=black, fill=white] (b4) at (0, 1) {4};
+		\node[circle, draw, color=black, fill=white] (b5) at (1, 0) {5};
+
+		% Draw the edges between nodes with labels
+		\draw[line width=0.5mm, color=black] (b1) -- (i1);
+		\draw[line width=0.5mm, color=black] (b3) -- (i1);
+		\draw[line width=0.5mm, color=black] (i1) -- (i2);
+		\draw[line width=0.5mm, color=black] (b2) -- (i2);
+		\draw[line width=0.5mm, color=black] (i2) -- (i3);
+		\draw[line width=0.5mm, color=black] (i3) -- (b4);
+		\draw[line width=0.5mm, color=black] (i3) -- (b5);
+
+		\node (G) at (-0.5, 4.25) {\{1,2,3\}};
+
+		\node[opacity=0] at (0,4) (start) {};
+		\node[opacity=0] at (0,2) (end) {};
+		\draw[dotted, thick] (start) to[bend left=60] (end);
 	\end{scope}

### 2024-08-19 paper/main.tex
Fix code snippets line ranges
Update paper

@@
 
-\title{Implementation of a minimal branch-decomposition algorithm for planar graphs.}
+\title{Implementation of a minimum branch-decomposition and branch-width algorithm for planar graphs.}
 \author{Kristoffer Højelse}
@@
 \section{Introduction}
-	Some graph optimization problems can be solved efficiently for graphs of small branchwidth.\cite{CNP+11}
+
+	\todo{change graph to multigraph}
+
+	In complexity theory, an NP-hard problem has no known algorithm with all 3 of these properties: (1) runs in polynomial time, (2) gives an exact solution, and (3) runs on all instances of the problem.
+
+	You can find brute-force algorithms for NP-hard problems that both give an exact solution and run on all instances of the problem, but they run in exponential time.
+
+	You can find approximation algorithms for NP-hard problems that both run in polynomial time and run on all instances, but they give an approximate solution.
+
+	Finally, you can find parameterized algorithms for NP-hard problems that both run in polynomial time and give an exact solution, but they only run on instances of the problem where some parameter is small.
+Treewidth is a common such parameter.
+
+	Treewidth is a measure of how tree-like a graph is, a value of 1 is exactly a tree and larger values are less tree-like.
+For some NP-hard problems, if a graph is sufficiently tree-like, then a parameterized algorithm can find an exact solution in polynomial time.
+For instance, the best-known running time for the \textsc{The Maximum Independent Set Problem} is exponential in $n$, particularly $O(1.1996^n \cdot n^{O(1)})$\cite{XN17}, but can be solved in $O(2^k \cdot k \cdot n)$ time for graphs of treewidth at most $k$, which is linear in $n$ and exponential in $k$.
+
+	Branch width and carving width are related to treewidth, and some NP-hard problems can be solved efficiently for graphs of small branchwidth.\cite{CNP+11}
+
+	Tree-, branch- and carving-width are all related. The treewidth of a graph is at least bw-1 and at most floor(3/2bw)-1.\cite{RS91}
+
+	The carving width is at least half the branch width and is at most the degree times the branch width.\cite{ED18}
+
+	You might have a parameterized algorithm with tree- or branch-width as a parameter, but to determine if a graph has a small tree- or branch-width, you might need to compute the tree- or branch-decomposition first.
+There is no known optimal tree decomposition/treewidth algorithm even for planar graphs.
+
+	A planar graph is, informally, a graph that you can draw on a plane surface without any edges crossing.
+There might be multiple such drawings of a planar graph.
+A particular such drawing is called a plane graph.
+
+	The algorithm by Seymour and Thomas\cite{ST93} computes a minimum branch-decomposition of a planar graph in $O(n^4)$ time.
+
+	You can obtain an approximate optimum tree decomposition from the branch decomposition. So computing an approximate treewidth for planar graphs is possible in polynomial time.
+
+	The concepts of medial graph and dual graph are used in the algorithm. Both are planar graphs, that are constructed from a plane graph.
+
+	The dual graph has vertices for faces and edges connecting the faces that share an edge.
+You might imagine a map of countries, with edges for borders, and a vertex for each country, plus a vertex for the surrounding sea.
+You could for instance query such a network for which countries are neighbors, or which countries are landlocked.
+
+	The medial graph also considers the faces but has vertices for edges, that are connected if the original edges were consecutive in some face. I have found no good metaphors for this.
 
@@
 
-	\todo{which? counting Hamiltonian cycles of planar cubic graphs}
+	\todo{Read params book about tree width}
 
@@
 
+	This section defines the terms and concepts used in the algorithm.
+
+	The algorithm deals with graphs that might have parallel edges.
+Distinguishing between two parallel edges is crucial.
+Therefore, the implementation gives labels to both vertices and edges, by encoding a graph as an adjacency list of edge IDs and a map from unique edge IDs to its vertex-pair.
+
 	A \textit{graph} $G$ consists of a vertex set $V(G)$, and an edge set $\mathbb{E}(G)$ and a function $\phi_G$, where $V(G) \subset \mathbb{N}^+$ and where $\mathbb{E}(G) \subset \mathbb{N}^+$ and where $\phi_G \st \mathbb{E}(G) \to \{\{u,v\} \st u,v \in V(G)\}$.
 
-	\textbf{Note.} Other authors might instead call this definition an undirected labelled pseudograph, with edges having own identity.
+	\textbf{Note.} Normally \textit{graph} refers to a simple graph, usually also without labels, which are normally encoded by an adjacency list or matrix.
+In this paper, a graph is a multigraph with labels.
 
 	\textbf{Note.} Regarding notation, $V$ and $\mathbb{E}$ are operations on graphs returning the vertex set and edge set respectively.
+
+	Let $E(G)$ return a multiset of all vertex-pairs of $G$; in other words, $E(G) = \{\phi_G(e) \st e \in \mathbb{E}(G)\}$.
 
@@
 
-	Let $E(G)$ return a multiset of all vertex-pairs of $G$; in other words, $E(G) = \{\phi_G(e) \st e \in \mathbb{E}(G)\}$.
-
 	A \textit{self-loop}, is an edge $e$ where $\phi_G(e) = \{u,v\}$ and $u = v$.
@@
 
-	A \textit{multi-graph} is a graph that is loop-less.
-
-	\begin{center}
-		\input{images/multigraph.tex}
-	\end{center}
-
-	A multi-graph $G$ is \textit{simple}, if it has no parallel edges; in other words, if elements of $E(G)$ are pair-wise distinct.
+	A graph $G$ is \textit{simple}, if it has no parallel edges; in other words, if all elements of $E(G)$ are pair-wise distinct.
 
@@
 
-		A contraction is a function that given a multi-graph $G$ and pair of distinct vertices $u,v \in V(G)$ such that $\{u,v\} \in E(G)$, then for all edges $e \in \mathbb{E}(G)$ if $\phi_G(e) = \{u,v\}$ then removes $e$ else if $\phi_G(e) = \{v,w\}$ then $\phi_G(e) = \{u,w\}$, and finally returns the resulting graph.
+		A contraction is a function that given a multi-graph (with no self-loops) $G$ and pair of distinct vertices $u,v$ of some edge $\{u,v\} \in E(G)$, removes $e$ if $\phi_G(e) = \{u,v\}$ and makes any edge that is adjacent to $v$, $\phi_G(e) = \{w,v\}$, adjacent to $u$ instead, $\phi_G(e) = \{w,u\}$, and finally returns the resulting graph.
 	\end{definition}
+
+	\begin{corollary}
+		The resulting graph of a contraction will not have any self-loops.
+	\end{corollary}
 
@@
 
-	An \textit{internal vertex} $v$ of a tree $T$, is a vertex $v \in V(T) - L(T)$ that is not a leaf. An internal vertex therefore has degree at least 2.
+	An \textit{internal vertex} $v$ of a tree $T$, is a vertex $v \in V(T) - L(T)$ that is not a leaf. An internal vertex therefore has at least degree 2.
 
@@
 
+	Removing any edge $e \in \mathbb{E}(B_G)$ partitions $B_G$ into 2 trees $P_e$ and $Q_e$.
+The set $\cup\{\phi_G(e) \st e \in L(P_e)\} \cap \cup\{\phi_G(e) \st e \in L(Q_e)\}$ is called a \textit{middle set} of $B_G$ given $e$, denoted $Z(B_G,e)$.
+The maximal cardinality of any middle set of $B_G$ given any $e \in \mathbb{E}(B_G)$ is the width of $B_G$; in other words, the width of $B_G$ is $\max\{ |Z(B, e)| \st e \in \mathbb{E}(B) \}$.
+
+	\todo{middle set definition is convoluted}
+
+	A \textit{Minimum Branch Decomposition} of $G$ is any branch decomposition of $G$ of minimum width among all branch decompositions of $G$, as there might exist many branch decompositions of a graph $G$.
+
 	\begin{center}
@@
 
-	Removing any edge $e \in \mathbb{E}(B_G)$ partitions $B_G$ into 2 trees $P_e$ and $Q_e$.
-The set $L(P_e) \cap L(Q_e)$ is called a \textit{middle set} of $B_G$ given $e$, denoted $Z(B_G,e)$.
-The maximal cardinality of any middle set of $B_G$ given any $e$ of $B_G$ is the width of $B_G$; in other words, the width of $B_G$ is $\max\{ |Z(B, e)| \st e \in \mathbb{E}(B) \}$.
-
-	There might exist many branch decompositions of a graph $G$.
-
-	A \textit{Minimal Branch Decomposition} of $G$ is any branch decomposition of $G$ of minimal width among all branch decompositions of $G$.
-
 	A \textit{Carving Decomposition} $(C_G, \lambda_G)$ of a simple graph $G$ consists of firstly, an unrooted binary tree $C_G$ and secondly a bijection $\lambda_G$ between $V(G)$ and $L(C_G)$.
@@
 	Removing any edge $e \in \mathbb{E}(C_G)$ partitions $C_G$ into 2 trees $P_e$ and $Q_e$.
-The set $L(P_e) \cap L(Q_e)$ is called a \textit{middle set} of $C_G$ given $e$, denoted $Z(C_G,e)$.
-The maximal cardinality of any middle set of $C_G$ given any $e$ of $C_G$ is the width of $C_G$; in other words, the width of $C_G$ is $\max\{ |Z(C_G, e)| \st e \in \mathbb{E}(C_G) \}$.
+The set of edges $e \in \mathbb{E}(G)$ that cross $L(P_e)$ and $L(Q_e)$ is called a \textit{crossing set} of $C_G$ given an $e \in \mathbb{E}(C_G)$, denoted $Y(C_G,e)$.
+The maximum cardinality of any crossing set of $C_G$ given any $e \in \mathbb{E}(C_G)$ is the width of $C_G$; in other words, the width of $C_G$ is $\max\{ |Y(C, e)| \st e \in \mathbb{E}(C) \}$.
 
-	There might exist many carving decompositions of a graph $G$.
+	A \textit{Minimum Carving Decomposition} of $G$ is any carving decomposition of $G$ of minimum width among all carving decompositions of $G$, as there might exist many carving decompositions of a graph $G$.
 
-	A \textit{Minimal Carving Decomposition} of $G$ is any carving decomposition of $G$ of minimal width among all carving decompositions of $G$.
+	\begin{center}
+		\input{images/carving.tex}
+	\end{center}
+
+	\todo{find example with non-mimimum carving decomposition}
 
@@
 
-	Given a simple connected planar graph $G$, the algorithm computes a minimal branch decomposition of $G$.
+	Given a simple connected planar graph $G$, the algorithm computes a minimum branch decomposition of $G$.
 
-	To compute a minimal branch decomposition of $G$, the algorithm first computes the medial graph $G^\times$ of $G$, secondly computes a minimal carving decomposition of $G^\times$, and finally computes a minimal branch decomposition of $G$ from the minimal carving decomposition of $G^\times$.
+	To compute a minimum branch decomposition of $G$, the algorithm first computes the medial graph $G^\times$ of $G$, secondly computes a minimum carving decomposition of $G^\times$, and finally computes a minimum branch decomposition of $G$ from the minimum carving decomposition of $G^\times$.
 
-	To compute the minimal carving decomposition of $G^\times$, the algorithm uses a contraction algorithm that finds a series of contractions that does not increase the carving width and then assembles the series of contractions into a minimal carving decomposition of $G^\times$.
+	To compute the minimum carving decomposition of $G^\times$, the algorithm uses a contraction algorithm that finds a series of contractions that does not increase the carving width and then assembles the series of contractions into a minimum carving decomposition of $G^\times$.
 
@@
 
-	The main computational problem of this paper is \textsc{The Planar Minimal Branch Decomposition Problem}.
+	The main computational problem of this paper is \textsc{The Planar Minimum Branch Decomposition Problem}.
 
 	\begin{definition}\label{def:mbdp}
-		\textsc{The Planar Minimal Branch Decomposition Problem}
+		\textsc{The Planar Minimum Branch Decomposition Problem}
 
@@
 
-		Output: A minimal branch decomposition of $G$.
+		Output: A minimum branch decomposition of $G$.
 	\end{definition}
 
-	The algorithm described in this paper solves \textsc{The Planar Minimal Branch Decomposition Problem} in polynomial time.
+	The algorithm described in this paper solves \textsc{The Planar Minimum Branch Decomposition Problem} in polynomial time.
 
@@
 
-	Considering a plane graph $G$, you can compute a minimal branch decomposition $(B_G, \delta_G)$ of $G$ from a minimal carving decomposition $(C_{G^\times}, \lambda_{G^\times})$ of the medial graph $G^\times$ of $G$.
+	Considering a plane graph $G$, you can compute a minimum branch decomposition $(B_G, \delta_G)$ of $G$ from a minimum carving decomposition $(C_{G^\times}, \lambda_{G^\times})$ of the medial graph $G^\times$ of $G$.
 
@@
 	\begin{problem}\label{problem:medial}
-		Given a plane graph $G$, output a medial graph $G^\times$, along with a bijectional relation between medial nodes $V(G^\times)$ and vertex-pairs $E(G)$.
+		Given a plane graph $G$, output a medial graph $G^\times$, along with a bijectional relation between medial nodes $V(G^\times)$ and edges $\mathbb{E}(G)$.
 	\end{problem}
@@
 	\begin{problem}\label{problem:mcd}
-		Given a plane graph $M$, output a minimal carving decomposition of $M$.
+		Given a plane graph $M$, output a minimum carving decomposition of $M$.
 	\end{problem}
@@
 	\begin{problem}\label{problem:mcd-to-mbd}
-		Given a minimal carving decomposition of a medial graph of $G$, output a minimal branch decomposition of $G$.
+		Given a minimum carving decomposition of a medial graph of $G$, output a minimum branch decomposition of $G$.
 	\end{problem}
@@
 
-	\pyfile{branch_decomposition.py}{0}{100000}
+	\pyfile{branch_decomposition.py}{6}{22}
 
@@
 
-		\pyfile{medial_graph.py}{4}{24}
+		\pyfile{medial_graph.py}{5}{27}
 
-	\subsection{Minimal Branch Decomposition from Minimal Carving Decomposition}
+	\subsection{Minimum Branch Decomposition from Minimum Carving Decomposition}
 
@@
 
-	\subsection{Minimal Carving Decomposition}
+	\subsection{Minimum Carving Decomposition}
 
@@
 
-		By doing a series of edge contractions on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices, can be assembled into a minimal carving decomposition of $M$.
+		By doing a series of edge contractions on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices, can be assembled into a minimum carving decomposition of $M$.
 
 		The implementation finds a nonincreasing contraction by doing a linear search over every edge.
+
 		No consideration has yet been given to any potential clever orderings of the edges that might improve the running time.
@@
 
-		\pyfile{carving_decomposition.py}{8}{48}
+		\pyfile{carving_decomposition.py}{7}{50}
 
@@
 
-			\pyfile{carving_width.py}{0}{8}
-			\pyfile{carving_width.py}{132}{157}
+			\pyfile{carving_width.py}{7}{8}
+			\pyfile{carving_width.py}{132}{158}
 
@@
 
-		\pyfile{carving_width.py}{7}{9}
 		\pyfile{carving_width.py}{78}{130}
 
-		\subsubsection{Quiet subgraph}
+		\subsection{Quiet subgraph}
 
@@
 
-		\subsubsection{Dual graph}\label{impl:dual}
+		\subsection{Dual graph}\label{impl:dual}
 
@@
 
-			The dual has a vertex for each face of the input graph. The faces are found by selecting an unmarked half-edge, and then marking all the edges of the face it belongs to.
+			The dual has a vertex for each face of the input graph.
+The faces are found by selecting an unmarked half-edge, then marking all the edges of the face it belongs to, and repeating this until all half-edges are marked.
 
@@
 
-			\pyfile{dual_graph.py}{21}{22}
-
-			\pyfile{dual_graph.py}{4}{50}
+			\pyfile{dual_graph.py}{4}{56}
 
@@
 	The implementation has been tested on a set of graphs, and the results are as expected. \todo{more}
+
+	\todo{Agreement with Andreas' implementation on 6-100 vertices graphs, 10 different graphs}
+
+	\todo{note Tested on cubic planar graph}
+
+	\todo{tested by permuting vertex labels should not change the branch width}
+
+	\todo{test that bw follows theoretical bound}
+
+	Fomin and Thilikos 2006, New upper bound obn the decomposability pf planar graphs, bw <= 4.5 * sqrt(N) on planar
+
+	Semour Thomas 1994, Call routing and ratcatcher, bw(G) = bw(dual(G)) assuming G is bridgeless planar graphs
+
+	then the dual has half the vertices of the primal
+
+	4.5 * sqrt(N/2)
+
+	<=> 3.18199 * sqrt(n)
+
+	Test bw <= 3.18199 * sqrt(N) on cubic planar graphs
+
+	\todo{graphing theoretical bound}
 

### 2024-09-01 paper/main.tex
Update notes
Lots of changes

@@
 \title{Implementation of a minimum branch-decomposition and branch-width algorithm for planar graphs.}
-\author{Kristoffer Højelse}
-\date{February 2024}
+\author{Kristoffer Højelse (krbh)}
+
+\date{2 Sep 2024}
 
@@
 
+\begin{center}
+	\includegraphics[width=0.9\textwidth]{images/dalle-pied-piper.png}
+\end{center}
+\footnote{Illustration by Dall-E via ChatGPT (GPT-4o) with the prompt "Generate an image of pied piper the rat-catcher along with a computer science graph".}
+
+\newpage
+
 % west https://dwest.web.illinois.edu/grammar.html
+
 % The abstract states the results as fully as possible in a brief presentation.
@@
 
-	\todo{change graph to multigraph}
-
-	In complexity theory, an NP-hard problem has no known algorithm with all 3 of these properties: (1) runs in polynomial time, (2) gives an exact solution, and (3) runs on all instances of the problem.
+	In algorithmic complexity theory, an NP-hard problem has no known algorithm with all 3 of these properties: (1) runs in polynomial time, (2) gives an exact solution, and (3) runs on all instances of the problem.
 
@@
 
-	Treewidth is a measure of how tree-like a graph is, a value of 1 is exactly a tree and larger values are less tree-like.
-For some NP-hard problems, if a graph is sufficiently tree-like, then a parameterized algorithm can find an exact solution in polynomial time.
-For instance, the best-known running time for the \textsc{The Maximum Independent Set Problem} is exponential in $n$, particularly $O(1.1996^n \cdot n^{O(1)})$\cite{XN17}, but can be solved in $O(2^k \cdot k \cdot n)$ time for graphs of treewidth at most $k$, which is linear in $n$ and exponential in $k$.
+	Treewidth is a measure of how tree-like a graph is. A graph with a treewidth of 1 is a tree. Larger values are less tree-like.
 
-	Branch width and carving width are related to treewidth, and some NP-hard problems can be solved efficiently for graphs of small branchwidth.\cite{CNP+11}
+	For some NP-hard problems, if a graph has a small enough treewidth, then a parameterized algorithm can find an exact solution in polynomial time.
 
-	Tree-, branch- and carving-width are all related. The treewidth of a graph is at least bw-1 and at most floor(3/2bw)-1.\cite{RS91}
+	Branch width and carving width are related to treewidth, and some NP-hard problems can be solved efficiently for graphs of small branch width.
 
-	The carving width is at least half the branch width and is at most the degree times the branch width.\cite{ED18}
+	The treewidth of a graph is at least $b-1$ and at most $\lfloor \frac{3}{2} b\rfloor-1$, where $b$ is the branch width.\cite{RS91}
 
-	You might have a parameterized algorithm with tree- or branch-width as a parameter, but to determine if a graph has a small tree- or branch-width, you might need to compute the tree- or branch-decomposition first.
-There is no known optimal tree decomposition/treewidth algorithm even for planar graphs.
+	The carving width is at least $\frac{b}{2}$ and is at most $\Delta(G) \cdot b$, where $b$ is the branch width.\cite{ED18}
 
-	A planar graph is, informally, a graph that you can draw on a plane surface without any edges crossing.
+	A decomposition is generally a derivative structure that encapsulates some aspect of a graph.
+Tree-, branch- and carving-width all have related decompositions.
+These decompositions can be used in conjunction with dynamic programming to design efficient parameterized algorithms.
+For instance, Pino\cite{Pin16} applies branch decompositions on some NP-hard connectivity problems.
+
+	It is NP-complete to determine whether a graph $G$ has a branch-decomposition of width at most $k$, when $G$ and $k$ are both considered as inputs to the problem.\cite{ST93}
+
+	Informally, a planar graph is a graph that you can draw on a plane surface without any edges crossing.
 There might be multiple such drawings of a planar graph.
 A particular such drawing is called a plane graph.
+A plane graph can be said to divide the plane into regions called \textit{faces}.
 
@@
 
-	You can obtain an approximate optimum tree decomposition from the branch decomposition. So computing an approximate treewidth for planar graphs is possible in polynomial time.
+	Bian, Gu, and Zhu\cite{BGZ15} describe and benchmark some implementations.
 
-	The concepts of medial graph and dual graph are used in the algorithm. Both are planar graphs, that are constructed from a plane graph.
+	This paper describes an implementation of the algorithm and publishes the source code for anyone to examine, use, and modify.
 
-	The dual graph has vertices for faces and edges connecting the faces that share an edge.
-You might imagine a map of countries, with edges for borders, and a vertex for each country, plus a vertex for the surrounding sea.
-You could for instance query such a network for which countries are neighbors, or which countries are landlocked.
+	% You can obtain an approximate optimum tree decomposition from the branch decomposition.
+So computing an approximate treewidth for planar graphs is possible in polynomial time.
+This is a significant result, as there is no known polynomial time algorithm for computing the treewidth of a planar graph.
 
-	The medial graph also considers the faces but has vertices for edges, that are connected if the original edges were consecutive in some face. I have found no good metaphors for this.
+	% The concepts of medial graph and dual graph are used in the algorithm. Both are graphs derived from some plane graph. Both are themselves plane.
 
-	Pino\cite{Pin16} applies branch decompositions.
+	% The dual graph has vertices for faces and edges connecting the faces that share an edge.
 
-	Seymour and Thomas\cite{ST93} give the rat-catching algorithm.
-
-	Bian, Gu and Zhu\cite{BGZ15} describe and benchmark some implementations.
-
-	\todo{Read params book about tree width}
+	% The medial graph also considers the faces, but has vertices for edges, that are connected if the original edges were consecutive in some face.
 
@@
 
-	This section defines the terms and concepts used in the algorithm.
+	This section defines the terminology and concepts used in the algorithm.
 
@@
 
-	A \textit{graph} $G$ consists of a vertex set $V(G)$, and an edge set $\mathbb{E}(G)$ and a function $\phi_G$, where $V(G) \subset \mathbb{N}^+$ and where $\mathbb{E}(G) \subset \mathbb{N}^+$ and where $\phi_G \st \mathbb{E}(G) \to \{\{u,v\} \st u,v \in V(G)\}$.
+	A \textit{graph} $G$ consists of a vertex set $V(G)$, and an edge set $\mathbb{E}(G)$ and a relation $\phi_G$, where $V(G) \subset \mathbb{N}^+$ and where $\mathbb{E}(G) \subset \mathbb{N}^+$ and where $\phi_G \st \mathbb{E}(G) \to \{\{u,v\} \st u,v \in V(G)\}$.
 
@@
 
-	Here is a drawing of a graph $G$.
-
-	\begin{center}
-		\input{images/unlabeled-graph.tex}
-	\end{center}
-
-	Here is a labeled drawing of the same graph $G$ and its function $\phi_G$.
+	Here is a drawing of a graph $G$ and its relation $\phi_G$.
 
@@
 
-	A \textit{self-loop}, is an edge $e$ where $\phi_G(e) = \{u,v\}$ and $u = v$.
+	Two edges $e_1,e_2 \in \mathbb{E}(G)$ are \textit{parallel}, if $\phi_G(e_1) = \phi_G(e_2)$.
 
-	A graph $G$ is \textit{loop-less}, if no edge $e \in \mathbb{E}(G)$ is a self-loop. The edge with label 7, in the graph above, is an example of a self-loop.
+	An edge $e_1 \in \mathbb{E}(G)$ is a \textit{parallel edge} if there exists an edge $e_2 \in \mathbb{E}(G)$ such that $e_1$ and $e_2$ are parallel.
+The edge labeled $5$ and $6$, in the graph above, are examples of parallel edges.
+
+	A \textit{self-loop}, is an edge $e$ where $\phi_G(e) = \{u,v\}$ and $u = v$. The edge with label 7, in the graph above, is an example of a self-loop.
+
+	A graph $G$ is \textit{loop-less}, if no edge $e \in \mathbb{E}(G)$ is a self-loop.
 
@@
 
-	The \textit{maximum degree} of a graph $G$, denoted $\Delta(G)$, is the maximal degree of any vertex of $G$.
+	The \textit{maximum degree} of a graph $G$, denoted $\Delta(G)$, is the maximum degree of any vertex of $G$.
 
@@
 
-	A \textit{rotation system} is an encoding of a graph, in particular, it is an adjacency list such that the neighborhood, around any vertex, is in clockwise ordering according to some plane embedding of $G$.
+	A \textit{rotation system} is an encoding of a graph, in particular, an adjacency list such that the list of neighbors, of every vertex, is sorted in clockwise ordering according to some drawing of $G$.
+If you encode a plane graph as a rotation system, you lose the coordinates of the vertices, but you can still recover the faces, to recover an equivalent plane graph.
 
@@
 
-		A contraction is a function that given a multi-graph (with no self-loops) $G$ and pair of distinct vertices $u,v$ of some edge $\{u,v\} \in E(G)$, removes $e$ if $\phi_G(e) = \{u,v\}$ and makes any edge that is adjacent to $v$, $\phi_G(e) = \{w,v\}$, adjacent to $u$ instead, $\phi_G(e) = \{w,u\}$, and finally returns the resulting graph.
+		A contraction is an operation that given a graph $G$ and pair of distinct vertices $u,v$ of some edge $\{u,v\} \in E(G)$, removes $e$ if $\phi_G(e) = \{u,v\}$ and makes any edge that is adjacent to $v$, $\phi_G(e) = \{w,v\}$, adjacent to $u$ instead, $\phi_G(e) = \{w,u\}$, and finally returns the resulting graph.
 	\end{definition}
@@
 	\begin{corollary}
-		The resulting graph of a contraction will not have any self-loops.
+		The resulting graph of a contraction will not have any new self-loops.
 	\end{corollary}
@@
 
-	I will refer to vertices and edges of the medial graph as "nodes" and "links" in an attempt at disambiguation.
+	I will refer to vertices and edges of the medial graph as \textit{nodes} and \textit{links} in an attempt at disambiguation.
 
@@
 The set $\cup\{\phi_G(e) \st e \in L(P_e)\} \cap \cup\{\phi_G(e) \st e \in L(Q_e)\}$ is called a \textit{middle set} of $B_G$ given $e$, denoted $Z(B_G,e)$.
-The maximal cardinality of any middle set of $B_G$ given any $e \in \mathbb{E}(B_G)$ is the width of $B_G$; in other words, the width of $B_G$ is $\max\{ |Z(B, e)| \st e \in \mathbb{E}(B) \}$.
+The maximum cardinality of any middle set of $B_G$ given any $e \in \mathbb{E}(B_G)$ is the width of $B_G$; in other words, the width of $B_G$ is $\max\{ |Z(B, e)| \st e \in \mathbb{E}(B) \}$.
 
-	\todo{middle set definition is convoluted}
-
-	A \textit{Minimum Branch Decomposition} of $G$ is any branch decomposition of $G$ of minimum width among all branch decompositions of $G$, as there might exist many branch decompositions of a graph $G$.
+	A \textit{Minimum Branch Decomposition} of $G$ is any branch decomposition of $G$ of minimum width among all branch decompositions of $G$.
+Note there might exist many branch decompositions of a graph $G$, as well as, multiple minimum branch decompositions of a graph $G$.
 
@@
 	Removing any edge $e \in \mathbb{E}(C_G)$ partitions $C_G$ into 2 trees $P_e$ and $Q_e$.
-The set of edges $e \in \mathbb{E}(G)$ that cross $L(P_e)$ and $L(Q_e)$ is called a \textit{crossing set} of $C_G$ given an $e \in \mathbb{E}(C_G)$, denoted $Y(C_G,e)$.
+The set of edges $e \in \mathbb{E}(G)$ that cross from $L(P_e)$ to $L(Q_e)$ is called a \textit{crossing set} of $C_G$ given an $e \in \mathbb{E}(C_G)$, denoted $Y(C_G,e)$.
 The maximum cardinality of any crossing set of $C_G$ given any $e \in \mathbb{E}(C_G)$ is the width of $C_G$; in other words, the width of $C_G$ is $\max\{ |Y(C, e)| \st e \in \mathbb{E}(C) \}$.
 
-	A \textit{Minimum Carving Decomposition} of $G$ is any carving decomposition of $G$ of minimum width among all carving decompositions of $G$, as there might exist many carving decompositions of a graph $G$.
+	A \textit{Minimum Carving Decomposition} of $G$ is any carving decomposition of $G$ of minimum width among all carving decompositions of $G$.
 
@@
 	\end{center}
-
-	\todo{find example with non-mimimum carving decomposition}
 
@@
 
-	To compute the carving width of $G$, the algorithm uses the rat-catching algorithm.
-By an analogy, the rat-catching algorithm can be described by a game of two players, the rat and rat-catcher, who take turns moving around on $G$.
-The game is played to determine, if the rat-catcher can corner the rat by making noise and thereby scaring away the rat from some subgraph, the size of which is dependent on $k$ or if the rat can escape indefinitely.
-Larger $k$ makes the rat-catching game easier for the rat-catcher.
-The smallest $k$ where the rat can escape indefinitely is the carving width of $G$.
+	To compute the carving width of a graph $H$, the algorithm uses the rat-catching algorithm.
+By analogy, the rat-catching algorithm can be described as a game of two players, the rat, and the rat-catcher, who take turns moving around on $H$.
+The game is played to determine, if the rat can escape indefinitely or if the rat-catcher can corner the rat by making noise and thereby scaring away the rat from some subgraph, the size of which is dependent on $k$.
+Larger $k$ makes the rat-catching game easier for the rat-catcher and vice versa.
+The smallest $k$ where the rat can escape indefinitely is the carving width of $H$.
 
-	To compute a noisy subgraph of $G$ given a noise level $k$ and a rat-catchers position, the algorithm considers the dual graph of $G$ and finds the set of dual edges reachable by a closed walk of length at most $k-1$, the dual of these edges is the noisy subgraph.
+	To compute a noisy subgraph of $H$ given a noise level $k$ and a rat-catcher position, the algorithm considers the dual graph of $H$ and finds the set of dual edges reachable by a closed walk of length at most $k-1$, the dual of these edges is the noisy subgraph.
+
+	All of these subroutines are described in more detail in the following sections.
 
@@
 
-\section{Data Structures and implementation considerations}
+\section{Data Structures and Implementation Considerations}
 
-	For many of the subroutines, the algorithm needs to deal with parallel edges and be able to tell them apart, therefore the implementation encodes a graph as an adjacency list of edges and a map from unique edge IDs its vertex-pair.
+	For many of the subroutines, the algorithm needs to deal with parallel edges and be able to tell them apart.
+Therefore the implementation encodes a graph as an adjacency list of edge IDs and a map from unique edge IDs its vertex-pair.
+
+	A \textit{half-edge} is one of the two directed edges that make up an undirected edge in an adjacency list of edge IDs.
 
@@
 	\pyfile{graph.py}{0}{33}
+
+	Rotation systems are of type \mintinline{py}{Graph} but the \mintinline{py}{adj_edges} list is sorted in clockwise ordering according to some drawing of the graph.
+
+	Carving-decompositions are adjacency lists \mintinline{py}{dict[int, list[int]]}.
+
+	Branch-decompositions are adjacency lists with either integers or integer pairs as vertices\\ \mintinline{py}{dict[int, list[Union[int, tuple[int, int]]]]}.
 
@@
 
-	Therefore problem \ref{def:mbdp} breaks down into subproblems \ref{problem:medial}, \ref{problem:mcd} and \ref{problem:mcd-to-mbd}.
+	Therefore problem \ref{def:mbdp} breaks down into subproblems \ref{problem:medial}, \ref{problem:mcd}, and \ref{problem:mcd-to-mbd}.
 
@@
 
-	\pyfile{branch_decomposition.py}{6}{22}
+	\pyfile{branch_decomposition.py}{19}{27}
 
-	Here are three sections, one for each subproblem.
+	The next 3 subsections describe the implementation of the subroutines for the problems \ref{problem:medial}, \ref{problem:mcd}, and \ref{problem:mcd-to-mbd}.
 
@@
 
-		Recall the definition \ref{def:medial}.
-
 		I assume that the input graph $G$ is a rotation system.
@@
 
-		\begin{figure}[H]
-			\centering
-			\includegraphics[width=\textwidth]{./images/medial-implementation-4-reg-argument.png}
-			\caption{Two iterations of the algorithm computing a medial rotation system}
-			\label{fig:overview}
-		\end{figure}
+		A medial graph is 4-regular; every node has degree 4.
+Exploiting this property, the implementation iterates over the faces of $G$, and for each pair of clockwise consecutive edges $w$ and $v$ in some face, adds a link between the nodes corresponding to $w$ and $v$ in the medial graph.
 
-		A medial graph is 4-regular; every node has degree 4.
-From the perspective of some medial node $v$, in some single iteration of the loop on line 14, two links are added to the neighborhood of $v$ in counterclockwise ordering, and later the two other links are added to the neighborhood of $v$ also in counterclockwise ordering.
+		From the perspective of a medial node $v^\times$, the first two links are added to its neighborhood in counterclockwise ordering, after processing the first face, and later the final two links are added in counterclockwise ordering, after processing the second face.
 Therefore the medial graph $M$ is a rotation system.
 
-		\pyfile{medial_graph.py}{5}{27}
+		\pyfile{medial_graph.py}{4}{54}
 
@@
 
-		For both branch- and carving-decompositions, I have chosen a data structure of tuples of either more tuples or integers.
-This has a straightforward translation to the Newick Tree Format, a concise notation for tree structures.
+		The implementation identifies the leaf vertices and maps them to vertex-pairs according to the relation \mintinline{py}{node_to_vertexpair} from the computation of the medial graph.
 
-		To then solve the above-mentioned problem, the implementation recursively returns either returns a copy of any tuple it encounters or returns a tuple of integers for any integer in encounters, using the relation node\_to\_vertexpair.
-
-		This effectively replaces the leaf nodes of the carving decomposition with vertex-pairs for the branch decomposition.
-
-		\pyfile{branch_decomposition.py}{5}{19}
+		\pyfile{branch_decomposition.py}{6}{17}
 
@@
 
-		To solve \ref{problem:mcd} \ref{} gives a contraction algorithm.
+		Solving problem \ref{problem:mcd}.
 
-		By doing a series of edge contractions on a graph $M$, where the carving width does not increase until 3 vertices remain, then the series of contracted edges along with the three vertices, can be assembled into a minimum carving decomposition of $M$.
+		By doing a series of edge contractions on a graph $M$, where the carving width does not increase, until 3 vertices remain, then the series of contracted edges along with the 3 vertices, can be assembled into a minimum carving decomposition of $M$.
 
 		The implementation finds a nonincreasing contraction by doing a linear search over every edge.
-
 No consideration has yet been given to any potential clever orderings of the edges that might improve the running time.
 
-		The "contraction" function, given a pair of vertices $u,v$, returns a new unique vertex ID $w$ (instead of reusing either $u$ or $v$), therefore by keeping track of which vertex is a contraction of which vertex-pair, in say a dictionary, constructing the decomposition is then a matter of recursively looking up vertices in the dictionary.
-Repeating this until only vertices of $M$ remain gives a carving decomposition in the aforementioned Newick-like nested tuple format.
+		The \mintinline{py}{contraction} function, given a pair of vertices $u,v$, returns a new unique vertex ID $w$ (instead of reusing $u$ as definition \ref{def:contraction} suggests), therefore by keeping track of which vertex is a contraction of which vertex-pair, in a dictionary, constructing the decomposition is then a matter of recursively looking up vertices in the dictionary.
+Repeating this until only vertices of $M$ remain gives a carving decomposition.
 
-		\pyfile{carving_decomposition.py}{7}{50}
+		\pyfile{carving_decomposition.py}{7}{79}
 
-		The contraction algorithm depends on a function to compute a contraction and a function to compute the carving width of a graph. This is problems \ref{problem:contraction} and \ref{problem:cw}.
+		The contraction algorithm depends on a function to compute a contraction and a function to compute the carving width of a graph.
+These are referred to as problems \ref{problem:contraction} and \ref{problem:cw}.
 
@@
 
-			As the resulting graph is later given as an argument to functions assuming a rotation system of a planar graph, the implementation needs to preserve this invariant when contracting.
+			The implementation needs to preserve the rotation system, as the resulting graph is later given as an argument to functions assuming a rotation system of a planar graph.
 
 			Recall the definition \ref{def:contraction}.
-As this contraction removes all edges connecting a pair of vertices, the resulting graph will not exhibit any self-loops.
+As this contraction removes all edges connecting a pair of vertices, the resulting graph will not create any new self-loops.
 I suspect reconciling this and the rotation system could be difficult, but in this context, it is irrelevant.
@@
 This edge has some ID $e$ and the other half-edge, ID $-e$, will therefore be in the neighborhood of $b$.
-Now "rotating" the lists representing the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0, in their respective lists, means that a concatenation of the lists will preserve the clockwise ordering around the new vertex $c$.
+Now \textit{rotating} the lists representing the neighborhoods of $a$ and $b$ such edge $e$ and $-e$ is at index 0, in their respective lists, means that a concatenation of the lists will preserve the clockwise ordering around the new vertex $c$.
 And finally, remove any edges connecting $a$ and $b$.
@@
 			This is where telling apart two edges that are parallel, by having edge IDs, becomes very useful.
-Inferring where to stitch together the neighborhoods to preserve the clockwise ordering, solely from an adjacency list, is way less practical.
+Inferring where to stitch together the neighborhoods to preserve the clockwise ordering, solely from an adjacency list, is not a trivial problem.
 
-			\pyfile{contraction.py}{4}{49}
+			\pyfile{contraction.py}{2}{46}
 
@@
 
-			\pyfile{carving_width.py}{7}{8}
-			\pyfile{carving_width.py}{132}{158}
+			\pyfile{carving_width.py}{5}{6}
+			\pyfile{carving_width.py}{130}{156}
 
@@
 
-		Additionally, if $\Delta(M) \geq k$ then the rat wins. The argument for why this is true is glossed over in \ref{BGZ15}. This is discussed in section \ref{}.
+		Additionally, if $\Delta(M) \geq k$ then the rat wins. The argument for why this is true is glossed over in \ref{BGZ15}.
 
@@
 
-		For some noise level and location of the rat-catcher, exactly which edges are noisy and which are quiet are definitions \ref{def:edge-quiet} and \ref{def:face-quiet}.
+		For some noise level and location of the rat-catcher, exactly which edges are noisy and which are quiet are definitions \ref{def:edge-noisy} and \ref{def:face-quiet}.
 
-		\begin{definition}\label{def:edge-quiet}
+		\begin{definition}\label{def:edge-noisy}
 			When the rat-catcher is on some edge $e_1$, then edge $e_2$ is noisy iff. there is a closed walk of length scrictly less than $k$ containing $e_1^*$ and $e_2^*$ in the dual $M^*$.
@@
 
-		A quiet subgraph $Q(M, k, e)$, for some graph $M$, some noise level $k$ and some $e \in \mathbb{E}(M)$, is a subgraph of $M$ with the vertex set $V(Q(M, k, e)) = V(M)$ and the edge set
+		A \textit{quiet subgraph} $Q(M, k, e)$, for some graph $M$ and some noise level $k$ and some edge $e \in \mathbb{E}(M)$, is a subgraph of $M$ with the vertex set $V(Q(M, k, e)) = V(M)$ and the edge set
 
@@
 
-		Let $S$ be every possible state when the rat-catcher is in a face some of which might be losing states. $S = \{ (f, v) \st v \in V(M) \land f$ is a face of $ M \}$.
+		Let $S$ be every possible state when the rat-catcher is in a face, some of which might be losing states. $S = \{ (f, v) \st v \in V(M) \land f$ is a face of $ M \}$.
 
@@
 
-		You can mark states/vertices of the graph $H$ that are losing states, and then repeatedly mark any state that leads to a losing state, until either every state is marked or no more states can be marked.
+		You can mark vertices of the graph $H$ that are losing states, and then repeatedly mark any state that leads to a losing state, until either every state is marked or no more states can be marked.
 If every state is marked then the rat-catcher wins, otherwise the rat wins.
@@
 
-		Considering a tuple $(e, C) \in T$, if all $(f, v)$ where $v \in V(C)$ is losing then $(e, C)$ is losing.
+		Considering a state $(e, C) \in T$, if all $(f, v)$ where $v \in V(C)$ is losing then $(e, C)$ is losing.
 
-		Considering a tuple $(f, v)$, if there exists a tuple $(e, C)$ that is losing where $e \in f$ and $v \in V(C)$ then $(f, v)$ is losing.
+		Considering a state $(f, v)$, if there exists a state $(e, C)$ that is losing where $e \in f$ and $v \in V(C)$ then $(f, v)$ is losing.
 
-		\pyfile{carving_width.py}{78}{130}
+		\pyfile{carving_width.py}{79}{128}
 
@@
 
-			Recall the definition \ref{def:edge-quiet}.
+			Recall the definition of a noisy edge \ref{def:edge-noisy}.
 
@@
 
-			Using the mapping from links to edges, and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained in the natural way.
+			Using the mapping from links to edges, and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained by applying the map.
 
@@
 
-			\pyfile{carving_width.py}{10}{76}
+			\pyfile{carving_width.py}{6}{74}
 
@@
 \section{Results}
+	The algorithm has been implemented in Python 3.12.2. The source code is available on GitHub \footnote{\url{https://github.com/hojelse/thesis}}.
 
-	The algorithm has been implemented in Python 3.12.2. The source code is available at \url{https://github.com/hojelse/thesis}.
+	The implementation has been tested against two sets of graphs, using various methods.
 
-	The implementation has been tested on a set of graphs, and the results are as expected. \todo{more}
+	The first set of graphs are random cubic planar graphs.
+The second set of graphs are small simple planar graphs generated by \textit{Plantri}\footnote{\url{http://users.cecs.anu.edu.au/~bdm/plantri/}}\footnote{\url{https://github.com/mishun/plantri/tree/master}}.
 
-	\todo{Agreement with Andreas' implementation on 6-100 vertices graphs, 10 different graphs}
+	The random cubic planar graphs is a set of 23 graphs, that were generated by a script, created by Andreas Björklund.
+See appendix \ref{graphs/gen-random-planar-cubic.py} for the generator script.
+The graphs are available on GitHub \footnote{\url{https://github.com/hojelse/thesis/tree/main/code/graphs/random-planar-cubic}}.
+They also created a branch width algorithm that was cross-validated with the implementation from this paper.
+The two algorithms agreed on every value.
+See appendix \ref{test-cubic-cross-validate.py} for the script.
 
-	\todo{note Tested on cubic planar graph}
+	The second set of small simple planar graphs was generated by Plantri, which exhaustively generated all $2228$ simple planar graphs on 2 to 7 vertices.
+Testing the implementation on these graphs, the results are shown in figure \ref{fig:bw-planar}.
+The scripts for generating, processing, and testing this set of graphs are in appendix \ref{graphs/gen-planar.py}, \ref{graphs/process_planar_code.py}, and \ref{test-planar-graphs.py}.
+The graphs are available on GitHub \footnote{\url{https://github.com/hojelse/thesis/tree/main/code/graphs/simple-planar}}
 
-	\todo{tested by permuting vertex labels should not change the branch width}
+	Fomin and Thilikos\cite{FT06} find a new theoretical upper bound on the branch width of planar graphs, $bw(G) \leq 4.5 \cdot \sqrt{n}$, where $n$ is the number of vertices in the graph.
 
-	\todo{test that bw follows theoretical bound}
+	Some of the graphs are suspected to be cause the implementation to go into an infinite loop, as the branch width is not computed within a reasonable time.
+So the implementation is suspectedly not correct for all planar graphs.
 
-	Fomin and Thilikos 2006, New upper bound obn the decomposability pf planar graphs, bw <= 4.5 * sqrt(N) on planar
+	\begin{figure}[H]
+		\centering
+		\includegraphics[width=\textwidth]{./images/bw-planar.png}
+		\caption{Branch width of planar graphs generated by Plantri.
+The x-axis is the number of vertices in the graph, and the y-axis is the branch width.
+The red line is $\lfloor 4.5 \cdot \sqrt{n} \rfloor$.
+The blue dots are the branch width and the number directly next to it is the number of the graph with the same unique (n, bw) pair.}
+		\label{fig:bw-planar}
+	\end{figure}
 
-	Semour Thomas 1994, Call routing and ratcatcher, bw(G) = bw(dual(G)) assuming G is bridgeless planar graphs
+	On the same set of graphs, the implementation was tested against itself by permuting the vertex labels of input graphs, which should not change the branch width.
+This was not done exhaustively, but the results were as expected.
 
-	then the dual has half the vertices of the primal
+	A tool, \textit{grafer}, was developed for visualizing and editing planar graphs and used for debugging during this thesis \url{https://github.com/hojelse/grafer}.
 
-	4.5 * sqrt(N/2)
+\section{Future work}
+	The implementation was only tested on cubic planar graphs until late in the process of writing this thesis, as there were plans to use this branch-decomposition algorithm as a subroutine in a larger algorithm.
+As seen in the results, the implementation does not behave nicely for all planar graphs.
+Debugging the implementation on the simple planar graphs generated by Plantri that fail is a good next step.
 
-	<=> 3.18199 * sqrt(n)
-
-	Test bw <= 3.18199 * sqrt(N) on cubic planar graphs
-
-	\todo{graphing theoretical bound}
+\section{Conclusion}
+	This paper presents a clear and accessible implementation of the rat-catching algorithm for planar graphs, based on the work of Seymour and Thomas.
+The implementation follows the original algorithm and is supported by a framework for testing correctness to ensure its reliability.
+The results regrettably show that there are still bugs to eliminate.
+The source code provided alongside this paper is intended as a practical resource for those working on algorithm design involving branch- and carving decompositions.
+I hope this work will be useful for students, researchers, or practitioners looking to learn about, use, or develop a branch-decomposition algorithm by examining, using, or modifying the code provided alongside this paper.
+\url{https://github.com/hojelse/thesis}.
 
@@
 \section{Appendix}
-	% \pyfile{branch_decomposition.py}{0}{1000}
-	% \pyfile{branch_width_brute_force.py}{0}{1000}	
-	% \pyfile{branch_width.py}{0}{1000}
-	% \pyfile{carving_decomposition.py}{0}{1000}
-	% \pyfile{carving_width_brute_force.py}{0}{1000}
-	% \pyfile{carving_width.py}{0}{1000}
-	% \pyfile{contraction.py}{0}{1000}
-	% \pyfile{dual_graph.py}{0}{1000}
-	% \pyfile{Graph.py}{0}{1000}
-	% \pyfile{medial_graph.py}{0}{1000}
-	% \pyfile{parse_graph.py}{0}{1000}
-	% \pyfile{parse_newick.py}{0}{1000}
-
+	\subsection{}
+	\pyfile{branch_decomposition.py}{0}{1000}\label{branch_decomposition.py}
+	\subsection{}
+	\pyfile{branch_width.py}{0}{1000}\label{branch_width.py}
+	\subsection{}
+	\pyfile{carving_decomposition.py}{0}{1000}\label{carving_decomposition.py}
+	\subsection{}
+	\pyfile{carving_width.py}{0}{1000}\label{carving_width.py}
+	\subsection{}
+	\pyfile{contraction.py}{0}{1000}\label{contraction.py}
+	\subsection{}
+	\pyfile{dual_graph.py}{0}{1000}\label{dual_graph.py}
+	\subsection{}
+	\pyfile{Graph.py}{0}{1000}\label{Graph.py}
+	\subsection{}
+	\pyfile{medial_graph.py}{0}{1000}\label{medial_graph.py}
+	\subsection{}
+	\pyfile{test-cubic-cross-validate.py}{0}{1000}\label{test-cubic-cross-validate.py}
+	\subsection{}
+	\pyfile{test-planar-graphs.py}{0}{1000}\label{test-planar-graphs.py}
+	\subsection{}
+	\pyfile{util.py}{0}{1000}\label{util.py}
+	\subsection{}
+	\pyfile{graphs/gen-random-planar-cubic.py}{0}{1000}\label{graphs/gen-random-planar-cubic.py}
+	\subsection{}
+	\pyfile{graphs/gen-planar.py}{0}{1000}\label{graphs/gen-planar.py}
+	\subsection{}
+	\pyfile{graphs/process_planar_code.py}{0}{1000}\label{graphs/process_planar_code.py}
+	\subsection{}
 \end{document}

### 2024-09-01 code/Graph.py

@@ -28,6 +28,21 @@ class Graph:
 
 	def N(self, v: int) -> list[int]:
 		return [self.edge_to_vertexpair[e][1] if self.edge_to_vertexpair[e][0] == v else self.edge_to_vertexpair[e][0] for e in self.adj_edges[v]]
+
+	def rename_vertex(self, old: int, new: int):
+		self.adj_edges[new] = self.adj_edges.pop(old)
+		for e in self.E():
+			u,v = self.edge_to_vertexpair[e]
+			if u == old:
+				self.edge_to_vertexpair[e] = (new, v)
+			if v == old:
+				self.edge_to_vertexpair[e] = (u, new)
+
+	def rename_edge(self, old: int, new: int):
+		self.edge_to_vertexpair[new] = self.edge_to_vertexpair.pop(old)
+		for u,vs in self.adj_edges.items():
+			if old in vs:
+				self.adj_edges[u][vs.index(old)] = new
 	
 	def adj(self) -> dict[int, list[int]]:
 		return dict([(x, self.N(x)) for x in self.adj_edges.keys()])
@@ -67,10 +82,14 @@ class Graph:
 	
 def read_lmg_from_stdin() -> str:
 	N,M = map(int, input().split())
-	stdin_multiline_string = f"{N}, {M}\n"
+	stdin_multiline_string = f"{N} {M}\n"
 	for _ in range(N+M):
 		stdin_multiline_string += input() + "\n"
 	return stdin_multiline_string
+
+def read_lmg_from_file(filename: str) -> str:
+	with open(filename) as f:
+		return f.read()
 
 if __name__ == "__main__":
 	G = Graph()

### 2024-09-01 code/branch_decomposition.py

@@ -1,27 +1,32 @@
-import log
-from parse_graph import parse_text_to_adj
-from medial_graph import medial_graph, medial_graph_2
+from typing import Union
+from Graph import Graph, read_lmg_from_stdin
+from medial_graph import medial_graph
 from carving_decomposition import carving_decomposition
+from util import adj_to_str
+
+def carving_decomposition_to_branch_decomposition(
+	cd: dict[int, list[int]],
+	node_to_vertexpair: dict[int, tuple[int, int]]
+) -> dict[int, list[Union[int, tuple[int, int]]]]:
+
+	leafs = [u for u,vs in cd.items() if len(vs) == 1]
+
+	bd = dict()
+	for u,vs in cd.items():
+		bd[u if u not in leafs else node_to_vertexpair[u]] = [v if v not in leafs else node_to_vertexpair[v] for v in vs]
+
+	return bd
 
 # Construct a branch decomposition of a graph
-def branch_decomposition(G_adj: dict[int, list[int]]):
-	# Contruct the carving decomposition of the medial graph
-	Gx, node_to_vertexpair = medial_graph(G_adj)
-
+def branch_decomposition(G: Graph) -> dict[int, list[Union[int, tuple[int, int]]]]:
+	Gx, node_to_vertexpair = medial_graph(G)
 	cd = carving_decomposition(Gx.copy())
-
-	# Convert the carving decomposition of M to a branch decomposition of G
-	def decomp(t):
-		if isinstance(t, int):
-			return node_to_vertexpair[t]
-		return tuple([decomp(a) for a in t])
-	
-	bd = decomp(cd)
-
-	log.add("Branch decomposition: " + str(bd))
+	bd = carving_decomposition_to_branch_decomposition(cd, node_to_vertexpair)
+	print(adj_to_str(bd))
 	return bd
 
 if __name__ == "__main__":
-	adj = parse_text_to_adj()
-	bd = branch_decomposition(adj)
-	print(bd)
+	G = Graph()
+	G.from_lmg(read_lmg_from_stdin())
+	bd = branch_decomposition(G)
+	print("bd", bd)

### 2024-09-01 code/branch_width.py

@@ -1,23 +1,9 @@
-import log
+from typing import Union
+from Graph import Graph
 from branch_decomposition import branch_decomposition
-from parse_graph import parse_text_to_adj, adj_to_text
+from util import adj_from_stdin
 
-def branch_width_of_branch_decomposition(bd):
-	# Create an adjacency list from the branch decomposition
-	T_adj = dict()
-	def aux(subtree, depth, name):
-		if len(subtree) == 2 and isinstance(subtree[0], int) and isinstance(subtree[1], int):
-			T_adj[subtree] = []
-			return subtree
-		else:
-			T_adj[name] = []
-			for i,a in enumerate(subtree):
-				child_name = aux(a, depth+1, name+str(i))
-				T_adj[name].append(child_name)
-				T_adj[child_name].append(name)
-			return name
-	aux(bd, 0, "i0")
-
+def branch_width_of_branch_decomposition(bd: dict[int, list[Union[int, tuple[int, int]]]]) -> int:
 	# Get the vertex set of the leafs of the subtree of x (not y)
 	def leafs_set(x, y):
 		leafs = set()
@@ -30,30 +16,35 @@ def branch_width_of_branch_decomposition(bd):
 				continue
 			if v not in visited:
 				visited.add(v)
-				for w in T_adj[v]:
+				for w in bd[v]:
 					stack.append(w)
 		return leafs
 
 	# Find the maximal width of any middle set
 	width = 0
-	for x,ys in T_adj.items():
+	for x,ys in bd.items():
 		for y in ys:
 			a = leafs_set(x, y)
 			b = leafs_set(y, x)
 			middle_set = len(a.intersection(b))
 			width = max(width, middle_set)
-	
-	log.add("Branch width of branch decomposition: " + str(width))
+
 	return width
 
-def branch_width(adj: dict[int, list[int]]):
-	bd = branch_decomposition(adj)
+def branch_width(G: Graph) -> int:
+	bd = branch_decomposition(G)
 	bw = branch_width_of_branch_decomposition(bd)
-	log.add("Branch width: " + str(bw))
+
 	return bw
 
 if __name__ == "__main__":
-	adj = parse_text_to_adj()
-	bd = branch_decomposition(adj)
-	bw = branch_width_of_branch_decomposition(bd)
-	print("bw", bw)
+	# G = Graph()
+	# G.from_lmg(read_lmg_from_stdin())
[…]

### 2024-09-01 code/bw_from_bd.py

@@ -1,9 +1,0 @@
-from branch_width import branch_width_of_branch_decomposition
-from parse_newick import parse_newick
-
-s = input()
-print(s)
-nw = parse_newick(s)
-print(nw)
-bw = branch_width_of_branch_decomposition(nw)
-print(bw)

### 2024-09-01 code/carving_decomposition.py

@@ -1,55 +1,85 @@
-import log
+from collections import defaultdict
+from typing import Union
 from carving_width import carving_width
 from contraction import contraction
-from parse_graph import parse_text_to_adj, adj_to_text
-from Graph import Graph
+from Graph import Graph, read_lmg_from_stdin
 
 # Find a contraction that does not increase the carving width
-def nonincreasing_cw_contraction(G: Graph, cw1: int) -> tuple:
+def nonincreasing_cw_contraction(G: Graph, cw1: int) -> Union[tuple[Graph, tuple[int, int], int, int], None]:
 	for e in G.E():
 		u, v = G.edge_to_vertexpair[e]
+		if u == v:
+			continue
 		G2, w = contraction(G, u, v)
 		cw2 = carving_width(G2)
 		if cw2 <= cw1:
-			log.add(f"Graph after contracting edge {e}: \n{str(G2)}")
 			return G2, (u, v), cw2, w
-	return None, None, None, None
+	return None
 
 # Contract edges that do not increase the carving width
 # until only 3 vertices remain.
 # Return the resulting graph and the edges that were contracted
-def gradient_descent_contractions(G: Graph) -> Graph:
+def gradient_descent_contractions(G: Graph) -> tuple[Graph, dict[int, tuple[int, int]]]:
 	G2 = G.copy()
 	cw1 = carving_width(G)
 	edges = dict()
 	while True:
-		G3, uv, cw2, w = nonincreasing_cw_contraction(G2, cw1)
-		if G3 is not None and len(G3.V()) >= 3:
-			G2 = G3
-			cw1 = cw2
-			edges[w] = uv
-		if len(G2.V()) == 3:
+		if len(G2.V()) <= 3:
 			return G2, edges
+		res = nonincreasing_cw_contraction(G2, cw1)
+		if res is not None:
+			G3, uv, cw2, w = res
+			if len(G3.V()) >= 3:
+				G2 = G3
+				cw1 = cw2
+				edges[w] = uv
 
 # Contruct a carving decomposition of a graph
-def carving_decomposition(G: Graph) -> tuple:
+def carving_decomposition(G: Graph) -> dict[int, list[int]]:
 	G2, edges = gradient_descent_contractions(G)
 
-	# Construct the decomposition from the edges that were contracted
[…]

### 2024-09-01 code/carving_width.py

@@ -1,7 +1,5 @@
 import math
-
-from Graph import Graph
-from parse_graph import adj_to_text, adj_to_text_2, parse_text_to_adj
+from Graph import Graph, read_lmg_from_stdin
 from dual_graph import dual_graph
 
 def carving_width(G: Graph) -> int:
@@ -89,13 +87,13 @@ def carving_width(G: Graph) -> int:
 		if max([len(G.N(v)) for v in G.V()]) >= k:
 			return True
 
-		# Set up the game states
+		# Set up the game graph
 		halfedges = edge_to_link.keys()
 
 		T = set([(e, tuple(C)) for e in halfedges for C in quiet_components(e, k)])
 		S = set([(f, v) for f in node_to_face.keys() for v in G.V()])
 
-		# Set up the losing states
+		# Set up the initial losing states
 		losing_T = set()
 		losing_S = set()
 
@@ -158,9 +156,7 @@ def carving_width(G: Graph) -> int:
 	return cw
 
 if __name__ == "__main__":
-	adj = parse_text_to_adj()
-
 	G = Graph()
-	G.from_adj(adj)
+	G.from_lmg(read_lmg_from_stdin())
 	cw = carving_width(G)
 	print("cw", cw)

### 2024-09-01 code/compare_bw.py

@@ -1,17 +1,0 @@
-import subprocess
-import time
-import datetime
-import math
-
-ns = [6,8,10,14,18,22,26,30,34,38,42,46,50,54,58,62,66,70,74,78,82,86,90,94,98]
-
-for n in ns:
-	with open(f"../graphs/g/{n}-5.in", 'rb') as f:
-		start_time = time.time()
-		r = subprocess.run(['python3', 'branch_width.py'], stdin=f, text=True, capture_output=True)
-		res = r.stdout.strip()
-		seconds_elapsed = time.time() - start_time
-		print("n", n, end=' ')
-		# print("bound", 2.122*math.sqrt(n))
-		print(res)
-		# print("time", str(datetime.timedelta(seconds=seconds_elapsed)))

### 2024-09-01 code/contraction.py

@@ -1,15 +1,30 @@
-from Graph import Graph
-from parse_graph import adj_to_text, parse_text_to_adj
+from Graph import Graph, read_lmg_from_stdin
 
-# assume G might have parallel edges
-# assume G do not have self-loops
+# assume G might have parallel edges and self-loops
 # assume adjacency list of G has clockwise ordering of neighbors
-def contraction(G: Graph, a: int, b: int) -> Graph:
+# assume a != b
+def contraction(G: Graph, a: int, b: int) -> tuple[Graph, int]:
 	# copy G
 	G1 = G.copy()
 
 	# create new vertex c
 	c = max(G1.adj_edges.keys()) + 1
+
+	# find edges connecting a and b
+	edges_a_b = []
+	for e in G1.adj_edges[a] + G1.adj_edges[b]:
+		u,v = G1.edge_to_vertexpair[e]
+		if (u == a and v == b) or (u == b and v == a):
+			edges_a_b.append(e)
+
+	# create neighborhood of c while preserving clockwise ordering
+	idx1 = G1.adj_edges[a].index(edges_a_b[0])
+	rotated_Ga = G1.adj_edges[a][idx1:] + G1.adj_edges[a][:idx1]
+
+	idx2 = G1.adj_edges[b].index(-edges_a_b[0])
+	rotated_Gb = G1.adj_edges[b][idx2:] + G1.adj_edges[b][:idx2]
+
+	G1.adj_edges[c] = rotated_Ga + rotated_Gb
 
 	# let every edge incident to a or b be incident to c instead
 	for e in G1.E():
@@ -20,27 +35,9 @@ def contraction(G: Graph, a: int, b: int) -> Graph:
 		if v == a or v == b:
 			G1.edge_to_vertexpair[e] = (u, c)
 
-	# create neighborhood of c
-	def index_of_first(lst, pred):
-		for i, v in enumerate(lst):
-			if pred(v):
-				return i
-		return None
-
-	index_of_first_shared_edge = index_of_first(G1.adj_edges[a], lambda e: G1.edge_to_vertexpair[e][0] == c and G1.edge_to_vertexpair[e][1] == c)
-	first_shared_edge = G1.adj_edges[a][index_of_first_shared_edge]
-
-	idx1 = G1.adj_edges[a].index(first_shared_edge)
[…]

### 2024-09-01 code/dual_graph.py

@@ -1,15 +1,14 @@
-from Graph import Graph
-from parse_graph import adj_to_text, parse_text_to_adj
+from Graph import Graph, read_lmg_from_stdin
 
 # Assume G is a rotation system
-def dual_graph(G: Graph) -> Graph:
+def dual_graph(G: Graph) -> tuple[Graph, dict[int, int], dict[int, int], dict[int, list[int]], dict[int, int]]:
 	edges = [e for e in G.E()]
 
 	D = Graph()
-	edge_to_link = dict()
-	link_to_edge = dict()
-	node_to_face = dict() # nodeid to edgeid list
-	edge_to_node = dict() # half-edge to the faceid/node to its either left/right
+	edge_to_link: dict[int, int] = dict()
+	link_to_edge: dict[int, int] = dict()
+	node_to_face: dict[int, list[int]] = dict()  # nodeid to edgeid list
+	edge_to_node: dict[int, int] = dict()        # half-edge to the faceid/node to its either left/right
 
 	# Find faces
 	next_nodeid = -1
@@ -38,7 +37,7 @@ def dual_graph(G: Graph) -> Graph:
 	for i,f1 in node_to_face.items():
 		for j,f2 in node_to_face.items():
 			if i < j:
-				common_edges = set(list(map(abs, f1))).intersection(set(map(abs, f2)))
+				common_edges = set(set(map(abs, f1))).intersection(set(map(abs, f2)))
 				for e in common_edges:
 					D.edge_to_vertexpair[next_linkid] = (i, j)
 					D.edge_to_vertexpair[-next_linkid] = (j, i)
@@ -56,11 +55,10 @@ def dual_graph(G: Graph) -> Graph:
 	return D, edge_to_link, link_to_edge, node_to_face, edge_to_node
 
 if __name__ == "__main__":
-	adj = parse_text_to_adj()
 	G = Graph()
-	G.from_adj(adj)
+	G.from_lmg(read_lmg_from_stdin())
 	D, edge_to_link, link_to_edge, node_to_face, edge_to_node = dual_graph(G)
-	adj_to_text(D.adj())
+	print_adj(D.adj())
 	print("edge_to_link", edge_to_link)
 	print("link_to_edge", link_to_edge)
 	print("node_to_face", node_to_face)

### 2024-09-01 code/is_planar.py

@@ -1,5 +1,0 @@
-import networkx as nx
-from parse_graph import parse_text_to_adj, adj_to_nx
-
-G = adj_to_nx(parse_text_to_adj())
-print(nx.is_planar(G))

### 2024-09-01 code/log.py

@@ -1,7 +1,0 @@
-def init():
-	global log
-	log = []
-
-def add(o):
-	log.append(o)
-	print(o)

### 2024-09-01 code/main.py

@@ -1,23 +1,0 @@
-import log
-import branch_width
-import sys
-import Graph
-
-graph_file_name = sys.argv[1]
-
-log.init()
-log.add("Reading graph from file " + graph_file_name)
-with open(graph_file_name, 'r') as file:
-	graph_str = file.read()
-
-	G = Graph.Graph()
-	if graph_file_name.endswith(".m.in"):
-		G.from_lmg(graph_str)
-	else:
-		print("Unsupported file extension")
-		exit()
-
-	log.add("Graph: " + str(G))
-	log.add("Graph adjacency list: " + str(G.adj()))
-
-	bw = branch_width.branch_width(G.adj())

### 2024-09-01 code/medial_graph.py

@@ -1,45 +1,15 @@
-import log
-from Graph import Graph
-from parse_graph import adj_to_text, adj_to_text_2, parse_text_to_adj
-
-# assume G is simple
-# assume G is planar
-# assume G_adj is a rotation system
-# guarantee M is a rotation system
-def medial_graph(G_adj: dict[int, list[int]]) -> Graph:
-	vertexpairs = set([tuple(sorted((i, j))) for i in G_adj for j in G_adj[i]])
-
-	vertexpair_to_node = dict([(e, i+1) for i,e in enumerate(vertexpairs)])
-	node_to_vertexpair = dict([(i+1, e) for i,e in enumerate(vertexpairs)])
-
-	medial = dict([(i+1, []) for i in range(len(vertexpairs))])
-
-	for u,vs in G_adj.items():
-		nodes = [vertexpair_to_node[tuple(sorted((u, v)))] for v in vs]
-		for i in range(len(nodes)):
-			medial[nodes[i]].append(nodes[(i-1)%len(nodes)])
-			medial[nodes[i]].append(nodes[(i+1)%len(nodes)])
-
-	M = Graph()
-	M.from_adj(medial)
-	# log.add("Medial graph: " + str(M))
-	# log.add("Node to vertexpair: " + str(node_to_vertexpair))
-	return M, node_to_vertexpair
+from Graph import Graph, read_lmg_from_stdin
 
 # assume G is simple
 # assume G is planar
 # assume G_adj is a rotation system
 # guarantee an edge of G have the same id as its corresponding medial node
 # guarantee M is a rotation system
-def medial_graph_2(G_adj: dict[int, list[int]]):
-	G = Graph()
-	G.from_adj(G_adj)
-
+def medial_graph(G: Graph):
 	M = Graph()
 	M.adj_edges = dict((abs(n), []) for n in G.edge_to_vertexpair.keys())
 
-	# print(M.adj_edges)
-
+	# Find faces
 	edges = [e for e in G.E()]
 	faces = []
 	while len(edges) > 0:
@@ -56,33 +26,30 @@ def medial_graph_2(G_adj: dict[int, list[int]]):
 			face.append(next_e)
 		faces.append(face)
 
+	# Add edges to medial graph
 	next_linkid = 1
 	for face in faces:
 		for i in range(len(face)):
[…]

### 2024-09-01 code/misc/branch_width_brute_force.py

@@ -1,9 +1,19 @@
-from parse_graph import parse_graph_to_adj
+def adj_from_stdin() -> dict[int, list[int]]:
+	adj = dict()
+	n = int(input())
+	for _ in range(n):
+		ys = list(map(int, input().split()))
+		x = ys[0]
+		adj[x] = []
+		for y in ys[1:]:
+			adj[x].append(y)
+	return adj
 
 # The branchwidth of G is the minimum width of any of its branch-decompositions.
 def branch_width(G):
 	Ts = branch_decompositions(G)
 	min_T = min(Ts, key=width_of_branch_decomposition)
+	print(min_T)
 	return width_of_branch_decomposition(min_T)
 
 # A branch-decomposition of a graph G is a tree T such that:
@@ -133,6 +143,6 @@ def enumerate_trees(leaves):
 
 		return new_trees
 
-adj = parse_graph_to_adj()
+adj = adj_from_stdin()
 
 print(branch_width(adj))

### 2024-09-01 code/misc/carving_width_brute_force.py

@@ -1,4 +1,4 @@
-from parse_graph import parse_graph_to_adj
+from util import parse_graph_to_adj
 
 G = parse_graph_to_adj()
 vertex_set = set(G.keys())

### 2024-09-01 code/misc/count_hamcyc_brute_force.py

@@ -1,7 +1,7 @@
 import itertools
-from parse_graph import parse_text_to_adj
+from util import adj_from_stdin
 
-G: dict[int, list[int]] = parse_text_to_adj()
+G: dict[int, list[int]] = adj_from_stdin()
 vertex_set = G.keys()
 N = len(vertex_set)
 

### 2024-09-01 code/misc/count_hamcyc_dp_thrice_prune.py

@@ -1,5 +1,5 @@
 import sys
-from parse_graph import parse_graph_to_adj
+from util import parse_graph_to_adj
 
 def find_faces(adj: dict[int, list[int]]) -> list[list[(int, int)]]:
 	# assume planar graph

### 2024-09-01 code/misc/create_png.py

@@ -1,7 +1,7 @@
 import matplotlib.pyplot as plt
 import networkx as nx
 import sys
-from code.parse_graph import parse_graph_to_nx
+from parse_graph import parse_graph_to_nx
 
 G = parse_graph_to_nx()
 

### 2024-09-01 code/misc/log.py

@@ -1,0 +1,24 @@
+import os
+import sys
+from Graph import Graph
+from util import print_adj
+
+def log_graph(graph: Graph, file_name: str):
+	if len(sys.argv) > 1:
+		graph_name = sys.argv[1]
+		print("Logging to ./", graph_name, "/", file_name, sep="")
+		if not os.path.exists(graph_name):
+			os.makedirs(graph_name)
+
+		with open(f"./{graph_name}/{file_name}", "w") as f:
+			f.write(graph.to_lmg())
+
+def log_adj(adj: dict[int, list[int | tuple[int, int]]], file_name: str):
+	if len(sys.argv) > 1:
+		graph_name = sys.argv[1]
+		print("Logging to ./", graph_name, "/", file_name, sep="")
+		if not os.path.exists(graph_name):
+			os.makedirs(graph_name)
+
+		with open(f"./{graph_name}/{file_name}", "w") as f:
+			f.write(print_adj(adj))

### 2024-09-01 code/misc/parse_graph.py

@@ -1,4 +1,5 @@
 import sys
+import networkx as nx
 
 def parse_bin_to_adj() -> dict[int, list[int]]:
 	adj = dict()
@@ -29,12 +30,6 @@ def adj_to_text(adj):
 	print(len(adj))
 	for v,xs in adj.items():
 		print(v, *xs)
-
-def adj_to_text_2(adj):
-	s = str(len(adj)) + "\n"
-	for v,xs in adj.items():
-		s += str(v) + " " + " ".join(map(str, xs)) + "\n"
-	return s
 
 def adj_to_nx(adj):
 	G = nx.MultiDiGraph()

### 2024-09-01 code/permute_vertices.py

@@ -1,15 +1,0 @@
-from parse_graph import parse_text_to_adj, adj_to_text
-adj = parse_text_to_adj()
-
-vertices = list(adj.keys())
-rotated = vertices[1:] + vertices[:1]
-
-mapping = dict([(vertices[i], rotated[i]) for i in range(len(vertices))])
-
-P = dict([(v, []) for v in vertices])
-
-for i in adj:
-	for j in adj[i]:
-		P[mapping[i]].append(mapping[j])
-
-adj_to_text(P)

### 2024-09-01 code/test-cubic-cross-validate.py

@@ -1,0 +1,51 @@
+import subprocess
+import threading
+
+expected = {
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{10}-5.in": "3",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{14}-5.in": "3",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{18}-5.in": "4",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{22}-5.in": "4",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{26}-5.in": "4",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{30}-5.in": "4",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{34}-5.in": "4",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{38}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{42}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{46}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{50}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{54}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{58}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{62}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{66}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{70}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{74}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{78}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{82}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{86}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{90}-5.in": "5",
+	f"python3 branch_width.py < ./graphs/random-planar-cubic/{94}-5.in": "5",
[…]

### 2024-09-01 code/test-planar-graphs-vis.py

@@ -1,0 +1,96 @@
+import sys
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from collections import defaultdict
+
+data = []
+
+# Data extraction from stdin
+for line in sys.stdin:
+    line = line.strip()
+    if line == "":
+        continue
+    tokens = line.split(";")
+    
+    # Parse the data
+    file = tokens[0]
+    n = int(tokens[1])
+    if tokens[2].startswith("Timeout"):
+        bw = -1
+    else:
+        bw = int(tokens[2])
+    bw_bound = int(tokens[3])
+    
+    # Append to data list
+    data.append({"file": file, "n": n, "bw": bw, "bw_bound": bw_bound})
+
+# Extracting values for plotting
+n_values = [entry["n"] for entry in data]
+bw_values = [entry["bw"] for entry in data]
+bw_bound_values = [entry["bw_bound"] for entry in data]
+
+# Count occurrences of (n, bw) pairs
+count_dict = defaultdict(int)
+for entry in data:
+    count_dict[(entry["n"], entry["bw"])] += 1
+
+# Create a list of counts for alpha and size calculation
+counts = [count_dict[(n, bw)] for n, bw in zip(n_values, bw_values)]
+
+# Handle empty counts list
+if counts:
+    max_count = max(counts)
+    min_count = min(counts)
+    # Normalize counts to range [0, 1]
+    normalized_counts = [(count - min_count) / (max_count - min_count) if max_count > min_count else 1 for count in counts]
+    # Scale normalized counts to range [0.3, 1.0] for alpha
+    alpha_values = [0.3 + 0.7 * alpha for alpha in normalized_counts]
+    
+    # Scale normalized counts to range [10, 1000] for size
+    sizes = [10 + 290 * alpha for alpha in normalized_counts]
+else:
+    # Default alpha values and sizes if counts list is empty
+    alpha_values = [0.3] * len(n_values)
+    sizes = [50] * len(n_values)
+
+# Separate values for coloring
+n_values_beyond_bound = [n for n, bw in zip(n_values, bw_values) if bw == -1]
+bw_values_beyond_bound = [bw for bw in bw_values if bw == -1]
+alpha_beyond_bound = [alpha_values[i] for i, bw in enumerate(bw_values) if bw == -1]
[…]

### 2024-09-01 code/test-planar-graphs.py

@@ -1,0 +1,50 @@
+import math
+import os
+import subprocess
+from subprocess import PIPE
+import time
+
+planar_graphs_dir = "./graphs/planar2/"
+program = "branch_width.py"
+
+files = os.listdir(planar_graphs_dir)
+files.sort()
+
+def run_command(file):
+	process = None
+	try:
+		command = f"python3.12 {program} < {planar_graphs_dir}{file}"
+		process = subprocess.Popen(command, shell=True, stdout=PIPE, stderr=PIPE)
+		stdout, stderr = process.communicate(timeout=2)
+		if stderr:
+			return f"Error encountered in {file}: {stderr.decode().strip()}"
+		return stdout.decode().strip()
+	except subprocess.TimeoutExpired:
+		if process:
+			process.terminate()
+			try:
+				process.wait(timeout=2)
+			except subprocess.TimeoutExpired:
+				process.kill()
+		return f"Timeout in {file}"
+	except Exception as e:
+		if process:
+			process.terminate()
+			try:
+				process.wait(timeout=5)
+			except subprocess.TimeoutExpired:
+				process.kill()
+		return f"Exception in {file}: {e}"
+
+for file in files:
+	if file <= "007-001957.in":
+		continue
+	st = time.time()
+	response = run_command(file)
+	elapsed_time = time.time() - st
+
+	formatted_time = time.strftime("%H:%M:%S", time.gmtime(elapsed_time)) + f".{int(elapsed_time * 1000) % 1000:03d}"
+	n = int(file.split('-')[0])
+	bw_bound = math.floor(4.5 * math.sqrt(n))
+
+	print(f"{file};{n};{response.lstrip().strip()};{bw_bound};{formatted_time}")

### 2024-09-01 code/util.py

@@ -1,0 +1,50 @@
+import sys
+
+def read_bin_to_adj() -> dict[int, list[int]]:
+	adj = dict()
+	n = ord(sys.stdin.buffer.read(1))
+	for i in range(1, n+1):
+		adj[i] = []
+	i = 1
+	while i <= n:
+		x = ord(sys.stdin.buffer.read(1))
+		if x == 0:
+			i += 1
+			continue
+		adj[i].append(x)
+	return adj
+
+def adj_from_stdin() -> dict[int, list[int]]:
+	adj = dict()
+	n = int(input())
+	for _ in range(n):
+		ys = list(map(int, input().split()))
+		x = ys[0]
+		adj[x] = []
+		for y in ys[1:]:
+			adj[x].append(y)
+	return adj
+
+def adj_to_text(adj):
+	print(len(adj))
+	for v,xs in adj.items():
+		print(v, *xs)
+
+def adj_to_str(adj):
+	s = str(len(adj)) + "\n"
+	for v,xs in adj.items():
+		s += str(v) + " " + " ".join(map(str, xs)) + "\n"
+	return s
+
+def adj_to_nx(adj):
+	G = nx.MultiDiGraph()
+	for v,xs in adj.items():
+		G.add_node(v)
+		for x in xs:
+			G.add_edge(v, x)
+	return G
+
+def adj_to_bytes(adj):
+	print(chr(len(adj)), end="")
+	for v,xs in adj.items():
+		print("".join(map(chr, [*xs])), end="\x00")

### 2024-09-01 code/util/andreas_to_bin.py

@@ -1,9 +1,0 @@
-import sys
-from code.parse_graph import adj_to_bytes
-
-adj = dict()
-
-for i,line in enumerate(sys.stdin):
-	adj[i+1] = list(map(lambda x: int(x)+1, line.strip().split()))
-
-adj_to_bytes(adj)

### 2024-09-01 code/util/time.sh

@@ -1,6 +1,0 @@
-python3 cubic_planar_gen.py 30 && time python3 count_hamcyc_faster.py < graphs2/blah.bin \
-&& python3 cubic_planar_gen.py 32 && time python3 count_hamcyc_faster.py < graphs2/blah.bin \
-&& python3 cubic_planar_gen.py 34 && time python3 count_hamcyc_faster.py < graphs2/blah.bin \
-&& python3 cubic_planar_gen.py 36 && time python3 count_hamcyc_faster.py < graphs2/blah.bin \
-&& python3 cubic_planar_gen.py 38 && time python3 count_hamcyc_faster.py < graphs2/blah.bin \
-&& python3 cubic_planar_gen.py 40 && time python3 count_hamcyc_faster.py < graphs2/blah.bin \

### 2024-09-01 code/util/validate_simple_cubic_planar.py

@@ -1,56 +1,0 @@
-import matplotlib.pyplot as plt
-import networkx as nx
-import sys
-from code.parse_graph import parse_graph_to_nx
-
-G = parse_graph_to_nx()
-
-def is_simple(G):
-	for v in G.nodes:
-		es = list(G.edges(v))
-		if len(es) <= 0:
-			print(es)
-			print(v)
-			exit()
-
-		# check for self loops
-		if (v, v) in es:
-			return False
-
-		# check for parallel edges
-		e = es[0]
-		for e2 in es[1:]:
-			if e2 == e:
-				return False
-			e = e2
-	return True
-
-def is_3_regular(G):
-	for v in G.nodes:
-		xs = list(G[v])
-		if len(xs) != 3:
-			return False
-	return True
-
-print("is simple" if is_simple(G) else "NOT simple")
-print("is cubic" if is_3_regular(G) else "NOT cubic")
-print("is planar" if nx.is_planar(G) else "NOT planar")
-
-if not (is_simple(G) and is_3_regular(G) and nx.is_planar(G)):
-	plt.figure()
-	pos = nx.planar_layout(G)
-	nx.draw_networkx_nodes(G, pos, node_color = 'r', node_size = 100, alpha = 1)
-	nx.draw_networkx_labels(G, pos, font_size=8)
-	ax = plt.gca()
-	for e in G.edges:
-		ax.annotate("",
-					xy=pos[e[0]], xycoords='data',
-					xytext=pos[e[1]], textcoords='data',
-					arrowprops=dict(arrowstyle="->", color="0.5",
-									shrinkA=5, shrinkB=5,
-									patchA=None, patchB=None,
-									connectionstyle="arc3,rad=rrr".replace('rrr',str(0.2*e[2])),
-									),
-					)
-	plt.axis('off')
-	plt.show()

### 2024-09-01 code/validate/is_connected.py

@@ -1,4 +1,4 @@
-from parse_graph import parse_text_to_adj
+from code.graphs.parse_graph import parse_text_to_adj
 
 G = parse_text_to_adj()
 

### 2024-09-01 code/validate/is_cubic.py

@@ -1,7 +1,7 @@
 import networkx as nx
-from parse_graph import parse_text_to_adj, adj_to_nx
+from util import adj_from_stdin, adj_to_nx
 
-G = adj_to_nx(parse_text_to_adj())
+G = adj_to_nx(adj_from_stdin())
 
 def is_3_regular(G):
 	for v in G.nodes:

### 2024-09-01 code/validate/is_planar.py

@@ -1,0 +1,5 @@
+import networkx as nx
+from util import adj_from_stdin, adj_to_nx
+
+G = adj_to_nx(adj_from_stdin())
+print(nx.is_planar(G))

### 2024-09-01 code/validate/is_simple.py

@@ -1,7 +1,7 @@
 import networkx as nx
-from parse_graph import parse_text_to_adj, adj_to_nx
+from util import adj_from_stdin, adj_to_nx
 
-G = adj_to_nx(parse_text_to_adj())
+G = adj_to_nx(adj_from_stdin())
 
 def is_simple(G):
 	for v in G.nodes:

### 2024-09-01 code/validate/is_undirected.py

@@ -1,6 +1,6 @@
-from parse_graph import parse_text_to_adj
+from util import adj_from_stdin
 
-G = parse_text_to_adj()
+G = adj_from_stdin()
 
 def is_undirected(G):
 	for x,ys in G.items():

### 2024-09-01 python-api/app.py

@@ -1,25 +1,0 @@
-from flask import Flask, request, jsonify
-
-app = Flask(__name__)
-
-countries = [
-	{"id": 1, "name": "Thailand", "capital": "Bangkok", "area": 513120},
-	{"id": 2, "name": "Australia", "capital": "Canberra", "area": 7617930},
-	{"id": 3, "name": "Egypt", "capital": "Cairo", "area": 1010408},
-]
-
-def _find_next_id():
-	return max(country["id"] for country in countries) + 1
-
-@app.get("/countries")
-def get_countries():
-	return jsonify(countries)
-
-@app.post("/countries")
-def add_country():
-	if request.is_json:
-		country = request.get_json()
-		country["id"] = _find_next_id()
-		countries.append(country)
-		return country, 201
-	return {"error": "Request must be JSON"}, 415

### 2024-09-02 paper/main.tex
Add graphs

@@
 \begin{center}
-	\includegraphics[width=0.9\textwidth]{images/dalle-pied-piper.png}
+	\includegraphics[width=0.9\textwidth]{images/dalle-pied-piper-2.png}
 \end{center}
@@
 
-	With subroutines for all three problems, the implementation obtains a branch decomposition like so.
+	With subroutines for all three problems, the implementation obtains a branch decomposition.
 
-	\pyfile{branch_decomposition.py}{19}{27}
+	\pyfile{branch_decomposition.py}{20}{25}
 
@@
 
-		\pyfile{medial_graph.py}{4}{54}
+		\pyfile{medial_graph.py}{4}{50}
 
@@
 
-			\pyfile{carving_width.py}{5}{6}
+			\pyfile{carving_width.py}{5}{5}
 			\pyfile{carving_width.py}{130}{156}
@@
 
-		The rat-catching algorithm decides $cw(M) \geq k$ with $k \in \mathbb{N}^+$.
+		The rat-catching algorithm decides $cw(M) \geq k$ with $M$ being a graph and $k \in \mathbb{N}^+$.
 
@@
 Considering a graph $M$, the edges of a face can be thought of as walls of a room and vertices as the corners of some rooms.
-The rat moves from corner to corner along the walls and the rat-catcher moves from room to room through some wall.
+The rat moves from corner to corner along the walls and the rat-catcher moves from room to wall to room again.
 The rat-catcher can force the rat away from some walls by making noise.
@@
 
-		Additionally, if $\Delta(M) \geq k$ then the rat wins. The argument for why this is true is glossed over in \ref{BGZ15}.
+		Additionally, if $\Delta(M) \geq k$ then the rat wins. The argument for why this is true, is glossed over in \cite{BGZ15}, but is important to include.
 
@@
 
+		An edge $e$ is called quiet iff. $e$ is not noisy.
+
 		\begin{definition}\label{def:edge-noisy}
@@
 		\end{definition}
-
-		An edge $e$ is called quiet iff. $e$ is not noisy.
 
@@
 
-		A \textit{quiet subgraph} $Q(M, k, e)$, for some graph $M$ and some noise level $k$ and some edge $e \in \mathbb{E}(M)$, is a subgraph of $M$ with the vertex set $V(Q(M, k, e)) = V(M)$ and the edge set
+		A \textit{quiet subgraph} $Q$, for some graph $M$ and some noise level $k$ and some edge $e \in \mathbb{E}(M)$, is a subgraph of $M$ with the vertex set $V(Q) = V(M)$ and the edge set
 
-		$\mathbb{E}(Q(M, k, e)) = \{e_1 \st $ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
+		$\mathbb{E}(Q) = \{e_1 \st $ every closed walk of $M^*$ containing $e_1^*$ and $e_2^*$ has length at least $k\}$
 
 		\begin{problem}\label{problem:noisy-subgraph}
-			Given a plane graph $M$ that might have parallel edges, an edge $e \in \mathbb{E}(M)$, and noise level $k \in \mathbb{N}^+$, output the quiet subgraph $Q(M, k, e)$.
+			Given a plane graph $M$ that might have parallel edges, an edge $e \in \mathbb{E}(M)$, and noise level $k \in \mathbb{N}^+$, output the quiet subgraph $Q$.
 		\end{problem}
@@
 
-		Let $F(M)$ be the set of faces of $M$.
-
 		Let $S$ be every possible state when the rat-catcher is in a face, some of which might be losing states. $S = \{ (f, v) \st v \in V(M) \land f$ is a face of $ M \}$.
 
-		Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \st e \in \mathbb{E}(M) \land C$ is a component of $Q(M, k, e) \}$.
+		Let $T$ be every possible state when the rat-catcher is on an edge. $T = \{ (e, C) \st e \in \mathbb{E}(M) \land C$ is a component of $Q \}$.
 
@@
 If every state is marked then the rat-catcher wins, otherwise the rat wins.
+Exactly how to propagate losing states is discussed below.
 
@@
 
-		Losing states (the tuples $(f,v) \in S$ where $v \in f$) are marked as losing.
+		The initial losing states (the tuples $(f,v) \in S$ where $v \in f$) are marked as losing.
 
@@
 
+		Apply the two rules above until either no more new losing states are marked, or until every state is marked as losing.
+
+		It is useful for the implementation to have a relation \mintinline{py}{edge_to_node} mapping the two half-edges of an edge to the two faces to either side of the edge, and a relation \mintinline{py}{node_to_face} to enumerate the faces.
+These relation are computed in the \mintinline{py}{dual_graph} function.
+
+		\pyfile{carving_width.py}{5}{6}
 		\pyfile{carving_width.py}{79}{128}
@@
 
-			Using the mapping from links to edges, and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained by applying the map.
+			Using the relation \mintinline{py}{link_to_edge} and the fact that an edge $e$ is called quiet iff. $e$ is not noisy, the quiet edges can be obtained by applying the relation.
 
@@
 
-			\pyfile{carving_width.py}{6}{74}
+			\pyfile{carving_width.py}{5}{74}
 
@@
 
+			The relations \mintinline{py}{node_to_face} and \mintinline{py}{edge_to_node} are useful for previously mentioned subproblems, and are computed here while enumerating the faces.
+
 			\pyfile{dual_graph.py}{4}{56}
@@
 	The first set of graphs are random cubic planar graphs.
-The second set of graphs are small simple planar graphs generated by \textit{Plantri}\footnote{\url{http://users.cecs.anu.edu.au/~bdm/plantri/}}\footnote{\url{https://github.com/mishun/plantri/tree/master}}.
+The second set of graphs are small simple planar graphs generated by \textit{Plantri}\footnote{\url{http://users.cecs.anu.edu.au/~bdm/plantri/}}\footnote{\url{https://github.com/mishun/plantri/tree/master}}.\cite{BM}
 
@@
 
-	Some of the graphs are suspected to be cause the implementation to go into an infinite loop, as the branch width is not computed within a reasonable time.
+	Some of the graphs are suspected to cause the implementation to go into an infinite loop, as the branch width is not computed within a reasonable time.
 So the implementation is suspectedly not correct for all planar graphs.
@@
 The red line is $\lfloor 4.5 \cdot \sqrt{n} \rfloor$.
-The blue dots are the branch width and the number directly next to it is the number of the graph with the same unique (n, bw) pair.}
+The blue dots are the branch width and the number directly next to it is the number of the graph with both the same number of vertices and the same branch width.}
 		\label{fig:bw-planar}
@@
 
+\section{Acknowledgements}
+	I would like to thank my supervisors, Thore Husfeldt and Andreas Björklund, for their guidance and support throughout this project.
+I would also like to thank Andreas Björklund for providing the random cubic planar graphs and values for cross-validation.
+I would like to thank my family and friends for their support and encouragement.
+
 \newpage
-\section{References}
 \printbibliography
@@
 	\pyfile{graphs/process_planar_code.py}{0}{1000}\label{graphs/process_planar_code.py}
-	\subsection{}
 \end{document}

### 2024-09-02 code/branch_decomposition.py

@@ -2,7 +2,7 @@ from typing import Union
 from Graph import Graph, read_lmg_from_stdin
 from medial_graph import medial_graph
 from carving_decomposition import carving_decomposition
-from util import adj_to_str
+from util import adj_from_stdin, adj_to_str
 
 def carving_decomposition_to_branch_decomposition(
 	cd: dict[int, list[int]],
@@ -22,11 +22,17 @@ def branch_decomposition(G: Graph) -> dict[int, list[Union[int, tuple[int, int]]
 	Gx, node_to_vertexpair = medial_graph(G)
 	cd = carving_decomposition(Gx.copy())
 	bd = carving_decomposition_to_branch_decomposition(cd, node_to_vertexpair)
-	print(adj_to_str(bd))
 	return bd
 
 if __name__ == "__main__":
+	# Uncomment for multigraphs
+	# G = Graph()
+	# G.from_lmg(read_lmg_from_stdin())
+	# bd = branch_decomposition(G)
+	# print("bd", bd)
+
+	adj = adj_from_stdin()
 	G = Graph()
-	G.from_lmg(read_lmg_from_stdin())
+	G.from_adj(adj)
 	bd = branch_decomposition(G)
-	print("bd", bd)
+	print(adj_to_str(bd))

### 2024-09-02 code/branch_width.py

@@ -38,6 +38,7 @@ def branch_width(G: Graph) -> int:
 	return bw
 
 if __name__ == "__main__":
+	# Uncomment for multigraphs
 	# G = Graph()
 	# G.from_lmg(read_lmg_from_stdin())
 	# bw = branch_width(G)

### 2024-09-02 code/carving_width.py

@@ -12,7 +12,7 @@ def carving_width(G: Graph) -> int:
 		s,t = D.edge_to_vertexpair[l]
 		links = link_to_edge.keys()
 
-		def dists(n: int) -> dict[int, int]:
+		def dists(n: int) -> dict[int, int]: # BFS
 			dist = {v: -1 for v in D.V()}
 			dist[n] = 0
 			queue = [n]
@@ -57,7 +57,7 @@ def carving_width(G: Graph) -> int:
 		components = []
 		unseen = set(quiet_subgraph.keys())
 
-		while len(unseen) > 0:
+		while len(unseen) > 0: # DFS
 			v = unseen.pop()
 			component = [v]
 			stack = [v]
@@ -106,12 +106,12 @@ def carving_width(G: Graph) -> int:
 
 		# Play the game
 		while True:
-			new_deletion = False
+			at_least_one_new_losing = False
 
 			for (e, C) in T:
 				if all([(edge_to_node[e], v) in losing_S for v in C]):
 					if (e, C) not in losing_T:
-						new_deletion = True
+						at_least_one_new_losing = True
 						losing_T.add((e, C))
 
 			for (e, C) in losing_T:
@@ -119,12 +119,12 @@ def carving_width(G: Graph) -> int:
 				f2 = edge_to_node[-e]
 				for (f, v) in [(f1, v) for v in C] + [(f2, v) for v in C]:
 					if (f, v) not in losing_S:
-						new_deletion = True
+						at_least_one_new_losing = True
 						losing_S.add((f, v))
 
 			if len(T) == len(losing_T) or len(S) == len(losing_S):
 				return False
-			elif not new_deletion:
+			elif not at_least_one_new_losing:
 				return True
 
 	def binary_search_cw():

### 2024-09-02 code/graphs/gen-cubic-planar.py

@@ -1,0 +1,8 @@
+import os
+
+ns = [
+	2, 3, 4, 5, 6,
+]
+
+for n in ns:
+	os.system(f"./plantri.bin -pm1c1 {n} | python3 process_planar_code.py planar")

### 2024-09-02 code/graphs/gen-planar.py

@@ -1,0 +1,11 @@
+import os
+
+# ns = [
+# 	2, 3, 4, 5, 6,
+# ]
+ns = [
+	7
+]
+
+for n in ns:
+	os.system(f"./plantri.bin -pm1c1 {n} | python3 process_planar_code.py planar2")

### 2024-09-02 code/graphs/gen-random-planar-cubic.py

@@ -1,0 +1,87 @@
+# Author: Andreas Björklund
+from random import random, seed
+import sys
+
+if len(sys.argv)>2:
+  NVERTICES=int(sys.argv[1])
+  RSEED=int(sys.argv[2])
+  seed(RSEED)
+elif len(sys.argv)>1:
+  NVERTICES=int(sys.argv[1])
+else:  
+  NVERTICES=6
+
+if NVERTICES%2==1:
+  sys.stderr.write("Number of vertices must be even\n")
+  exit(1)
+if NVERTICES<4:
+  sys.stderr.write("Number of vertices must be larger than 4\n")
+  exit(1)
+
+DEPTH=(NVERTICES-4)//2
+def start_instance():
+  # Original graph K_4  
+  #   0
+  #
+  #   3
+  #1     2
+  C=[[2,3,1], [0,3,2], [1,3,0], [0,2,1]] 
+  return C  
+                         
+# Generate random cubic 3-connected planar graph
+def random_instance(C):
+ 
+  # Triangle
+  D=[[0,4,5],[1,5,3],[2,3,4]]
+  
+  # Single Bridge
+  # 041
+  # 253
+  F=[[0,1,5],[2,4,3]]
+  
+  B = C[:]
+
+  for rounds in range(1):
+    n=len(B)
+    who = int(random()*n)
+   
+    nx = int(random()*3)
+    cycle=[who]
+    while B[who][nx]!=cycle[0]:
+      cycle.append(B[who][nx])
+      ix=[x for x in range(3) if B[B[who][nx]][x]==who]
+      ix=(ix[0]+2)%3
+      who = B[who][nx]
+      nx = ix
+  
+    m=len(cycle)
+    p = int(random()*(m-3))+2;
+    T = cycle[:2] + cycle[p:p-2:-1]
+    G = B[:]
+    H = [(list(map(sum, zip(x, [n-4,n-4,n-4])))) for x in F]
+    for i in range(4):
+      for ix2 in range(3):
+        if B[T[i]][ix2]==T[i^1]:
+          G[T[i]]= (G[T[i]][:ix2]+[n+(i // 2)]+G[T[i]][ix2+1:])
+      for j in range(len(F)):
+        for k in range(3):
+          if F[j][k]==i:
+            H[j] = (H[j][:k]+[T[i]]+H[j][k+1:])
+    B=G+H
+   
+  return B
+
+def gen_instance():
+  C=start_instance()
[…]

### 2024-09-02 code/graphs/process_planar_code.py

@@ -1,0 +1,45 @@
+import sys
+import os
+
+def read_bin_to_adj() -> dict[int, list[int]]:
+	adj = dict()
+	n = ord(sys.stdin.buffer.read(1))
+	for i in range(1, n+1):
+		adj[i] = []
+	i = 1
+	while i <= n:
+		x = ord(sys.stdin.buffer.read(1))
+		if x == 0:
+			i += 1
+			continue
+		adj[i].append(x)
+	return adj
+
+def adj_to_str(adj):
+	s = str(len(adj)) + "\n"
+	for v,xs in adj.items():
+		s += str(v) + " " + " ".join(map(str, xs)) + "\n"
+	return s
+
+# skip header
+byte_count = 0
+while byte_count < 15:
+	byte_s = sys.stdin.buffer.read(1)
+	if not byte_s:
+		break
+	byte_count += 1
+
+# read graphs
+graph_idx = 1
+while True:
+	adj = read_bin_to_adj()
+
+	folder_name = sys.argv[1]
+
+	# create file if not exists
+	if not os.path.exists(f"./{folder_name}"):
+		os.makedirs(f"./{folder_name}")
+
+	with open(f"./{folder_name}/{len(adj.keys()):03}-{graph_idx:06}.in", "w") as outfile:
+		outfile.write(adj_to_str(adj))
+	graph_idx += 1

### 2024-09-02 paper/images/k4/bd.tex
Add drawings

\begin{tikzpicture}
	\begin{scope}[xshift=6cm, yshift=0cm]
		\node[circle, draw, color=black, fill=white] (b1) at (3, 4) {\{1,3\}};
		\node[circle, draw, color=black, fill=white] (b3) at (-0.5, 3) {\{1,4\}};
		\node[circle, draw, color=black, fill=white] (10) at (1.5, 4) {10};
		\node[circle, draw, color=black, fill=white] (9) at (1, 3) {9};
		\node[circle, draw, color=black, fill=white] (8) at (1.5, 2) {8};
		\node[circle, draw, color=black, fill=white] (7) at (1, 1) {7};
		\node[circle, draw, color=black, fill=white] (b2) at (3, 2) {\{2,4\}};
		\node[circle, draw, color=black, fill=white] (b4) at (-0.5, 1) {\{2,3\}};
		\node[circle, draw, color=black, fill=white] (b5) at (1, -0.5) {\{3,4\}};
		\node[circle, draw, color=black, fill=white] (b6) at (1.5, 5.5) {\{1,2\}};

		% Draw the edges between nodes with labels
		\draw[line width=0.5mm, color=black] (10) -- (b6);
		\draw[line width=0.5mm, color=black] (b1) -- (10);
		\draw[line width=0.5mm, color=black] (b3) -- (9);
		\draw[line width=0.5mm, color=black] (10) -- (9);
		\draw[line width=0.5mm, color=black] (9) -- (8);
		\draw[line width=0.5mm, color=black] (b2) -- (8);
		\draw[line width=0.5mm, color=black] (8) -- (7);
		\draw[line width=0.5mm, color=black] (7) -- (b4);
		\draw[line width=0.5mm, color=black] (7) -- (b5);

		\node (G) at (3, -0.25) {\{2,3,4\}};

		\node[opacity=0] at (-0.5, 2) (start) {};
		\node[opacity=0] at (2.5, 0) (end) {};
		\draw[dotted, thick] (start) .. controls +(2,0) and +(0,0.5) .. (end);
	\end{scope}
\end{tikzpicture}

### 2024-09-02 paper/images/k4/cd.tex
Add drawings

--- 2024-09-02 paper/images/k4/bd.tex
@@
 	\begin{scope}[xshift=6cm, yshift=0cm]
-		\node[circle, draw, color=black, fill=white] (b1) at (3, 4) {\{1,3\}};
-		\node[circle, draw, color=black, fill=white] (b3) at (-0.5, 3) {\{1,4\}};
-		\node[circle, draw, color=black, fill=white] (10) at (1.5, 4) {10};
-		\node[circle, draw, color=black, fill=white] (9) at (1, 3) {9};
-		\node[circle, draw, color=black, fill=white] (8) at (1.5, 2) {8};
-		\node[circle, draw, color=black, fill=white] (7) at (1, 1) {7};
-		\node[circle, draw, color=black, fill=white] (b2) at (3, 2) {\{2,4\}};
-		\node[circle, draw, color=black, fill=white] (b4) at (-0.5, 1) {\{2,3\}};
-		\node[circle, draw, color=black, fill=white] (b5) at (1, -0.5) {\{3,4\}};
-		\node[circle, draw, color=black, fill=white] (b6) at (1.5, 5.5) {\{1,2\}};
+		\node[circle, draw, color=black, fill=white] (b1) at (3, 4) {2};
+		\node[circle, draw, color=black, fill=white] (b3) at (-0.5, 3) {3};
+		\node[circle, draw, color=black, fill=white] (10) at (1.5, 4) {};
+		\node[circle, draw, color=black, fill=white] (9) at (1, 3) {};
+		\node[circle, draw, color=black, fill=white] (8) at (1.5, 2) {};
+		\node[circle, draw, color=black, fill=white] (7) at (1, 1) {};
+		\node[circle, draw, color=black, fill=white] (b2) at (3, 2) {4};
+		\node[circle, draw, color=black, fill=white] (b4) at (-0.5, 1) {5};
+		\node[circle, draw, color=black, fill=white] (b5) at (1, -0.5) {6};
+		\node[circle, draw, color=black, fill=white] (b6) at (1.5, 5.5) {1};
 
@@
 
-		\node (G) at (3, -0.25) {\{2,3,4\}};
+		\node (G) at (3, -0.25) {\{2,3,4,6,7,9\}};
 

### 2024-09-02 paper/images/k4/quiet/k5/1.tex
Add drawings

\begin{tikzpicture}
	% Define the nodes with specific positions
	\begin{scope}[xshift=0cm, yshift=0cm, scale=0.4]
		\node[circle, draw] (1) at (8, -0) {1};
		\node[circle, draw] (2) at (3, -8) {2};
		\node[circle, draw] (3) at (4, -3.5) {3};
		\node[circle, draw] (4) at (3, -1.5) {4};
		\node[circle, draw] (5) at (-2, -0) {5};
		\node[circle, draw] (6) at (2, -3.5) {6};

		% Draw the edges between nodes with labels
		\draw[dashed, line width=3] (5) -- node[below left] {1} (6);
		\draw[dashed] (5) -- node[above right] {2} (4);
		\draw[dashed] (4) -- node[above left] {3} (6);
		\draw[dashed] (3) -- node[below] {4} (6);
		\draw[draw, line width=1] (3) -- node[right] {5} (2);
		\draw[dashed] (6) -- node[left] {6} (2);
		\draw[dashed] (2) -- node[below left] {7} (5);
		\draw[draw, line width=1] (1) -- node[below right] {8} (2);
		\draw[dashed] (1) -- node[above] {9} (5);
		\draw[draw, line width=1] (1) -- node[above left] {10} (4);
		\draw[draw, line width=1] (1) -- node[below right] {11} (3);
		\draw[draw, line width=1] (4) -- node[above right] {12} (3);
	\end{scope}
\end{tikzpicture}

### 2024-09-02 paper/images/k4/quiet/k5/11.tex
Add drawings

--- 2024-09-02 paper/images/k4/quiet/k5/1.tex
@@
 		% Draw the edges between nodes with labels
-		\draw[dashed, line width=3] (5) -- node[below left] {1} (6);
-		\draw[dashed] (5) -- node[above right] {2} (4);
-		\draw[dashed] (4) -- node[above left] {3} (6);
+		\draw[draw, line width=1] (5) -- node[below left] {1} (6);
+		\draw[draw, line width=1] (5) -- node[above right] {2} (4);
+		\draw[draw, line width=1] (4) -- node[above left] {3} (6);
 		\draw[dashed] (3) -- node[below] {4} (6);
-		\draw[draw, line width=1] (3) -- node[right] {5} (2);
-		\draw[dashed] (6) -- node[left] {6} (2);
-		\draw[dashed] (2) -- node[below left] {7} (5);
-		\draw[draw, line width=1] (1) -- node[below right] {8} (2);
+		\draw[dashed] (3) -- node[right] {5} (2);
+		\draw[draw, line width=1] (6) -- node[left] {6} (2);
+		\draw[draw, line width=1] (2) -- node[below left] {7} (5);
+		\draw[dashed] (1) -- node[below right] {8} (2);
 		\draw[dashed] (1) -- node[above] {9} (5);
-		\draw[draw, line width=1] (1) -- node[above left] {10} (4);
-		\draw[draw, line width=1] (1) -- node[below right] {11} (3);
-		\draw[draw, line width=1] (4) -- node[above right] {12} (3);
+		\draw[dashed] (1) -- node[above left] {10} (4);
+		\draw[dashed, line width=3] (1) -- node[below right] {11} (3);
+		\draw[dashed] (4) -- node[above right] {12} (3);
 	\end{scope}

### 2024-09-02 paper/images/k4/quiet/k5/12.tex
Add drawings

--- 2024-09-02 paper/images/k4/quiet/k5/11.tex
@@
 		\draw[draw, line width=1] (5) -- node[below left] {1} (6);
-		\draw[draw, line width=1] (5) -- node[above right] {2} (4);
-		\draw[draw, line width=1] (4) -- node[above left] {3} (6);
+		\draw[dashed] (5) -- node[above right] {2} (4);
+		\draw[dashed] (4) -- node[above left] {3} (6);
 		\draw[dashed] (3) -- node[below] {4} (6);
@@
 		\draw[draw, line width=1] (2) -- node[below left] {7} (5);
-		\draw[dashed] (1) -- node[below right] {8} (2);
-		\draw[dashed] (1) -- node[above] {9} (5);
+		\draw[draw, line width=1] (1) -- node[below right] {8} (2);
+		\draw[draw, line width=1] (1) -- node[above] {9} (5);
 		\draw[dashed] (1) -- node[above left] {10} (4);
-		\draw[dashed, line width=3] (1) -- node[below right] {11} (3);
-		\draw[dashed] (4) -- node[above right] {12} (3);
+		\draw[dashed] (1) -- node[below right] {11} (3);
+		\draw[dashed, line width=3] (4) -- node[above right] {12} (3);
 	\end{scope}

### 2024-09-02 paper/images/k4/zcontract1.tex
Add drawings

\begin{tikzpicture}
	\node (G) at (2, 1) {Contract $(5,6)$};
	% Define the nodes with specific positions
	\begin{scope}[xshift=0cm, yshift=0cm, scale=0.6]
		\node[circle, draw] (1) at (8, -0) {1};
		\node[circle, draw] (2) at (3, -8) {2};
		\node[circle, draw] (3) at (4, -3.5) {3};
		\node[circle, draw] (4) at (3, -1.5) {4};
		\node[circle, draw] (7) at (-2, -0) {7};

		% Draw the edges between nodes with labels
		\draw[draw] (7) to[bend left=5] node[above right] {2} (4);
		\draw[draw] (4) to[bend left=5] node[above left] {3} (7);
		\draw[draw] (3) -- node[below] {4} (7);
		\draw[draw] (3) -- node[right] {5} (2);
		\draw[draw] (7) to[bend left=5] node[left] {6} (2);
		\draw[draw] (2) to[bend left=5] node[below left] {7} (7);
		\draw[draw] (1) -- node[below right] {8} (2);
		\draw[draw] (1) -- node[above] {9} (7);
		\draw[draw] (1) -- node[above left] {10} (4);
		\draw[draw] (1) -- node[below right] {11} (3);
		\draw[draw] (4) -- node[above right] {12} (3);
	\end{scope}
\end{tikzpicture}

### 2024-09-02 paper/images/k4/zcontract2.tex
Add drawings

--- 2024-09-02 paper/images/k4/zcontract1.tex
@@
 \begin{tikzpicture}
-	\node (G) at (2, 1) {Contract $(5,6)$};
+	\node (G) at (2, 1) {Contract $(7,4)$};
 	% Define the nodes with specific positions
@@
 		\node[circle, draw] (3) at (4, -3.5) {3};
-		\node[circle, draw] (4) at (3, -1.5) {4};
-		\node[circle, draw] (7) at (-2, -0) {7};
+		\node[circle, draw] (8) at (-2, -0) {8};
 
 		% Draw the edges between nodes with labels
-		\draw[draw] (7) to[bend left=5] node[above right] {2} (4);
-		\draw[draw] (4) to[bend left=5] node[above left] {3} (7);
-		\draw[draw] (3) -- node[below] {4} (7);
+		\draw[draw] (3) to[bend left=5] node[below] {4} (8);
 		\draw[draw] (3) -- node[right] {5} (2);
-		\draw[draw] (7) to[bend left=5] node[left] {6} (2);
-		\draw[draw] (2) to[bend left=5] node[below left] {7} (7);
+		\draw[draw] (8) to[bend left=5] node[left] {6} (2);
+		\draw[draw] (2) to[bend left=5] node[below left] {7} (8);
 		\draw[draw] (1) -- node[below right] {8} (2);
-		\draw[draw] (1) -- node[above] {9} (7);
-		\draw[draw] (1) -- node[above left] {10} (4);
+		\draw[draw] (1) to[bend right=5] node[above] {9} (8);
+		\draw[draw] (1) to[bend left=5] node[above left] {10} (8);
 		\draw[draw] (1) -- node[below right] {11} (3);
-		\draw[draw] (4) -- node[above right] {12} (3);
+		\draw[draw] (8) to[bend left=5] node[above right] {12} (3);
 	\end{scope}

### 2024-09-02 paper/images/k4/zcontract3.tex
Add drawings

--- 2024-09-02 paper/images/k4/zcontract2.tex
@@
 \begin{tikzpicture}
-	\node (G) at (2, 1) {Contract $(7,4)$};
+	\node (G) at (2, 1) {Contract $(8,3)$};
 	% Define the nodes with specific positions
@@
 		\node[circle, draw] (2) at (3, -8) {2};
-		\node[circle, draw] (3) at (4, -3.5) {3};
-		\node[circle, draw] (8) at (-2, -0) {8};
+		\node[circle, draw] (9) at (-2, -0) {9};
 
 		% Draw the edges between nodes with labels
-		\draw[draw] (3) to[bend left=5] node[below] {4} (8);
-		\draw[draw] (3) -- node[right] {5} (2);
-		\draw[draw] (8) to[bend left=5] node[left] {6} (2);
-		\draw[draw] (2) to[bend left=5] node[below left] {7} (8);
+		\draw[draw] (9) -- node[right] {5} (2);
+		\draw[draw] (9) to[bend left=5] node[left] {6} (2);
+		\draw[draw] (2) to[bend left=5] node[below left] {7} (9);
 		\draw[draw] (1) -- node[below right] {8} (2);
-		\draw[draw] (1) to[bend right=5] node[above] {9} (8);
-		\draw[draw] (1) to[bend left=5] node[above left] {10} (8);
-		\draw[draw] (1) -- node[below right] {11} (3);
-		\draw[draw] (8) to[bend left=5] node[above right] {12} (3);
+		\draw[draw] (1) to[bend right=5] node[above] {9} (9);
+		\draw[draw] (1) to[bend left=5] node[above left] {10} (9);
+		\draw[draw] (1) -- node[below right] {11} (9);
 	\end{scope}

### 2024-09-02 paper/main.tex
Add drawings

@@
 
+	The tetrahedral graph is a simple connected planar graph with 4 vertices and 6 edges. It is also know as the complete graph on 4 vertices $K_4$. The tetrahedral graph is used as a running example.
+
+	\begin{figure}[H]
+		\centering
+		% \includegraphics[width=\textwidth/2]{./images/k4/gm.png}
+		\input{./images/k4/g.tex}
+		\caption{The tetrahedral graph}
+		\label{fig:tetrahedral}
+	\end{figure}
+
 \section{The algorithm}
@@
 
+	\begin{figure}[H]
+		\centering
+		\input{images/k4/bd.tex}
+		\caption{Minimum Branch Decomposition of the Tetrahedral Graph with a branch width of 3.}
+		\label{fig:bd}
+	\end{figure}
+
 	The next 3 subsections describe the implementation of the subroutines for the problems \ref{problem:medial}, \ref{problem:mcd}, and \ref{problem:mcd-to-mbd}.
@@
 
+		\begin{figure}[H]
+			\centering
+			\input{./images/k4/gmedial.tex}
+			\caption{The Tetrahedral Graph (in black) and Medial graph of the Tetrahedral Graph (in blue)}
+			\label{fig:gmedial}
+		\end{figure}
+
 	\subsection{Minimum Branch Decomposition from Minimum Carving Decomposition}
@@
 		\pyfile{branch_decomposition.py}{6}{17}
+
+		\begin{figure}[H]
+			\centering
+			\input{images/k4/bd.tex}
+			\input{images/k4/cd.tex}
+			\caption{Minimum Branch Decomposition (left) of the Tetrahedral Graph and a Minimum Carving Decomposition (right) of the Medial Graph.}
+			\label{fig:m}
+		\end{figure}
 
@@
 		\pyfile{carving_decomposition.py}{7}{79}
+
+		\begin{figure}[H]
+			\centering
+			\input{./images/k4/m.tex}
+			\input{./images/k4/zcontract1.tex}
+			\input{./images/k4/zcontract2.tex}
+			\input{./images/k4/zcontract3.tex}
+			\input{images/k4/cd.tex}
+			\caption{A series of contractions of the medial graph of the Tetrahedral Graph, with nonincreasing carving width.
+All graphs have a carving width of $6$.
+First contracting $(5,6)$ into $7$, then $(7,4)$ into $8$, and finally $(8,3)$ into $9$.
+Then assembling the series of contractions into a minimum carving decomposition by introducing vertex $10$.}
+			\label{fig:m}
+		\end{figure}
 
@@
 
+			\begin{figure}[H]
+				\centering
+				\input{./images/k4/m.tex}
+				\input{./images/k4/zcontract1.tex}
+				\caption{A contraction of the medial graph of the Tetrahedral Graph given $(5,6)$ producing a new vertex $7$.}
+				\label{fig:m}
+			\end{figure}
+
 		\subsubsection{Carving Width}
@@
 			\pyfile{carving_width.py}{130}{156}
+
+			For the medial of the tetrahedral graph, the linear search gives for \mintinline{py}{rat_wins} for $k=0,1,2,3,4,5,6,7$ gives $True, True, True, True, True, True, True, False$, therefore the carving width of the medial of the tetrahedral graph is $6$.
 
@@
 
+		\subsection{Dual graph}\label{impl:dual}
+
+			Solving problem \ref{problem:dual}.
+
+			No other path of the implementation needs the assumption that the dual is planar, therefore the output doesn't need to be a rotation system. This simplifies the implementation.
+
+			The dual has a vertex for each face of the input graph.
+The faces are found by selecting an unmarked half-edge, then marking all the edges of the face it belongs to, and repeating this until all half-edges are marked.
+
+			The next halfedge $e_{i+1}$ after the current halfedge $e_i = \{u, v\}$ is the edge just before $-e_i$ in the neighborhood list for vertex $v$.
+
+			The relations \mintinline{py}{node_to_face} and \mintinline{py}{edge_to_node} are useful for previously mentioned subproblems, and are computed here while enumerating the faces.
+
+			\pyfile{dual_graph.py}{4}{56}
+
+			\begin{figure}[H]
+				\centering
+				\input{./images/k4/dual.tex}
+				\caption{The medial graph of the Tetrahedral Graph and its dual. Vertex 5 is the vertex for the outerface and therefore visually repeated.}
+				\label{fig:m}
+			\end{figure}
+
+		\subsection{Does the rat win?}
+
 			Solving problem \ref{problem:cw}.
@@
 			\pyfile{carving_width.py}{79}{128}
+
+			\begin{figure}[H]
+				\centering
+				\input{./images/k4/quiet/k5/1.tex}
+				\input{./images/k4/quiet/k5/2.tex}
+				\input{./images/k4/quiet/k5/3.tex}
+				\input{./images/k4/quiet/k5/4.tex}
+				\input{./images/k4/quiet/k5/5.tex}
+				\input{./images/k4/quiet/k5/6.tex}
+				\input{./images/k4/quiet/k5/7.tex}
+				\input{./images/k4/quiet/k5/8.tex}
+				\input{./images/k4/quiet/k5/9.tex}
+				\input{./images/k4/quiet/k5/10.tex}
+				\input{./images/k4/quiet/k5/11.tex}
+				\input{./images/k4/quiet/k5/12.tex}
+				\caption{The medial of the Tetrahedral Graph and the quiet components for noise level $k=5$, given the rat-catcher is on edge $1,...,12$.}
+				\label{fig:quiet}
+			\end{figure}
 
@@
 
-		\subsection{Dual graph}\label{impl:dual}
-
-			Solving problem \ref{problem:dual}
-
-			No other path of the implementation needs the assumption that the dual is planar, therefore the output doesn't need to be a rotation system. This simplifies the implementation.
-
-			The dual has a vertex for each face of the input graph.
-The faces are found by selecting an unmarked half-edge, then marking all the edges of the face it belongs to, and repeating this until all half-edges are marked.
-
-			The next halfedge $e_{i+1}$ after the current halfedge $e_i = \{u, v\}$ is the edge just before $-e_i$ in the neighborhood list for vertex $v$.
-
-			The relations \mintinline{py}{node_to_face} and \mintinline{py}{edge_to_node} are useful for previously mentioned subproblems, and are computed here while enumerating the faces.
-
-			\pyfile{dual_graph.py}{4}{56}
+			\begin{figure}[H]
+				\centering
+				\input{images/k4/quiet/quiet-k4-k5-e9.tex}
+				\caption{The quiet component for edge $9$ with noise level $k=5$.
+The quiet edges are in solid and the noisy edges are dashed.
+Vertex 5 is the vertex for the outerface and therefore visually repeated.}
+				\label{fig:quiet}
+			\end{figure}
 
